---
title: "FinalRecruitment_Results"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries 
```{r, echo=FALSE, message=FALSE}
library(patchwork)
library(tidyverse)
library(here)
library(PNWColors)
library(agricolae)
library(lme4)
library(lmerTest)
library(modelsummary)
library(performance)
library(purrr)
library(nlstools)
library(stringr)
library(emmeans)
library(modelr)
```

## Bring in datasets
```{r, echo=FALSE, message=FALSE}
RecruitmentCounts <- read_csv(here::here("Data","Recruitment_Tiles_Counts.csv"))
## raw tile counts/data

AllChemData_August <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/August2021/Allbiogeochemdata_QC.csv")

AllChemData_March <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/March2022/Allbiogeochemdata_QC_march_fdom2.csv")

pH <- read_csv(here::here("Data", "pHSlope.csv"))

EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv")) 
## this df has sum total recruit counts along with dead and alive counts and all mean, max, CV, and min environmental data 

Poc_cover_data <- read_csv(here::here("Data", "Species_Composition_2022_Danielle.csv"))
## Danielle Barnas' benthic taxa data 

PercentCover <- read_csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Full_Metadata.csv")
## Danielle Barnas' percent cover data 

PercentCover3 <- read_csv(here::here("Data", "PercentCover3.csv"))
##my updated percent cover data filtering out just relevant info 

RecruitmentCounts2.1 <- read_csv((here::here("Data", "RecruitmentCounts2.1.csv")))
## just the totaled recruit counts 

EnvData_wPocCover <- read_csv(here::here("Data", "EnvData_wPocCover.csv"))
## FULL dataset with pocillopora cover data, env data (nuts, etc), and recruitment counts 

```

## Create dataframe that summarize all recruit counts JUST for Poc 
-  use this df for analysis bc ONLY looking at poc recruits

```{r}
RecruitmentCounts_Poconly <- RecruitmentCounts %>% 
  group_by(CowTagID, Location) %>% 
  dplyr::summarise(sum_total = sum(Pocilloporidae), 
            sum_total_dead = sum(Dead), 
            sum_total_alive=sum(sum_total-sum_total_dead))
```

### Look at the diversity of non-Poc recruits on the tiles
**Results**: basically exactly the same Porites and Acropora densities found at both sites 

```{r}

nonPoc_recruits <- RecruitmentCounts %>% 
  group_by(CowTagID, Location) %>% 
  dplyr::summarise(total_Por = sum(Poritidae), 
                   total_Acrop = sum(Acroporidae), 
                   sum_total_nonPoc = sum(Poritidae+Acroporidae+Other))


nonPoc_recruits_bysite <- nonPoc_recruits %>% 
  group_by(Location) %>% 
  dplyr::summarise(totalPor = sum(total_Por), 
                   totalAcrop = sum(total_Acrop),
                   totalnonPoc = sum(sum_total_nonPoc))
  
```




## Figure 1: Poc Cover Data 
Plot: percent cover on the x, total recruit counts on the y, faceted by location 
-  shown on same plot to show differences between sites, however could also separate onto different plots 

-  Note: as of August 16th, 2024 - confirmed that missing P. acuta data for some CowTagIDs should be equal to 0 
-  finding: reference the Poc_Cover_Data script for the tidying and creation of df 
-  

Stats: 
-  run model comparison between linear and nonlinear 

**Results**: SEE POC_COVER_DATA MARKDOWN FOR RESULTS AND MODELS 

```{r}

recruits_by_PocCover <- EnvData_wPocCover %>% 
  ggplot(aes(x=percent_cover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              fill="cornsilk3", 
              alpha=0.2) + ## alpha adjusts the transparency 
  labs(x="% Cover of P. acuta Species found near tile location", 
       y="Logged Total of Recruits")

recruits_by_PocCover
ggsave(here::here("Outputs", "RecruitmentTiles", "Results", "Prelim", "recruits_by_PocCover.jpg"))


### run models here 


```


### Silicate with recruit data 
Plot: average silicate concentrations on the x, *logged* total recruit counts on the y, faceted by location 
-  shown on same plot to show differences between sites, however could also separate onto different plots 

Stats: 
-  run model comparison between linear and nonlinear
-  model: 

**Results**: 

```{r}
###############################
## plot 
###############################
recruits_by_silicate <- EnvData_wPocCover %>% 
  ggplot(aes(x=silicate_avg, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm,
              formula="y~poly(x,2)",
              fill="cornsilk3", 
              alpha=0.2) + ## alpha adjusts the transparency 
  labs(x="Avg Silicate", 
       y="Logged Total of Recruits")

recruits_by_silicate

#################
## join two plots together (silicate data and Poc cover data) into same plot 
recruits_by_silicate + recruits_by_PocCover
ggsave(here::here("Outputs", "RecruitmentTiles", "Results", "Prelim", "PocCoverANDSilicate.jpg"), 
       width=10, height=7)


###############################
## Statistics  
###############################

##### linear model 
silicate_recruit_model <- lm(log(sum_total+1) ~ silicate_avg*Location, data=EnvData_wPocCover)

anova(silicate_recruit_model)
summary(silicate_recruit_model)


##### nonlinear model 

silicate_recruit_model_nonlin <- lm(log(sum_total+1) ~ poly(silicate_avg,2)*Location, data=EnvData_wPocCover)

anova(silicate_recruit_model_nonlin)
summary(silicate_recruit_model_nonlin)


```


### Compare Poc cover vs Silicate to see if there is a relationship 
Plot: percent cover on the x, total recruit counts on the y, faceted by location 
- shown on same plot to show differences between sites, however could also separate onto different plots 

Stats: ran two separate models (bc sites are so different)
-  important note: there is missing percent cover data for some of the rebar/cowtags 

**Results**: 
-  need to check re-check models but looks like there is a strong nonlinear relationship at Varari but not Cabral 

```{r}
#######################
## plot 
######################

##### plot by site 
PocCover_vs_Silicate <- EnvData_wPocCover %>% 
  ggplot(aes(x=silicate_avg, 
             y=percent_cover, 
             color=Location)) + 
  geom_point(size=2) + 
  theme_classic() + 
  facet_wrap(~Location, scales="free") + 
  geom_smooth(method=lm, 
              formula="y~poly(x,2)",
              fill="cornsilk3", 
              alpha=0.2) ## alpha adjusts the transparency 
  
PocCover_vs_Silicate
ggsave(here::here("Outputs", "RecruitmentTiles", "Results", "Prelim", "silicate_by_PocCover.jpg"), 
       width=10, height=7)

############
## try plotting just as one site and with one geom_smooth 

PocCover_vs_Silicate2 <- EnvData_wPocCover %>% 
  ggplot(aes(x=silicate_avg, 
             y=percent_cover)) + 
  geom_point(size=2, aes(color=Location)) + 
  theme_classic() + 
 # facet_wrap(~Location, scales="free") + 
  geom_smooth(method=lm, 
              formula="y~poly(x,2)",
              fill="cornsilk3", 
              alpha=0.2) ## alpha adjusts the transparency 
  
PocCover_vs_Silicate2

ggsave(here::here("Outputs", "RecruitmentTiles", "Results", "Prelim", "silicate_by_PocCover.jpg"), 
       width=10, height=7)

```


-  run models (run sites separately and try a multiple regresssion )
```{r}
## Varari 
PocCover_vs_Silicate_Mregression_V <- lm(log(sum_total+1) ~ poly(silicate_avg, 2) + poly(percent_cover,2), data=EnvData_wPocCover %>% 
                                   filter(Location=="Varari") %>%  
                                   drop_na(percent_cover))

anova(PocCover_vs_Silicate_Mregression_V)
summary(PocCover_vs_Silicate_Mregression_V)

## Cabral
PocCover_vs_Silicate_Mregression_C <- lm(log(sum_total+1) ~ poly(silicate_avg, 2) + poly(percent_cover,2), data=EnvData_wPocCover %>% 
                                   filter(Location=="Cabral") %>%  
                                   drop_na(percent_cover))

anova(PocCover_vs_Silicate_Mregression_C)
summary(PocCover_vs_Silicate_Mregression_C)

################
### something is still going on with poly function so manually add percent cover squared column 
## basic regression 

EnvData_wPocCover2 <- EnvData_wPocCover %>% 
  mutate(percent_cover2 = percent_cover^2)

### for Varari 

### manual polyx,2
PocCover_vs_Silicate_model_V <- lm(silicate_avg~ percent_cover + percent_cover2, data=EnvData_wPocCover2 %>% 
                                   filter(Location=="Varari") %>%  
                                   drop_na(percent_cover))


### trouble with poly(x,2)
PocCover_vs_Silicate_model_V <- lm(silicate_avg~ poly(percent_cover,2), data=EnvData_wPocCover %>% 
                                   filter(Location=="Varari") %>% 
                                     drop_na(percent_cover))

anova(PocCover_vs_Silicate_model_V)
summary(PocCover_vs_Silicate_model_V)

### for Cabral 

### manual polyx,2
PocCover_vs_Silicate_model_C <- lm(silicate_avg~ percent_cover + percent_cover2, data=EnvData_wPocCover2 %>% 
                                   filter(Location=="Cabral") %>% 
                                     drop_na(percent_cover))

### trouble with poly(x,2)
PocCover_vs_Silicate_model_C <- lm(silicate_avg~ poly(percent_cover,2), data=EnvData_wPocCover %>% 
                                   filter(Location=="Cabral") %>% 
                                     drop_na(percent_cover))

anova(PocCover_vs_Silicate_model_C)
summary(PocCover_vs_Silicate_model_C)


##########################
## check residuals and plot normality 
##########################

#### plot models 
plot(PocCover_vs_Silicate_model_V)
plot(PocCover_vs_Silicate_model_C)


#### look at residuals 

## varai 
residuals_V <- resid(PocCover_vs_Silicate_model_V)
fitted_values_V <- fitted(PocCover_vs_Silicate_model_V)

plot(fitted_values_V, residuals_V)

## cabral 
residuals_C <- resid(PocCover_vs_Silicate_model_C)
fitted_values_C <- fitted(PocCover_vs_Silicate_model_C)

plot(fitted_values_C, residuals_C)

```



### Run a t-test to compare Poc cover with settlement and show in a boxplot 
-  compare between Location so x-axis is Location and y axis is mean settlement OR mean Poc cover (separate analyses) 
-  edit df to create mean column 
-  **Results**: there is no significant difference in the means b/w sites for either settlement or Poc cover

```{r}

EnvData_wPocCover_ttest <- EnvData_wPocCover %>% 
  select(!c("sum_total_dead":"total_count_biotics")) %>% 
  group_by(Location) %>% 
  mutate(log_settled = log(sum_total+1), 
         sq_percentcover = sqrt(percent_cover)) %>% 
  mutate(mean_settled = mean(log_settled), 
         mean_cover = mean(sq_percentcover))

##### settlement 
Settlement_Ttest<-t.test(log(sum_total+1)~Location, data=EnvData_wPocCover_ttest, na.rm=TRUE)

Settlement_Ttest

##### cover  
PocCover_Ttest<-t.test(sq_percentcover~Location, data=EnvData_wPocCover_ttest, na.rm=TRUE)

PocCover_Ttest

##### boxplot 
Settlement_boxplot <- EnvData_wPocCover_ttest %>% 
  ggplot(aes(x=Location, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_boxplot() + 
  theme_classic() 

Settlement_boxplot


PocCover_boxplot <- EnvData_wPocCover_ttest %>% 
  ggplot(aes(x=Location, 
             y=sq_percentcover, 
             color=Location)) + 
  geom_boxplot() + 
  theme_classic() 

PocCover_boxplot

##### patch together 
joined_pocandsilicate <- PocCover_boxplot + Settlement_boxplot

joined_pocandsilicate

```



