---
title: "Models and Figures for Chapter 2 Hypotheses"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Story: 
-  A bit of a chicken and an egg tale: 
There are three hypotheses that look at a full circle story from settlement to adulthood 
-  **Hypothesis 1a:** Does percent cover of adult corals drive the total number of settlers on a tile? (spoiler alert: yes)
-  **Hypothesis 1b:** Does silicate drive the total number of settlers on a tile? (spoiler alert: also yes, but slighly less than cover)
-  **Hypothesis 2a:** Once settled, what factors drive survivorship? Density dependence! (spoiler alert: duh!)
-  **Hypothesis 2b:** Beyond density dependence, does SGD impact the survivorship (spoiler alert: no)
-  **Hypothesis 3:** Of the settlers that survive, does the number impact the percent cover into adulthood? (spoiler alert: sort of, it's complicated) - NOT DOING THIS ANYMORE BECAUSE WAS NOT SIGNIFICANT 

# The Models: 
Although this story looks as if it should be a perfect fit for an SEM, there are a few roadblocks: 
-  percent cover and SGD are highly correlated/covaried. Need to figure out how to deal with that in an SEM because SEM does not accept separate models with the same response variable, so SGD and percent cover would need to be together, but then results are not accurate depiction of the situation. 

-  The zeros are removed from Hyp 3 because only a few indivdual recruits on the plates and they were skewing the data 

-  Considering Varari and Cabral in separate models; however, maybe combine all into one model and use Location as an interaction term

#### Load libraries 
NOTE: needed to download and install an old version of "heavy" package because it was removed from the CRAN repository in 2022 
```{r, echo=FALSE, message=FALSE}
library(ggeffects)
library(patchwork)
library(MASS)
library(car)
library(lavaan)
library(tidyverse)
library(lmerTest)
library(lme4)
library(nlme)
library(piecewiseSEM)
library(GGally)
library(emmeans)
library(mgcv)
library(semEff)
library(here)
library(PNWColors)
library(agricolae)
library(modelsummary)
library(performance)
library(modelr)
library(survival)
library(betareg)
library(statmod)
library(ggsurvfit)
library(survminer)
#library(heavy)
library(rstanarm)
library(devtools)
library(sjPlot)
library(effectsize)
library(marginaleffects)
library(MuMIn)
```


#### Read in Data and Tidy
-  Includes some background info like range of settler counts between sites and mean number of settlers 

```{r, echo=FALSE, message=FALSE}

SettlementDF_Full <- read_csv(here::here("Data", "EnvData_wPocCover.csv"))
## FULL dataset with pocillopora cover data, env data (nuts, etc), and recruitment counts 
## CommComp --> EnvDataPocCover 

CoralGenera_Tiles <- read_csv(here::here("Data", "Recruitment_Tiles_Counts.csv")) 

Poc_cover_data <- read_csv(here::here("Data", "Species_Composition_2022_Danielle.csv"))

full <- read_csv((here::here("Data", "full.csv")))


#### tidy the data 

SettlementDF_Full <- SettlementDF_Full %>% 
  dplyr::select(!c("temp_avg":"NN_avg", "ammonia_avg":"total_count_biotics")) %>% 
  dplyr::mutate(proportion_alive = sum_total_alive/sum_total, 
                proportion_alive_log = log(sum_total_alive+1)/log(sum_total+1)) %>% 
  filter(CowTagID!="C19") ## because no settlers at all so remove entirely from dataset 

##################################
## what is overall % survivorship? 
##################################
#mean(SettlementDF_FullC$proportion_alive)
#mean(SettlementDF_FullV$proportion_alive)


## is this significant? 
mean_survivorship <- SettlementDF_Full %>% 
  dplyr::select(c("Location","proportion_alive"))
  #summarize()

mean_survivorship_model <- lm(proportion_alive ~ Location, data = mean_survivorship)
anova(mean_survivorship_model)

##################################
## background stats 
##################################
## Cabral stats 
SettlementDF_FullC <- SettlementDF_Full %>% 
  filter(Location=="Cabral")

min(SettlementDF_FullC$sum_total)
max(SettlementDF_FullC$sum_total)
mean(SettlementDF_FullC$sum_total)
sd(SettlementDF_FullC$sum_total) / sqrt(length(SettlementDF_FullC$sum_total))


min(SettlementDF_FullC$proportion_alive)
max(SettlementDF_FullC$proportion_alive)
mean(SettlementDF_FullC$proportion_alive)
sd(SettlementDF_FullC$proportion_alive) / sqrt(length(SettlementDF_FullC$proportion_alive))


max(SettlementDF_FullC$silicate_avg)
mean(SettlementDF_FullC$silicate_avg)
sd(SettlementDF_FullC$silicate_avg) / sqrt(length(SettlementDF_FullC$silicate_avg))

min(SettlementDF_FullC$percent_cover)
max(SettlementDF_FullC$percent_cover)
mean(SettlementDF_FullC$percent_cover)
sd(SettlementDF_FullC$percent_cover) / sqrt(length(SettlementDF_FullC$percent_cover))




## Varari stats 
SettlementDF_FullV <- SettlementDF_Full %>% 
  filter(Location=="Varari")

min(SettlementDF_FullV$sum_total)
max(SettlementDF_FullV$sum_total)
mean(SettlementDF_FullV$sum_total)
sd(SettlementDF_FullV$sum_total) / sqrt(length(SettlementDF_FullV$sum_total))


min(SettlementDF_FullV$proportion_alive)
max(SettlementDF_FullV$proportion_alive)
mean(SettlementDF_FullV$proportion_alive)
sd(SettlementDF_FullV$proportion_alive) / sqrt(length(SettlementDF_FullV$proportion_alive))


max(SettlementDF_FullV$silicate_avg)
mean(SettlementDF_FullV$silicate_avg)
sd(SettlementDF_FullV$silicate_avg) / sqrt(length(SettlementDF_FullV$silicate_avg))

min(SettlementDF_FullV$percent_cover)
max(SettlementDF_FullV$percent_cover)
mean(SettlementDF_FullV$percent_cover)
sd(SettlementDF_FullV$percent_cover) / sqrt(length(SettlementDF_FullV$percent_cover))


```

#### Plot for the breakdown of coral genera on the tiles - old 
```{r}

## tidy data in way that makes sense for plot 

CoralGenera_Tiles_edit <- CoralGenera_Tiles %>% 
  dplyr::select(!"Notes") %>% 
  pivot_longer(cols="Pocilloporidae":"Other", names_to = "coral_genera" , values_to= "counts") %>% 
  group_by(CowTagID, Location, coral_genera) %>%  
  summarize(genera_total=sum(counts)) %>% 
  ungroup() %>% 
  group_by(CowTagID, Location) %>%  
  mutate(sum_total=sum(genera_total)) %>% 
  mutate(percentage = (genera_total/sum_total)*100) %>% 
  filter(CowTagID!="C19") %>% 
  ungroup() %>% 
  group_by(coral_genera, Location) %>%
  mutate(mean_percentage = mean(percentage),
         se = sd(percentage) / sqrt(n()))  # Standard error calculation

#######################################
## for stacked bar plot - no averages 
########################################

## calculate total at Varari and at Cabral 
      ## 433 at Cabral and 676 at Varari 

CoralGenera_Tiles_editV <- CoralGenera_Tiles %>% 
  dplyr::select(!"Notes") %>%
  filter(Location=="Varari") %>% 
  pivot_longer(cols="Pocilloporidae":"Other", names_to = "coral_genera" , values_to= "counts") %>% 
  group_by(Location, coral_genera) %>% 
  summarize(genera_total=sum(counts), 
            percent = (genera_total/676)*100)

CoralGenera_Tiles_editC <- CoralGenera_Tiles %>% 
  dplyr::select(!"Notes") %>%
  filter(Location=="Cabral") %>% 
  pivot_longer(cols="Pocilloporidae":"Other", names_to = "coral_genera" , values_to= "counts") %>% 
  group_by(Location, coral_genera) %>% 
  summarize(genera_total=sum(counts), 
            percent = (genera_total/433)*100)


CoralGenera_Tiles_edit2 <- CoralGenera_Tiles_editC %>% 
  bind_rows(CoralGenera_Tiles_editV)
  
########################################
## plot percent total of the genera
########################################

coral_generaplot1 <- CoralGenera_Tiles_edit2 %>% 
  mutate(grp = factor(coral_genera, levels = c("Other","Acroporidae","Poritidae","Pocilloporidae"))) %>% # put the levels in an order I want for plotting
  ggplot(aes(x=Location, 
             y=percent, 
             fill=grp)) + 
  geom_col() + 
  #geom_errorbar(aes(ymin = mean_percentage - se, 
                  #  ymax = mean_percentage + se), 
               # width = 0.2, position = position_dodge(0.7)) +  # Add error bars
  theme_classic() + 
  scale_fill_manual(values=c("gray", "coral2", "chocolate4", "burlywood"), 
                    limits=c("Other", "Acroporidae", "Poritidae", "Pocilloporidae")) +
  labs(x = "Location", 
       y= "Percentage of Total Settlers", 
       fill= "Coral Family") + 
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        legend.justification.inside = c(1, 1)) ### NOT WORKING RIGHT NOW 
  
coral_generaplot1
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "coral_genera_tiles.jpg"), 
       width=10, height=7)

```


#### Poc acuta cover - updated November 7th, 2024
-  dataframe "full" is coming from Poc_Cover_data script
```{r}



fullC <- full %>% 
  filter(Location=="Cabral")

min(fullC$pacuta_cover)
max(fullC$pacuta_cover)
mean(fullC$pacuta_cover)
sd(fullC$pacuta_cover) / sqrt(length(fullC$pacuta_cover))

fullV <- full %>% 
  filter(Location=="Varari")

##########################
## calculate % cover of corals
##########################
min(fullC$coral_percentcover)
max(fullC$coral_percentcover)
mean(fullC$coral_percentcover)
sd(fullC$coral_percentcover) / sqrt(length(fullC$coral_percentcover))

min(fullV$coral_percentcover)
max(fullV$coral_percentcover)
mean(fullV$coral_percentcover)
sd(fullV$coral_percentcover) / sqrt(length(fullV$coral_percentcover))

##########################
## calculate % cover of P acuta 
##########################
min(fullC$pacuta_cover)
max(fullC$pacuta_cover)
mean(fullC$pacuta_cover)
sd(fullC$pacuta_cover) / sqrt(length(fullC$pacuta_cover))

min(fullV$pacuta_cover)
max(fullV$pacuta_cover)
mean(fullV$pacuta_cover)
sd(fullV$pacuta_cover) / sqrt(length(fullV$pacuta_cover))


##########################
## calculate actual number of settlers, model, and se 
##########################

min(fullV$sum_total)
max(fullV$sum_total)
mean(fullV$sum_total)
sd(fullV$sum_total) / sqrt(length(fullV$sum_total))

##########################
## calculate actual number of survivors, model, and se 
##########################

min(fullV$sum_total_alive)
max(fullV$sum_total_alive)
mean(fullV$sum_total_alive)
sd(fullV$sum_total_alive) / sqrt(length(fullV$sum_total_alive))

min(fullC$sum_total_alive)
max(fullC$sum_total_alive)
mean(fullC$sum_total_alive)
sd(fullC$sum_total_alive) / sqrt(length(fullC$sum_total_alive))



```



## Hypothesis 1 (a and b):
-  Using a **generalized linear model**, negative binomial 
-  **Results:** There is a site difference for the effect of silicate, which makes sense because the two sites are so different in their dynamics (silicate concentration, flow, etc). However, there is no effect of site for percent cover. Percent cover significantly affects the number of total recruits that settle on the tiles; however silicate only affects the total number of settlers at Varari (linearly and non-linearly), but not at Cabral. According to the **AIC**, the silicate model is a lower rank, so percent cover is a better/stronger driver of total settlement, but silicate is still important. Looking at the **standardized effect sizes**, the power of the relationship is strong with .... 

-  Why separate it into 2 models? Because too many variables at play, and silicate & percent cover are too highly correlated 
-  Explanatory values are independent of the family... but they sometimes have to be logged or transformed if you have issues with your residuals being normal in your check model



### models 
-  logged percent cover 
```{r}
#######################
## decision is to keep both sites together and use "Location" as an interaction term in the model 
#######################

############
#### Hypothesis 1a - percent cover 
############
PocCover_negbinom_old <- glm.nb(sum_total ~ log(percent_cover+1)*Location, data=SettlementDF_Full)

PocCover_negbinom_new <- glm.nb(sum_total ~ log(pacuta_cover+1)*Location, data=full)
anova(PocCover_negbinom_old)
summary(PocCover_negbinom)  
## AIC = 341.62
r.squaredGLMM(PocCover_negbinom)
check_model(PocCover_negbinom)

############
#### Hypothesis 1b - SGD 
############
SGD_negbinom <- glm.nb(sum_total+1 ~ ((silicate_avg) + I(silicate_avg^2))*Location, data=SettlementDF_Full)

anova(SGD_negbinom) ## SIGNIFICANT FOR VARARI 
summary(SGD_negbinom)
## AIC = 355.44
check_model(SGD_negbinom)


############
#### September 13th, 2025: Check for collinearity between silicate and percent cover 
############
both_factors_collin <- glm.nb(sum_total+1 ~ ((silicate_avg) + I(silicate_avg^2)) + log(percent_cover+1)*Location, data=SettlementDF_Full)

cor(SettlementDF_Full$silicate_avg, SettlementDF_Full$percent_cover)


anova(both_factors_collin) 
summary(both_factors_collin)
check_model(both_factors_collin)

#### run models separately - remove site as interaction term 

#### Varari models 

#percent cover 
PocCover_V <- glm.nb(sum_total ~ log(percent_cover+1), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Varari"))


anova(PocCover_V) ## p = 0.037
summary(PocCover_V)
check_model(PocCover_V)

#SGD 
SGD_V <- glm.nb(sum_total+1 ~ ((silicate_avg) + I(silicate_avg^2)), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Varari"))


anova(SGD_V) ## p = 0.006 for linear, 0.005 for polynomial- lots of collinearity 
summary(SGD_V)
check_model(SGD_V)

#### Cabral models 

#percent cover 
PocCover_C <- glm.nb(sum_total ~ log(percent_cover+1), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Cabral"))


anova(PocCover_C) ## p = 0.02
summary(PocCover_C)
check_model(PocCover_C)

#SGD 
SGD_C <- glm.nb(sum_total+1 ~ ((silicate_avg) + I(silicate_avg^2)), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Cabral"))


anova(SGD_C) ## p = 0.38 for linear, 0.8 for polynomial 
summary(SGD_C)
check_model(SGD_C) # high collinearity 
```

### neg binomial plots 

```{r}

#######################
## plot the negative binomial 
#######################

#########################
#### Percent Cover 
#########################

PocCover_negbinom <- glm.nb(sum_total ~ percent_cover*Location, data=SettlementDF_Full)
PocCover_negbinom ## model 

anova(PocCover_negbinom)
summary(PocCover_negbinom)

######## plot 4  --> doesn't have predicted data in there though 

newdata_cover = predict_response(PocCover_negbinom, terms= "percent_cover")

newdata_cover2 <- SettlementDF_Full %>%
  mutate(fitted_values = predict(PocCover_negbinom, terms= "percent_cover")) 


test4_percentcover_predplot <- ggplot() + 
  geom_point(data = full,
             aes(x=pacuta_cover, 
                 y=sum_total, 
                 color=Location), 
             size=4) +
 scale_color_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari")) +
 # geom_line(data = newdata_cover, 
         #   aes(x=x, 
             #   y=predicted)) #+ 
  geom_smooth(data = full, 
             aes(x=pacuta_cover, 
                  y=sum_total), 
             method = "glm.nb", formula = "y~x", color="black", alpha=0.5) + 
 coord_cartesian(ylim = c(0, 250)) + 
  theme_classic() + 
  labs(x="Percent Cover of Adults", 
       y="Total Settlers Present on Tiles") + 
      theme(axis.text.x=element_text(size=40), 
        axis.text.y=element_text(size=40), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=30), 
        legend.text=element_text(size=40),  # Adjust legend text size
        legend.title=element_text(size=40), # Adjust legend title size
        strip.text=element_text(size=40)) + 
  guides(color="none")


test4_percentcover_predplot
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "Percent_Cover_glmnbPlot.jpg"), 
       width=10, height=7)

######################
#### SGD 
######################

SGD_negbinom ## model 
SGD_negbinom <- glm.nb(sum_total ~ (silicate_avg + I(silicate_avg^2))*Location, data=SettlementDF_Full)

anova(SGD_negbinom)
summary(SGD_negbinom)


SGD_V_glmnbPlot <- ggplot() + 
  geom_point(data = SettlementDF_Full %>% filter(Location=="Varari"),
             aes(x=(silicate_avg^2), 
                 y=sum_total), 
             size=4, 
             color="firebrick4") +
 # geom_line(data = newdata_cover, 
         #   aes(x=x, 
             #   y=predicted)) #+ 
  geom_smooth(data = SettlementDF_Full %>% filter(Location=="Varari"), 
             aes(x=I(silicate_avg^2), 
                  y=sum_total), 
             method = "glm.nb", formula = "y~poly(x,2)", color="black", alpha=0.5) + 
  theme_classic() + 
  labs(x = expression(Average~Silicate~(mu*mol~L^-1)), 
       y="Total Settlers Present on Tiles") + 
    theme(axis.text.x=element_text(size=40), 
        axis.text.y=element_text(size=40), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=30), 
        legend.text=element_text(size=40),  # Adjust legend text size
        legend.title=element_text(size=40), # Adjust legend title size
        strip.text=element_text(size=40))

SGD_V_glmnbPlot
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "SGD_V_glmnbPlot.jpg"), 
       width=11, height=9)



SGD_C_glmnbPlot <- ggplot() + 
  geom_point(data = SettlementDF_Full %>% filter(Location=="Cabral"),
             aes(x=(silicate_avg^2), 
                 y=sum_total), 
             size=4, 
             color="goldenrod") +
 # geom_line(data = newdata_cover, 
         #   aes(x=x, 
             #   y=predicted)) #+ 
 # geom_smooth(data = SettlementDF_Full %>% filter(Location=="Cabral"), 
        #     aes(x=I(silicate_avg^2), 
           #       y=sum_total), 
          #   method = "glm.nb", formula = "y~poly(x,2)", color="black", alpha=0.5) + 
  theme_classic() + 
  labs(x = expression(Average~Silicate~(mu*mol~L^-1)), 
       y="Total Settlers Present on Tiles") + 
    theme(axis.text.x=element_text(size=40), 
        axis.text.y=element_text(size=40), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=30), 
        legend.text=element_text(size=40),  # Adjust legend text size
        legend.title=element_text(size=40), # Adjust legend title size
        strip.text=element_text(size=40))

SGD_C_glmnbPlot
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "SGD_C_glmnbPlot.jpg"), 
       width=11, height=9)


```

### standardized effects plots 

```{r}
#######################
## standardized effect sizes  
#######################
# Note: won't work on dependent var bc changes count data to continuous which then doesn't use glm.nb()
## only calculate sef for independent vars of SGD, SGD^2, and percent cover
## and take sef before the model, create new df 

############
## get z-scores of just independent vars before going into the model 
## You can get the zscore by using scale(x, center=TRUE, scale=TRUE)
## but need to make new df that has "parameter" which is the silicate or cover  and "zscores" which is the standardized score 


#percentcover_sef <- tibble("percentcover_zscore" = scale(SettlementDF_Full_C$percent_cover, center = TRUE, scale = TRUE))


#SettlementDF_Full_C$z_percent_cover <- scale(SettlementDF_Full_C$percent_cover, center = TRUE, scale = TRUE)
#SettlementDF_Full_C$z_SGD <- scale(SettlementDF_Full_C$silicate_avg, center = TRUE, scale = TRUE)
#SettlementDF_Full_C$z_SGD2 <- scale(SettlementDF_Full_C$silicate_avg^2, center = TRUE, scale = TRUE)

#SettlementDF_Full_V$z_percent_cover <- scale(SettlementDF_Full_V$percent_cover, center = TRUE, scale = TRUE)
#SettlementDF_Full_V$z_SGD <- scale(SettlementDF_Full_V$silicate_avg, center = TRUE, scale = TRUE)
#SettlementDF_Full_V$z_SGD2 <- scale(SettlementDF_Full_V$silicate_avg^2, center = TRUE, scale = TRUE)


```


## Hypothesis 2a:
-  Using a **general linear model** (gaussian lm)
-  **Results:** Significant with or without the 0s dropped for the 4 tiles that had few recruits and 100% mortality. For time being, keep the 0s in the model if significant. It is clear the density dependence is happening and driving the survivorship of the corals that settle on the tiles. Location is once again insignificant, as expected 

```{r}

### model the proportion of settlers that are alive out of the total on the plates 

proportion_alive <- lm(proportion_alive ~ log(sum_total+1)*Location, data=SettlementDF_Full)

anova(proportion_alive) ## barely significant, but still significant, Location is not sig (as expected)
summary(proportion_alive) ## report R squared too in results 
check_model(proportion_alive)

### remove the 0s from both Varari and Cabral because skewing the data - 0s meaning the plates that had settlers, but none survived -- there are only a handful of recruits on those plates, so probably weren't even supposed to be there 
### for all of the plates with 0, 6 or less settlers 

proportion_alive_noZero <- lm(proportion_alive ~ log(sum_total+1)*Location, data=SettlementDF_Full %>%
                                filter(proportion_alive>0))

anova(proportion_alive_noZero) ## super significant, Location is not sig (as expected)
summary(proportion_alive_noZero)
check_model(proportion_alive_noZero) ## high collinearity ** 

######################
## plot
######################

### with 0s 
### don't worry about faceting by location because we are keeping both sites together in the model 
densitydep_plot <- SettlementDF_Full %>% 
  ggplot(aes(x=sum_total, 
             y=proportion_alive)) +
  geom_point(size=4, aes(color=Location)) + 
  scale_color_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari")) +
  geom_smooth(method="lm", color="black") + 
  #coord_trans(x="log") +
  theme_classic() + 
  labs(x= "Logged Total Settlers", 
       y= "Proportion of Settlers Alive") + 
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

densitydep_plot
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "DensityPlot.jpg"), 
       width=12, height=9)


### without 0s 

densitydep_plot_noZeros <- SettlementDF_Full %>% 
  filter(proportion_alive>0) %>% 
   ggplot(aes(x=log(sum_total+1), 
             y=proportion_alive)) +
  geom_point(size=4, aes(color=Location)) + 
  scale_color_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari")) +
  geom_smooth(method="lm", color="black") + 
  theme_classic() + 
  #coord_trans(x="log") +
  labs(x= "Logged Total Settlers", 
       y= "Proportion of Settlers Alive") + 
  theme(axis.text.x=element_text(size=30), 
        axis.text.y=element_text(size=30), 
        axis.title.x=element_text(size=30), 
        axis.title.y=element_text(size=30), 
        legend.text=element_text(size=30),  # Adjust legend text size
        legend.title=element_text(size=30), # Adjust legend title size
        strip.text=element_text(size=30))

densitydep_plot_noZeros
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "DensityPlot_NoZero2.jpg"), 
       width=12, height=9)

```


## Hypothesis 2b:
-  Using a **general linear model** to plot the residuals of density dependence (Hyp 2a) against silicate to see if survivorship beyond density dependence is being driven by silicate 
-  **Results:**  

```{r}
################################
## calculate residuals: plotting the residuals from previous model from hyp 2a against silicate 
################################

#### with zeroes 
resid(proportion_alive)

resid_proportionalive <- SettlementDF_Full %>% 
  bind_cols(residuals = resid(proportion_alive)) 

#### without zeroes
resid(proportion_alive_noZero)

resid_proportionalive_noZero <- SettlementDF_Full %>% 
  filter(proportion_alive>0) %>% 
  bind_cols(residuals = resid(proportion_alive_noZero)) 

#####################
## models 
#####################

#### with zeroes 

resids_silicateEffect_model <- lm(residuals ~ (silicate_avg + I(silicate_avg^2))*Location , data=resid_proportionalive)

anova(resids_silicateEffect_model) 
summary(resids_silicateEffect_model)
r.squaredGLMM(resids_silicateEffect_model)
check_model(resids_silicateEffect_model)

#### without zeroes 

resids_silicateEffect_model_noZero <- lm(residuals ~ (silicate_avg + I(silicate_avg^2))*Location, data=resid_proportionalive_noZero)

anova(resids_silicateEffect_model_noZero)
summary(resids_silicateEffect_model_noZero)
check_model(resids_silicateEffect_model_noZero)


#####################
## plot 
#####################
#### plot the residuals (this basically says that whatever points are above the best fit line are surviving more than expected due to density dependence because of SGD, anything below the line is surviving less than expected due to SGD)

#### with zeroes 
survivorship_silicateEffectC <- resid_proportionalive %>% 
  filter(Location=="Varari") %>% 
  ggplot(aes(x=silicate_avg, 
             y=residuals)) +
  geom_point(size=4, color= "darkgoldenrod3") + 
#  scale_color_manual(values="firebrick4") +
  geom_smooth(method="lm", formula = "y~poly(x,2)", color="black") + 
  theme_classic() + 
  labs(x = expression(Silicate~(mu*mol~L^-1)), 
       y= "Residuals") + 
  theme(axis.text.x=element_text(size=40), 
        axis.text.y=element_text(size=40), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=40), 
        legend.text=element_text(size=40),  # Adjust legend text size
        legend.title=element_text(size=40), # Adjust legend title size
        strip.text=element_text(size=40))
  
survivorship_silicateEffectC
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "Survivorship_SilicateEffectC.jpg"), 
       width=12, height=10)


#### without zeroes 
survivorship_silicateEffect_noZero <- resid_proportionalive_noZero %>% 
  ggplot(aes(x=silicate_avg, 
             y=residuals)) +
  geom_point(size=4, aes(color=Location)) + 
  scale_color_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari")) +
 # geom_smooth(method="lm", color="black") + 
  theme_classic() + 
  labs(x = expression(Silicate~(mu*mol~L^-1)), 
       y= "Residuals of Alive Settlers") + 
  geom_smooth(method=lm, formula="y~poly(x,2)", color="black") + 
  coord_trans(x="log") +
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))
  
survivorship_silicateEffect_noZero
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "Survivorship_SilicateEffect_NoZero2.jpg"), 
       width=10, height=9)

```








# NOT USING ANYMORE 
## Hypothesis 3:
-  Using a **general linear model** (double check correctness)
-  **Results:**
-  Notes: when remove 0s from percent cover and total alive (though there are quite a few between both sites), you see a pretty strong relationship. The limitation with this hypothesis is that the data previously collected was percent cover data, and we do not have a true density count, so the 0s might not be a true reflection of the actual number of adult corals near the tile. The proportion of alive settlers is also a bit stronger here than just the total number of settlers. 

```{r}

##################
## models 
##################

## total alive 
alive_predict_cover <- lm(log(percent_cover+1) ~ log(sum_total_alive+1)*Location, data = SettlementDF_Full)

anova(alive_predict_cover) ## insignificant, though logging makes it WAY better and almost significant (0.06)
summary(alive_predict_cover)

## proportion alive 

propalive_predict_cover <- lm(percent_cover ~ proportion_alive*Location, data = SettlementDF_Full)

anova(propalive_predict_cover) ## significant 
summary(propalive_predict_cover)
check_model(propalive_predict_cover)

##################
## models without ZERO
##################

## total alive 
alive_predict_cover_noZero <- lm(percent_cover ~ log(sum_total_alive+1), data = SettlementDF_FullC %>% 
                                   filter(percent_cover>0))

anova(alive_predict_cover_noZero) ## significant ONLY FOR LOCATION, without logging 
summary(alive_predict_cover_noZero)
check_model(alive_predict_cover_noZero)

## proportion alive 

propalive_predict_cover_noZero <- lm(percent_cover ~ proportion_alive*Location, data = SettlementDF_Full %>% 
                                       filter(percent_cover>0))

anova(propalive_predict_cover_noZero) ## very significant 
summary(propalive_predict_cover_noZero)
check_model(propalive_predict_cover_noZero)



##################
## plots  
##################

## total alive 

alive_affect_cover <- SettlementDF_Full %>% 
  filter(percent_cover>0) %>% ## decide whether to keep in or not **
  ggplot(aes(x=sum_total_alive, 
             y=percent_cover)) +
  geom_point(size=4, aes(color=Location)) + 
  scale_color_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari")) +
  geom_smooth(method="lm", color="black") + 
  theme_classic() + 
  labs(x = "Total Alive Settlers", 
       y = expression("Percent Cover of " * italic("P. acuta") * " adults ")) +
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

alive_affect_cover

ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "Cover_AliveEffect.jpg"), 
       width=12, height=9)


## proportion alive 

propalive_affect_cover <- SettlementDF_Full %>% 
  filter(percent_cover>0) %>% 
  ggplot(aes(x=proportion_alive, 
             y=percent_cover)) +
  geom_point(size=4, aes(color=Location)) + 
  scale_color_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari")) +
  geom_smooth(method="lm", color="black") + 
  theme_classic() + 
  labs(x = "Proportion of Alive Settlers", 
       y = expression("Percent Cover of " * italic("P. acuta") * " adults ")) +
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

propalive_affect_cover
ggsave(here::here("Outputs", "RecruitmentTiles", "Thesis_Figures", "Cover_PropAliveEffect.jpg"), 
       width=12, height=9)

```


## Percent Cover and SGD 

```{r}


percentcover_SGD <- lm(log(percent_cover+1) ~ silicate_avg*Location, data = SettlementDF_Full)

anova(percentcover_SGD)
check_model(percentcover_SGD)

```




#### testing salinity relationships 

```{r}

##################################
## salinity stats - see a pattern? 
##################################
## separate df for salinity stats 
SettlementDF_FullC 
SettlementDF_FullV


## Cabral stats

min(SettlementDF_FullC$sal_avg)
max(SettlementDF_FullC$sal_avg)
mean(SettlementDF_FullC$sal_avg)
sd(SettlementDF_FullC$sal_avg) / sqrt(length(SettlementDF_FullC$sal_avg))


## Varari stats 

min(SettlementDF_FullV$sal_avg)
max(SettlementDF_FullV$sal_avg)
mean(SettlementDF_FullV$sal_avg)
sd(SettlementDF_FullV$sal_avg) / sqrt(length(SettlementDF_FullV$sal_avg))

##################################
## models to see if there is any pattern, probably not since salinity is so small of a range  
##################################

## both sites together 
salavg_negbinom <- glm.nb(sum_total ~ sal_avg*Location, data=SettlementDF_Full)

anova(salavg_negbinom)
summary(salavg_negbinom) ## RESULT: definitely not significant 
check_model(salavg_negbinom) ## 

```



#### Total spp coverage 
```{r}
Poc_cover_data_edit2 <- Poc_cover_data %>% 
  dplyr::select(!c("Date", "PhotoNum", "Notes")) %>% 
  dplyr::filter(Taxa!="Bare Rock",Taxa!="Sand",Taxa!="Rubble", Location!="Varari_Maya") %>% 
  group_by(CowTagID) 

Poc_cover_data_edit3 <- Poc_cover_data_edit2 %>% 
  dplyr::summarise(total_count_biotics = sum(SpeciesCounts)) %>% 
  filter(CowTagID!="VSEEP") ## missing C14 

Poc_cover_data_edit4 <- Poc_cover_data_edit2 %>% 
  filter(Taxa=="Pocillopora acuta"|Taxa=="Porites rus"|Taxa=="Montipora verrilli"|Taxa=="Montipora setosa"|Taxa=="Montipora grisea"|Taxa=="Porites irregularis"|Taxa=="Porites lobata"|Taxa=="Pocillopora verrucosa"|Taxa=="Pavona chiriquiensis"|Taxa=="	
Pavona cactus"|Taxa=="Leptastrea transversa"|Taxa=="Acropora"|Taxa=="Acropora pulchra"|Taxa=="Montipora capitata") %>% 
  group_by(CowTagID) %>% 
  mutate(SpeciesCounts_Coral = SpeciesCounts) %>% 
  dplyr::select(!SpeciesCounts)
 
### joins correlated biotics counts for that CowTag -- still with missing cowtags without P acuta data 
Poc_cover_data_edit5 <- Poc_cover_data_edit4 %>% 
  left_join(Poc_cover_data_edit3) %>% 
  group_by(CowTagID, Location, SpeciesCounts_Coral,total_count_biotics) %>% 
  dplyr::mutate(percent_cover = (SpeciesCounts_Coral/total_count_biotics)*(100)) ### should be divided by 200 (all the points in the square) -- for all corals and then for just P acuta 

##############################
## total cover % and then what is Pocillopora of that total 
##############################

Poc_cover_data_edit6 <- Poc_cover_data_edit5 %>%
  group_by(CowTagID, Location) %>% 
  dplyr::mutate(total_coral_cover_forCowTag = sum(percent_cover))

Poc_cover_data_edit6_C <- Poc_cover_data_edit5 %>%
  filter(Location=="Cabral") %>% 
  group_by(CowTagID) %>% 
  dplyr::summarize(total_coral_cover_forCowTag = sum(percent_cover)) 

Poc_cover_data_edit6_V <- Poc_cover_data_edit5 %>%
  filter(Location=="Varari") %>% 
  group_by(CowTagID) %>% 
  dplyr::summarize(total_percent_cover = sum(percent_cover)) 


min(Poc_cover_data_edit6_C$total_percent_cover)
max(Poc_cover_data_edit6_C$total_percent_cover)
mean(Poc_cover_data_edit6_C$total_percent_cover)
sd(Poc_cover_data_edit6_C$total_percent_cover) / sqrt(length(Poc_cover_data_edit6$total_percent_cover))

min(Poc_cover_data_edit6_V$total_percent_cover)
max(Poc_cover_data_edit6_V$total_percent_cover)
mean(Poc_cover_data_edit6_V$total_percent_cover)
sd(Poc_cover_data_edit6_V$total_percent_cover) / sqrt(length(Poc_cover_data_edit6$total_percent_cover))

#######
#######

Poc_cover_data_edit7 <- Poc_cover_data_edit6 %>%
  filter(Taxa=='Pocillopora acuta') %>% 
  dplyr::mutate(Pacuta_percent_cover = (percent_cover/total_coral_cover_forCowTag)*100) #### percent cover 

Poc_cover_data_edit7_C <- Poc_cover_data_edit7 %>%
  filter(Location=="Cabral")

Poc_cover_data_edit7_V <- Poc_cover_data_edit7 %>%
  filter(Location=="Varari") 


min(Poc_cover_data_edit7_C$Pacuta_percent_cover)
max(Poc_cover_data_edit7_C$Pacuta_percent_cover)
mean(Poc_cover_data_edit7_C$Pacuta_percent_cover)
sd(Poc_cover_data_edit7_C$Pacuta_percent_cover) / sqrt(length(Poc_cover_data_edit7_C$Pacuta_percent_cover))

min(Poc_cover_data_edit7_V$Pacuta_percent_cover)
max(Poc_cover_data_edit7_V$Pacuta_percent_cover)
mean(Poc_cover_data_edit7_V$Pacuta_percent_cover)
sd(Poc_cover_data_edit7_V$Pacuta_percent_cover) / sqrt(length(Poc_cover_data_edit7_V$Pacuta_percent_cover))


###############################
## calculate percent cover by species between site
###############################

## Pocillopora  

Pocillopora <- Poc_cover_data_edit5 %>% 
  filter(Taxa=="Pocillopora acuta"|Taxa=="Pocillopora verrucosa") %>% 
  group_by(Location) %>% 
  dplyr::summarize(totalspp = sum(SpeciesCounts_Coral), 
                   totalpercentcover = sum(percent_cover))


## Montiporas 
Montiporas <- Poc_cover_data_edit5 %>% 
  filter(Taxa=="Montipora verrilli"|Taxa=="Montipora setosa"|Taxa=="Montipora grisea"|Taxa=="Montipora capitata") %>% 
  group_by(Location) %>% 
  dplyr::summarize(totalspp = sum(SpeciesCounts_Coral), 
                   totalpercentcover = sum(percent_cover))

## Porites 
Porites <- Poc_cover_data_edit5 %>% 
  filter(Taxa=="Porites lobata"|Taxa=="Porites rus"|Taxa=="Porites irregularis") %>% 
  group_by(Location) %>% 
  dplyr::summarize(totalspp = sum(SpeciesCounts_Coral), 
                   totalpercentcover = sum(percent_cover))


## Acropora 
Acropora <- Poc_cover_data_edit5 %>% 
  filter(Taxa=="Acropora"|Taxa=="Acropora pulchra") %>% 
  group_by(Location) %>% 
  dplyr::summarize(totalspp = sum(SpeciesCounts_Coral), 
                   totalpercentcover = sum(percent_cover))

## Leptastrea 
Leptastrea <- Poc_cover_data_edit5 %>% 
  filter(Taxa=="Leptastrea transversa") %>% 
  group_by(Location) %>% 
  dplyr::summarize(totalspp = sum(SpeciesCounts_Coral), 
                   totalpercentcover = sum(percent_cover))

## Pavona 
Pavona <- Poc_cover_data_edit5 %>% 
  filter(Taxa=="Pavona chiriquiensis"|Taxa=="Pavona cactus") %>% 
  group_by(Location) %>% 
  dplyr::summarize(totalspp = sum(SpeciesCounts_Coral), 
                   totalpercentcover = sum(percent_cover))


##############################################
## Calculate algal turf cover across SGD gradient 
##############################################

turf_cover <- Poc_cover_data_edit2 %>% 
  filter(Taxa=="Turf")

turf_cover_joined <- turf_cover %>% 
  full_join(Poc_cover_data_edit3) %>% 
  group_by(CowTagID, Taxa, Location) %>% 
  dplyr::summarize(total_algae = sum(SpeciesCounts), 
                   percent_cover_algae = (total_algae/total_count_biotics)*(100)) %>% 
  full_join(SettlementDF_Full) %>% 
  filter(CowTagID!= "C19",CowTagID!= "C14") %>% 
  dplyr::select(!c("sum_total":"sal_avg","percent_cover":"proportion_alive_log")) 
 

turf_silicate <- lm(percent_cover_algae ~ silicate_avg*Location, data = turf_cover_joined)
anova(turf_silicate)
    

CCA_cover <- Poc_cover_data_edit2 %>% 
  filter(Taxa=="Crustose Corallines")
         

CCA_cover_joined <- CCA_cover %>% 
  full_join(Poc_cover_data_edit3) %>% 
  group_by(CowTagID, Taxa, Location) %>% 
  dplyr::summarize(total_CCA = sum(SpeciesCounts), 
                   percent_cover_CCA = (total_CCA/total_count_biotics)*(100)) %>% 
  full_join(SettlementDF_Full) %>% 
  filter(CowTagID!= "C19",CowTagID!= "C14") %>% 
  dplyr::select(!c("sum_total":"sal_avg","percent_cover":"proportion_alive_log")) 

CCA_silicate <- lm(percent_cover_CCA ~ silicate_avg*Location, data = CCA_cover_joined)
anova(CCA_silicate)


## quick plot to see direction of relationship 

CCAplot <- CCA_cover_joined %>% 
  ggplot(aes(x=silicate_avg, 
             y=percent_cover_CCA)) + 
  geom_point() + 
  facet_wrap(~Location, scales="free_x")
  
CCAplot

```



## Incorporating an SEM? 





##### EXTRA 
- other old version of models - moving here to clean up space up top 
```{r}
PocCover_negbinom2 <- glm.nb((sum_total+1) ~ log(percent_cover+1)*Location, data=SettlementDF_Full)

anova(PocCover_negbinom)
summary(PocCover_negbinom2)
check_model(PocCover_negbinom2)
## RESULT: log percent cover is significant, no location interaction or just location effect 
#r.squaredGLMM(PocCover_negbinom)
## AIC = 345.31, p-value = 0.01 


## logged? 
SGD_negbinom2 <- glm.nb(sum_total+1 ~ (log(silicate_avg) + I(log(silicate_avg^2)))*Location, data=SettlementDF_Full)

anova(SGD_negbinom2)
summary(SGD_negbinom2) 
check_model(SGD_negbinom2) ## obviously very colinear bc linear and nonlinear 


## try logged silicate 
resids_silicateEffect_model_noZero_log <- lm(residuals ~ (log(silicate_avg+1) + I(log(silicate_avg+1)^2))*Location, data=resid_proportionalive_noZero)

anova(resids_silicateEffect_model_noZero_log)
summary(resids_silicateEffect_model_noZero_log)
check_model(resids_silicateEffect_model_noZero_log)

## try logged 
resids_silicateEffect_model_log <- lm(residuals ~ (log(silicate_avg+1) + I(log(silicate_avg+1)^2))*Location , data=resid_proportionalive)

anova(resids_silicateEffect_model_log) 
summary(resids_silicateEffect_model)
check_model(resids_silicateEffect_model)

```

## Hypothesis 2a:
-  Using a **generalized linear model** (family, student's t distribution OR quasi binomial?) -- student's t does not weight as heavily on outliers. Use instead of a normal distribution in gaussian model. Also works for smaller sample sizes.
-  **Results:** When trying quasibinomial, results are significant with log link (though, double check if that is correct because already logged sum_total in the model). 

```{r}

###########################
## student ts
###########################

############
## trying from stack overflow, old heavy package 
#install_url('http://cran.r-project.org/src/contrib/Archive/heavy/heavy_0.38.19.tar.gz') ## not working right now  

#fit <- heavyLm(m.marietta ~ CRSP, data = SettlementDF_Full, family = Student(df = 6))
#summary(fit)


#############
## trying with gaussian and identity link 
#studentt_version1 <- stan_glm(proportion_alive ~ log(sum_total+1)*Location, data = SettlementDF_Full, family = gaussian(link = "identity"))

#summary(studentt_version1) ## not sure how to interpret 



###########################
## quasibinomial 
###########################
#RESULT: neither of these with "logit" link are significant  
#BUT significant with "log" family. Without zero model did not converge. 

#### with zeroes 
quasibin_model <- glm(proportion_alive ~ log(sum_total+1)*Location, family = quasibinomial(link="log"), data=SettlementDF_Full)

summary(quasibin_model)
check_model(quasibin_model)

#### without zeroes 

quasibin_model_noZero <-  glm(proportion_alive ~ log(sum_total+1)*Location, family = quasibinomial(link="log"), data=SettlementDF_Full %>%
                                filter(proportion_alive>0))

summary(quasibin_model)
check_model(quasibin_model_noZero) ## some high collinearity 


```




