---
title: "SEM_PocCover_Recruits"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries 
```{r, echo = FALSE, message=FALSE}
library(ggeffects)
library(patchwork)
library(MASS)
library(car)
library(lavaan)
library(tidyverse)
library(lmerTest)
library(lme4)
library(nlme)
library(GGally)
```

## Read in Data 
```{r, echo = FALSE, message=FALSE}
EnvData_wPocCover <- read_csv(here::here("Data", "EnvData_wPocCover.csv"))
## FULL dataset with pocillopora cover data, env data (nuts, etc), and recruitment counts 

```


## What is a structural equation model? 
-  It integrates a number of different multivariate techniques into one model fitting framework. 
-  Numerous independent and dependent variables all affecting each other. More of a _system_ than regression organization. 
-  Indirect effects and direct effects 
-  There are certain shapes and arrows that describe the variables and their relationships to one another 
-  UCLA [page](https://stats.oarc.ucla.edu/r/seminars/rsem/)

-  **Steps:** 1) specify the model. You can use the lavaan package to write the model syntax. You can have multiple explanatory variables (denoted with a "=~") and one response variable (denoted with just a '~'). 2) Then determine the model fit. The output gives you the summary of the estimated model. 3) visualize model and causal relationships between variables using pathSEM package. 


## Models based on hypotheses
```{r}

poc_cover <- lm(percent_cover ~ silicate_avg + sum_total, data=EnvData_wPocCover)
recruits <- lm(sum_total ~ silicate_avg + log(percent_cover+1), data=EnvData_wPocCover)


qqp(resid(poc_cover),"norm") 
qqp(resid(recruits),"norm") ### may need to transform in a different way 

```


## Structural equation model for recruitment tiles and Poc cover 
- Separate Varari and Cabral  

```{r}
##############################
## Varari 
##############################

## create the sem model, needed to remove reciprocal relationsship/path of sumtotal ~ percentcover because it was breaking the information matrix in lavaaan 

sem_modelV <- '
  # Silicate_avg impacts recruits and percent_cover
  logsum_total ~ silicate_avg
  logpercent_cover ~ silicate_avg
  
  # Recruits impacts percent_cover
  logpercent_cover ~ logsum_total
'

# create the df and tidy as necessary 
EnvData_wPocCoverV <- EnvData_wPocCover %>%
  filter(Location=="Varari") %>% 
  mutate(silicate_avg = as.numeric(silicate_avg),
         sum_total = as.numeric(sum_total),
         percent_cover = as.numeric(percent_cover), 
         logsilicate_avg = log(silicate_avg+1),
         logsum_total = log(sum_total+1),
         logpercent_cover = log(percent_cover+1)) %>%
  drop_na(silicate_avg, sum_total, percent_cover)

## fit and summarize the model 
fitV <- sem(sem_modelV, data = EnvData_wPocCoverV)
summary(fitV, fit.measures = TRUE, standardized = TRUE)


##############################
## Cabral 
##############################

sem_modelC <- '
  # Silicate_avg impacts recruits and percent_cover
  logsum_total ~ silicate_avg
  logpercent_cover ~ silicate_avg
  
  # Recruits impacts percent_cover
  logpercent_cover ~ logsum_total
'

EnvData_wPocCoverC <- EnvData_wPocCover %>%
  filter(Location=="Cabral") %>% 
  mutate(silicate_avg = as.numeric(silicate_avg),
         sum_total = as.numeric(sum_total),
         percent_cover = as.numeric(percent_cover), 
         logsilicate_avg = log(silicate_avg+1),
         logsum_total = log(sum_total+1),
         logpercent_cover = log(percent_cover+1)) %>%
  drop_na(silicate_avg, sum_total, percent_cover)


fitC <- sem(sem_modelC, data = EnvData_wPocCoverC)
summary(fitC, fit.measures = TRUE, standardized = TRUE)




```


## Visualize results and assess the model fit 

```{r}

#############################
## Varari 
#############################

ggplot(EnvData_wPocCoverV, 
       aes(x = silicate_avg, 
           y = log(percent_cover+1))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Relationship between SGD and Percent Cover")


ggplot(EnvData_wPocCoverV, 
       aes(x = log(percent_cover+1) , 
           y = log(sum_total+1))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Relationship between POC Cover and Recruits")


fitMeasures(fitV, c("rmsea", "cfi", "srmr"))


#############################
## Cabral 
#############################

ggplot(EnvData_wPocCoverC, 
       aes(x = silicate_avg, 
           y = log(percent_cover+1))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Relationship between SGD and Percent Cover")


ggplot(EnvData_wPocCoverC, 
       aes(x = log(percent_cover+1) , 
           y = log(sum_total+1))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Relationship between POC Cover and Recruits")


fitMeasures(fitC, c("rmsea", "cfi", "srmr"))



```




