---
title: "SGD Grouping with Recruitment Tiles"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

## Load libraries 
```{r}
library(patchwork)
library(tidyverse)
library(here)
library(PNWColors)
library(agricolae)
library(lme4)
library(lmerTest)
library(modelsummary)
library(performance)
library(purrr)
library(nlstools)
library(stringr)
library(emmeans)
library(modelr)
```

## Bring in datasets
```{r}
RecruitmentCounts <- read_csv(here("Data","Recruitment_Tiles_Counts.csv"))
EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

AllChemData_August <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/August2021/Allbiogeochemdata_QC.csv")
AllChemData_March <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/March2022/Allbiogeochemdata_QC_march_fdom2.csv")

```

## Define groupings to have SGD categories 
- Because without groupings, each tile stands alone as its own replicate which is not strong statistically 
- So make three groups of low medium and high groundwater... maybe based on silicate? 

```{r}

## edit old dfs to remove all unnecessary information 
AllChemData_August_edit <- AllChemData_August %>% 
  dplyr::select(c(1,4,19))

AllChemData_March_edit <- AllChemData_March %>% 
  dplyr::select(c(1,4,21))


FullChem <- AllChemData_August_edit %>% 
  full_join(AllChemData_March_edit)

## now group by Cow Tag ID and average across season and tides to get one value for each CowTag
FullChem_avg <- FullChem %>% 
  drop_na() %>% 
  filter(CowTagID!="VRC",CowTagID!="VSEEP",CowTagID!="VSPRING",CowTagID!="Varari_Well",CowTagID!="CSPRING_ROAD",CowTagID!="CSPRING_BEACH2",CowTagID!="CSPRING_BEACH",CowTagID!="CSPRING",CowTagID!="CSEEP",CowTagID!="CRC",CowTagID!="CPIT_Bottom",CowTagID!="CPIT") %>% 
  group_by(CowTagID, Location) %>% 
  summarize(Silicate_umolL = mean(Silicate_umolL))

## separate out Varari and Cabral 
FullChem_avg_V <- FullChem_avg %>% 
  filter(Location=="Varari")


FullChem_avg_C <- FullChem_avg %>% 
  filter(Location=="Cabral")

############################
## now create groupings 
############################
# 20 plates/CowTags, 3 groups --> ~6 per group 

FullChem_avg_C_org <- arrange(FullChem_avg_C, Silicate_umolL)
#write_csv(FullChem_avg_C_org, here::here("Data", "FullChem_avg_C.csv"))
FullChem_avg_C_SGDlevel <- read_csv(here("Data", "FullChem_avg_C.csv"))

FullChem_avg_V_org <- arrange(FullChem_avg_V, Silicate_umolL)
#write_csv(FullChem_avg_V_org, here::here("Data", "FullChem_avg_V.csv"))
FullChem_avg_V_SGDlevel <- read_csv(here("Data", "FullChem_avg_V.csv"))


### join them together 
SGDlevels_bothsites <- FullChem_avg_V_SGDlevel %>% 
  full_join(FullChem_avg_C_SGDlevel)
```

### Groupings = 
Cabral: 
- Low = C6, C15, C18, C7, C9, C3 (<2.5)
- Medium = C14, C12, C11, C9, C16, C8 ( 2.5 < x < 4.0 ) 
- High = C20, C4, C1, C5, C10, C17, C13, C2 (>4.0)

Varari: 
- Low = V19, V16, V4, V20, V9, V12, V10 (<1.5)
- Medium = V13, V5, V3, V11, V2, V15, V1 ( 1.5 < x < 2.5)
- High = V8, V18, V6, V14, V17, V7 (> 2.5) 

### Bring in recruitment data to link with low medium and high categories  

```{r}
EnvData_wrecruitcounts_edited <- EnvData_wrecruitcounts %>% 
  dplyr::select(c(1:5))

### join recruitment data with env data 
Grouped_recruitment <- SGDlevels_bothsites %>% 
  full_join(EnvData_wrecruitcounts_edited)

```

### Plot it out so can see differences by SGD groups 
```{r}
### ONE way is to total ALL the receruits for the plates in that SGD group
Grouped_recruitment_totaled <- Grouped_recruitment %>% 
  group_by(SGD_level, Location) %>% 
  summarize(grouptotal= sum(sum_total), 
            groupdead = sum(sum_total_dead), 
            groupalive = sum(sum_total_alive))

plot1_recruitsbySGDlevels <- Grouped_recruitment_totaled %>% 
  ggplot(aes(x=SGD_level, 
             y=grouptotal, 
             fill=Location)) + 
  geom_col() + 
  theme_classic() + 
  facet_wrap(~Location) + 
  scale_x_discrete(limits=c("Low","Medium", "High")) + 
  labs(title = "Total Settlers by SGD Level") + 
  scale_fill_manual(values=c("gold2", "firebrick4"))

plot1_recruitsbySGDlevels

### ANOTHER WAY is to look at total alive vs total dead 

## pivot df to useful orientation 
Grouped_recruitment_totaled_pivot <- Grouped_recruitment_totaled %>% 
  pivot_longer(cols = groupdead:groupalive, names_to = "Dead_Alive", values_to="deadalive_counts")

plot2_recruitsbySGDlevels <- Grouped_recruitment_totaled_pivot %>% 
  ggplot(aes(x=SGD_level, 
             y=deadalive_counts, 
             fill=Dead_Alive)) + 
  geom_col(position="dodge") + 
  theme_classic() + 
  facet_wrap(~Location) + 
  scale_x_discrete(limits=c("Low","Medium", "High")) + 
  labs(title = "Dead vs Alive Settlers by SGD Level") + 
  scale_fill_manual(values=c("coral3", "gray"))

plot2_recruitsbySGDlevels

plot2_recruitsbySGDlevels + plot1_recruitsbySGDlevels
ggsave(here::here("Outputs", "RecruitmentTiles", "Results", "Prelim", "RecruitCountsbySGDlevel.jpg"), 
       width=10, height=7)

```


### create model to test for significance 
```{r}
SGDlevel_recruits_model <- lm(sum_total ~ SGD_level*Location, data = Grouped_recruitment)

summary(SGDlevel_recruits_model)
Anova(SGDlevel_recruits_model, type = "III")

plot(SGDlevel_recruits_model)
qqp(SGDlevel_recruits_model)

AIC(SGDlevel_recruits_model)

```





