---
title: "Table_EnvData"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries 
```{r}

library(tidyverse)
library(here)
library(kableExtra)
library(webshot2)
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(magick)

```


## Read in Data 
```{r, echo=FALSE, message=FALSE}

Table_EnvData <- read_csv(here("Data", "2024_WaterChemMeasurements.csv"))

#### data with settler counts and salinity 
SettlementCount_wsalinity <- read_csv(here::here("Data", "SettlementCounts_edited_wSALINITY.csv"))

SettlementCount_wsalinity <- SettlementCount_wsalinity %>% 
  filter(settled!=0)

SettlementCount_wsalinitypH <- read_csv(here::here("Data", "SettlementCounts_edited_wSALINITYpH.csv"))

SettlementCount_wsalinitypH <- SettlementCount_wsalinitypH %>% 
  filter(settled!=0)

WaterChem_withpH <- read_csv(here::here("Data", "pHSlope.csv"))

AllChemData_August <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/August2021/Allbiogeochemdata_QC.csv")

AllChemData_March <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/March2022/Allbiogeochemdata_QC_march_fdom2.csv")

```
# Settlement Experiment 
## Create table with environmental data 
```{r}
###############################
## in case someone wants to see all the raw values 
## too much for publication otherwise 
###############################
modTable <- WaterChem_withpH %>%  
  dplyr::select(!c("TrisCalDate":"TTris", "note", "sampling_period", "light", "experiment_number", "orion_pH", "mV", "mVTris")) %>%
  drop_na(TA) %>% 
  mutate(treatment_level = if_else(treatment_level == "Fresh", "Raw GW", treatment_level)) %>% 
  rename(Phosphate = phosphate, 
         Silicate = silicate, 
         Ammonia = ammonia,
         "Nitrate+Nitrite" = NN, 
         "Total Alkalinity" = TA, 
         "Treatment Rep" = treatmentID_rep, 
         "SGD Level"= treatment_level, 
         "Time of Measurement" = time, 
         Date = date, 
         Temperature = temp, 
         Salinity = salinity) 
  #mutate(AICc = round(AICc, 1)) %>% 
 
 # mutate(pVal = signif(pVal, 2)) %>% 
 # rename(Regression = Reg_Type,
       #  p = pVal,
       #  'Delta AICc' = delAICc) %>% 
 # separate(col=Parameter, into = c("Parameter", NA), sep = " +") %>% 
 # mutate(Y = factor(Y, levels = c("NbSpP", "NbFEsP", "Vol8D", "resSpp", "resFEp", "resVol")),
      #   Parameter = factor(Parameter,
                       #     levels = c("Rugosity", "Phosphate", "Nitrate+Nitrite", "pH", "Salinity", "Silicate", "Temperature"))) %>% 
 # arrange(date)

modTable

###############################
## Average for treatments so table is reasonable size 
###############################
modTable2 <- WaterChem_withpH %>%  
  dplyr::select(!c("TrisCalDate", "note", "sampling_period", "light", "experiment_number", "orion_pH")) %>% 
  drop_na(TA) %>% 
  mutate(treatment_level = if_else(treatment_level == "Fresh", "Raw GW", treatment_level)) %>% 
  group_by(Location, treatment_level) %>% ## removed time and just included date so it is even more manageable -- took out date
  summarize("AvgTemp" = mean(temp), 
            "AvgPhosphate" = mean(phosphate), 
            "AvgSilicate" = mean(silicate), 
            "AvgNitrateNitrite" = mean(NN), 
           # "AvgTotalAlkalinity" = mean(TA), 
            "AvgSalinity" = mean(salinity), 
            "AvgpH" = mean(pH)) %>% 
  mutate(AvgSalinity = round(AvgSalinity, 2), 
       #  AvgTotalAlkalinity = round(AvgTotalAlkalinity, 2), 
         AvgNitrateNitrite = round(AvgNitrateNitrite, 2), 
         AvgSilicate = round(AvgSilicate, 2), 
         AvgPhosphate = round(AvgPhosphate, 2), 
         AvgTemp = round(AvgTemp, 2), 
         AvgpH = round(AvgpH, 2)) %>% 
 # mutate(treatment_level = factor(treatment_level, levels = c("Low", "Medium", "High"))) %>%
  rename("SGD Level"= treatment_level, 
         #"Time of Measurement" = time, 
        # Date = date, 
         "Avg Temperature" = AvgTemp, 
         "Avg Phosphate" = AvgPhosphate, 
         "Avg Silicate" = AvgSilicate, 
         "Avg Nitrate+Nitrite" = AvgNitrateNitrite, 
       #  "Avg Total Alkalinity" = AvgTotalAlkalinity, 
         "Avg Salinity" = AvgSalinity, 
         "Avg pH" = AvgpH) 

modTable2

modTable2 <- modTable2 %>%
  mutate('SGD Level' = factor('SGD Level', levels = c('Raw GW', "Ambient", "Low", "Medium", "High")))  

# Create the table with the manually ordered SGD Level
EnvData_Table <- modTable2 %>%
  arrange('SGD Level') %>%  # Arrange by the ordered factor
  kbl() %>%
  kable_classic(html_font = "Times New Roman", font_size = 14) %>%
  row_spec(0, italic = TRUE, bold = TRUE) %>%
  pack_rows("Cabral", 1, 5) %>%
  pack_rows("Varari", 6, 10) %>%
  column_spec(1:6, background = "white")

```

## Re-design table with Kable 
```{r}
 
EnvData_Table <- modTable2 %>%
  arrange('SGD Level') %>%  # Arrange by the ordered factor
  kbl() %>%
  kable_classic(html_font = "Times New Roman", font_size = 14) %>%
  row_spec(0, italic = TRUE, bold = TRUE) %>%
  pack_rows("Cabral", 1, 5) %>%
  pack_rows("Varari", 6, 10) %>%
  column_spec(1:6, background = "white")

EnvData_Table

######################
## save table 
######################
EnvData_Table %>% 
  as_image(file = here("Outputs", "SettlementExperiment", "Thesis_Figures", "EnvData_Table.jpg"))



```


## Make a plot with averaged endmember groundwater data 

```{r}

modTable2_edit <- modTable2 %>% 
  rename(treatment_level ="SGD Level", 
         AvgTemp = "Avg Temperature",
         AvgPhosphate ="Avg Phosphate" , 
         AvgSilicate = "Avg Silicate", 
         AvgNitrateNitrite = "Avg Nitrate+Nitrite", 
         #AvgTotalAlkalinity = "Avg Total Alkalinity", 
         AvgSalinity = "Avg Salinity", 
         AvgpH = "Avg pH") %>% 
  dplyr::filter(treatment_level=="Raw GW"|treatment_level=="Ambient") %>% 
  group_by(Location, treatment_level) %>% 
  summarize(meantemp = mean(AvgTemp), 
            meanphos = mean(AvgPhosphate), 
            meansilicate = mean(AvgSilicate), 
            meanNN = mean(AvgNitrateNitrite),
         #   meanTA = mean(AvgTotalAlkalinity),
            meanpH = mean(AvgpH), 
            meansal = mean(AvgSalinity), 
            setemp = sd(AvgTemp) / sqrt(n()), 
            sephos = sd(AvgPhosphate) / sqrt(n()), 
            sesilicate = sd(AvgSilicate) / sqrt(n()),
            seNN = sd(AvgNitrateNitrite) / sqrt(n()), 
         #   seTA = sd(AvgTotalAlkalinity) / sqrt(n()), 
            sesal = sd(AvgSalinity) / sqrt(n()), 
            sepH = sd(AvgpH) / sqrt(n())) # %>% 
 # pivot_longer(cols = "meantemp":"meansal", names_to = "avg_params", values_to = "avg_values")

################
## plot 
################

## pH
endmember_plot_pH <- modTable2_edit %>% 
  filter(treatment_level=="Raw GW") %>% 
  ggplot(aes(x=Location, 
             y=meanpH, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meanpH - sepH, 
                    ymax = meanpH + sepH), 
                width = 0.2, position = position_dodge(0.7)) + 
  labs(x = "", 
      y = "") + 
 coord_cartesian(ylim = c(7, 8)) +
 # scale_y_continuous(limits = c(7, 8)) +
       # y = expression(Average~pH) + 
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_pH


## NN
endmember_plot_NN <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meanNN, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meanNN - seNN, 
                    ymax = meanNN + seNN), 
                width = 0.2, position = position_dodge(0.7)) + 
     labs(x = "", 
          y = "") + 
       # y = expression(Average~Nitrate~Nitrite(mu*mol~L^-1))) + 
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_NN

## silicate 

endmember_plot_sil <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meansilicate, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meansilicate - sesilicate, 
                    ymax = meansilicate + sesilicate), 
                width = 0.2, position = position_dodge(0.7)) + 
     labs(x = "", 
          y = "") + 
       # y = expression(Average~Silicate~(mu*mol~L^-1))) + 
   theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_sil

## salinity 

endmember_plot_sal <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meansal, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"),
                     guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meansal - sesal, 
                    ymax = meansal + sesal), 
                width = 0.2, position = position_dodge(0.7)) + 
     labs(x = "", 
          y = "") + 
        #y = expression(Average~Salinity~(psu))) + 
   theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_sal


## phosphate 

endmember_plot_phosphate <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meanphos, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meanphos - sephos, 
                    ymax = meanphos + sephos), 
                width = 0.2, position = position_dodge(0.7)) + 
   labs(x = "", 
        y = expression(Average~Phosphate~(mu*mol~L^-1))) + 
   theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_phosphate

## TA 

endmember_plot_TA <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meanTA, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meanTA - seTA, 
                    ymax = meanTA + seTA), 
                width = 0.2, position = position_dodge(0.7)) + 
    labs(x = "", 
        y = expression(Average~TA~(mu*mol~L^-1))) + 
   theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_TA

######################################
## add them all together 
######################################

endmember_plots <-  endmember_plot_sil + endmember_plot_NN + endmember_plot_sal + endmember_plot_pH

endmember_plots
ggsave(here::here("Outputs", "SettlementExperiment", "Thesis_Figures", "endmember_plots.jpg"), 
       width=12, height=8)

```

## ggridges plot of environmental data 
-  **DOES NOT INCLUDE AVERAGES BECAUSE WAS GETTING AN ERROR - JUST RAW VALUES** (only seems to work with settlement info NOT water chem sheet, would need to manually move pH data over)

```{r}

## re-arrange data for better format 

SettlementCount_salinitypH_pivot <- SettlementCount_wsalinitypH %>% 
  dplyr::select(!c("swimming":"spawning_day", "notes")) %>% 
  pivot_longer(cols = silicate:NN|salinity:pH, names_to= "parameters", values_to = "values")

# ggridges plot 

ggridges_plot <- SettlementCount_salinitypH_pivot %>% 
  mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = values, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
  facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot

##################
## silicate
##################
ggridges_plot_silicate <- SettlementCount_wsalinitypH %>% 
 # mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = silicate, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
 # facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot_silicate

##################
## nitrates
##################
ggridges_plot_NN <- SettlementCount_wsalinitypH %>% 
 # mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = NN, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
 # facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot_NN

##################
## salinity
##################
ggridges_plot_salinity <- SettlementCount_wsalinitypH %>% 
 # mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = salinity, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
 # facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot_salinity


##################
## pH
##################

ggridges_plot_pH <- SettlementCount_wsalinitypH %>% 
 # mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = pH, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
 # facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot_pH


## add together 
ggridges_plot_silicate + ggridges_plot_NN + ggridges_plot_salinity + ggridges_plot_pH


ggsave(here::here("Outputs", "SettlementExperiment", "Thesis_Figures", "GGridges.jpg"),
       width=12, height=10)




######################
## if should include averages instead to make the same as endmember then can use this df but was causing issues with ggridges 
######################

modTable3_edit <- modTable2 %>% 
  rename(treatment_level ="SGD Level", 
         AvgTemp = "Avg Temperature",
         AvgPhosphate ="Avg Phosphate" , 
         AvgSilicate = "Avg Silicate", 
         AvgNitrateNitrite = "Avg Nitrate+Nitrite", 
         AvgSalinity = "Avg Salinity",
         AvgTotalAlkalinity = "Avg Total Alkalinity") %>% 
  filter(treatment_level!="Fresh") %>% 
  pivot_longer(cols = AvgSilicate:AvgNitrateNitrite|AvgSalinity,  names_to= "parameters", values_to = "values")

modTable3_edit$Location <- factor(modTable3_edit$Location, levels = c("Cabral", "Varari"))
unique(modTable3_edit$Location)
```

# In situ recruitment tiles 

## average March and August together and get min and mean data 
```{r}

##### tidy and join dfs

AllChemData_August2 <- AllChemData_August %>% 
  select(!c("Top_Plate_ID":"Plate_Seep", "Time", "DateTime", "M_C":"Lignin_Like")) %>% 
  filter(CowTagID!="CSEEP", CowTagID!="VSEEP", CowTagID!="CSPRING", CowTagID!="Varari_Well") %>% 
  group_by(Location, Tide, Date) %>% 
  filter(Date!="2021-08-06") 

AllChemData_March$Date <- mdy(AllChemData_March$Date)

AllChemData_March2 <- AllChemData_March %>% 
  select(!c("Top_Plate_ID":"Plate_Seep", "Time", "DateTime", "TimeBlock", "M_C":"Lignin_Like")) %>% 
  filter(CowTagID!="CSEEP", CowTagID!="VSEEP", CowTagID!="VRC", CowTagID!="CRC", CowTagID!="CPIT", CowTagID!="CPIT_Bottom", CowTagID!="CSPRING_ROAD", CowTagID!="CSPRING_BEACH", CowTagID!="CSPRING_BEACH2", CowTagID!="VSPRING") %>% 
  filter(TA > 2096) ##outlier TA 

##### tidy and join dfs with percent cover data 

AllChemData_seasonaverage <- AllChemData_August2 %>% 
  full_join(AllChemData_March2)

full_CommunityRecruit <- PercentCover %>% 
  left_join(RecruitmentCounts, join_by(CowTagID, Location))

full_envdataRecruits_avgseason <- full_CommunityRecruit %>% 
  left_join(AllChemData_seasonaverage, join_by(Location, CowTagID, lat, lon)) 

#####################################
## calculating MIN and MEAN across both seasons  
#####################################
full_envdataRecruits_avgseason2 <- full_envdataRecruits_avgseason %>% 
  dplyr::select(!c("Date", "Tide", "Day_Night")) %>% 
  filter(!CowTagID %in% c("CSEEP","VSEEP","Reef_Ambient_1","Reef_Ambient_2","Reef_Ambient_3","Reef_SGD_1","Reef_SGD_2", "Reef_SGD_3","Sand_Ambient_1","Sand_Ambient_2","Sand_Ambient_3","Sand_SGD_1","Sand_SGD_2","RockWall_SGD_1", "RockWall_SGD_2", "RockWall_SGD_3")) %>% 
  group_by(CowTagID) %>% 
  dplyr::summarise('Mean Salinity'=mean(Salinity, na.rm=TRUE), 
                   'SE Salinity' = sd(Salinity) / sqrt(length(Salinity)), 
                   'Min Salinity' = min(Salinity, na.rm=TRUE), 
                   'Mean pH'=mean(pH, na.rm=TRUE),
                   'SE pH' = sd(pH) / sqrt(length(pH)), 
                   'Min pH' = min(pH, na.rm=TRUE),
                   'Mean Silicate'=mean(Silicate_umolL, na.rm=TRUE),
                   'SE Silicate' = sd(Silicate_umolL) / sqrt(length(Silicate_umolL)), 
                   'Min Silicate'=min(Silicate_umolL, na.rm=TRUE), 
                   'Mean NN'=mean(NN_umolL, na.rm=TRUE), 
                   'SE NN' = sd(NN_umolL) / sqrt(length(NN_umolL)),
                   'Min NN' = min(NN_umolL, na.rm=TRUE)) %>% 
  mutate(CowTagID = factor(CowTagID, levels = c("C1", "C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10", "C11", "C12", "C13", "C14", "C15", "C16", "C17", "C18", "C19", "C20", "V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14", "V15", "V16", "V17", "V18", "V19", "V20")))  

#####################################
## make a table   
#####################################
InsituEnvData_Table <- full_envdataRecruits_avgseason2 %>%
  arrange(CowTagID) %>% 
  kbl() %>%
  kable_classic(html_font = "Times New Roman", font_size = 12) %>%
  row_spec(0, italic = TRUE, bold = TRUE) %>%
 pack_rows("Cabral", 1, 20) %>%
  pack_rows("Varari", 21, 40) %>%
  column_spec(1:6, background = "white")

InsituEnvData_Table
InsituEnvData_Table %>% 
  as_image(file = here("Outputs", "RecruitmentTiles", "Thesis_Figures", "InsituEnvData_Table.jpg"))



#####################################
## calculating MIN and MEAN across WHOLE SITES
#####################################
full_envdataRecruits_avgseason3 <- full_envdataRecruits_avgseason %>% 
  dplyr::select(!c("Date", "Tide", "Day_Night")) %>% 
  filter(!CowTagID %in% c("CSEEP","VSEEP","Reef_Ambient_1","Reef_Ambient_2","Reef_Ambient_3","Reef_SGD_1","Reef_SGD_2", "Reef_SGD_3","Sand_Ambient_1","Sand_Ambient_2","Sand_Ambient_3","Sand_SGD_1","Sand_SGD_2","RockWall_SGD_1", "RockWall_SGD_2", "RockWall_SGD_3")) %>% 
  group_by(Location) %>% 
  dplyr::summarise('Mean Salinity'=mean(Salinity, na.rm=TRUE), 
                   'SE Salinity' = sd(Salinity) / sqrt(length(Salinity)), 
                   'Min Salinity' = min(Salinity, na.rm=TRUE), 
                   'Mean pH'=mean(pH, na.rm=TRUE),
                   'SE pH' = sd(pH) / sqrt(length(pH)), 
                   'Min pH' = min(pH, na.rm=TRUE),
                   'Mean Silicate'=mean(Silicate_umolL, na.rm=TRUE),
                   'SE Silicate' = sd(Silicate_umolL) / sqrt(length(Silicate_umolL)), 
                   'Min Silicate'=min(Silicate_umolL, na.rm=TRUE), 
                   'Mean NN'=mean(NN_umolL, na.rm=TRUE), 
                   'SE NN' = sd(NN_umolL) / sqrt(length(NN_umolL)),
                   'Min NN' = min(NN_umolL, na.rm=TRUE))  

#####################################
## make a table   
#####################################
InsituEnvData_Table2 <- full_envdataRecruits_avgseason3 %>%
 #arrange(CowTagID) %>% 
  kbl() %>%
  kable_classic(html_font = "Times New Roman", font_size = 14) %>%
  row_spec(0, italic = TRUE, bold = TRUE) %>%
 #pack_rows("Cabral", 1, 20) %>%
 # pack_rows("Varari", 21, 40) %>%
  column_spec(1:6, background = "white")

InsituEnvData_Table2



```

