---
title: "Poc Cover Data with Biogeochem Params and Recruit Counts"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

### Load libraries 
```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(factoextra)
library(devtools)
library(ggbiplot)
library(car)
```


### Load Data
```{r}
EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

Poc_cover_data <- read_csv(here::here("Data", "Species_Composition_2022_Danielle.csv"))

#PercentCover <- read_csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Full_Metadata.csv")

PercentCover3 <- read_csv(here::here("Data", "PercentCover3.csv"))

RecruitmentCounts2.1 <- read_csv((here::here("Data", "RecruitmentCounts2.1.csv")))
```

## December, 2025: 


## Tidy data and join in Danielle's Poc cover data 
-  a) filter out the sand and hard substrate categories and 
-  b) sum the totals by cowtagid to get percent cover of the biotics 
-  c) calcualte percent cover of P. acuta based on total amount of biotics in the area 
-  Add in 0s for the cowtags that don't have P acuta counts (confirmed with Danielle, August 2024)
```{r}

#######################
## NOVEMBER 7th, 2024 update 
######################

Corals <- c("Pocillopora acuta","Porites rus","Montipora verrilli","Montipora setosa","Montipora grisea","Porites irregularis","Porites lobata","Pocillopora verrucosa","Pavona chiriquiensis","Pavona cactus","Leptastrea transversa","Acropora","Acropora pulchra","Montipora capitata")

## creating df of just corals using Danielle's df + new "Corals" only df 
df_corals <- Poc_cover_data %>% 
  filter(Taxa %in% Corals, 
         Location %in% c("Cabral", "Varari")) %>% 
  filter(Date!="7/19/22"|CowTagID!="C8") %>% 
  group_by(CowTagID) %>% 
  summarize(coral_counts = sum(SpeciesCounts)) 

## filtering out just P. acuta 
df_Pacuta <- Poc_cover_data %>% 
  filter(Taxa == "Pocillopora acuta", 
         Location %in% c("Cabral", "Varari")) %>% 
  filter(Date!="7/19/22"|CowTagID!="C8") %>% 
  group_by(CowTagID) %>% 
  summarize(Pacuta_counts = sum(SpeciesCounts)) ## getting a column of just P acuta counts 


#Poc_cover_dataedit <- Poc_cover_data %>% 
  #filter(Location %in% c("Cabral", "Varari")) %>% 
 # filter(Date!="7/19/22"|CowTagID!="C8") %>% ##filtering out 7/19/22 bc focused only on 6/25 ?? and C8 bc ??? (coming back Dec 2025)
 # group_by(CowTagID) %>% 
 # summarize(total_counts = sum(SpeciesCounts)) ## getting total species counts including Poc and all other spp per CowTagID BUT STILL HAS DATE AND COWTAGID C8 IN DF 
    

## filtering out sand, bare rock, rubble, and maya's locations so that left with just living orgs 
Poc_cover_data2 <- Poc_cover_data %>% 
  dplyr::select(!c("Date", "PhotoNum", "Notes")) %>% 
  filter(Taxa!="Bare Rock",Taxa!="Sand",Taxa!="Rubble", Location!="Varari_Maya") %>% 
  group_by(CowTagID) 

## getting species count for all biotic orgs 
Poc_cover_data3 <- Poc_cover_data2 %>% 
  dplyr::summarise(total_count_biotics = sum(SpeciesCounts)) %>% 
  filter(CowTagID!="VSEEP") ## missing C14 

## getting just P acuta counts 
## NOTE: not all CowTagID have P acuta present 
Poc_cover_data4 <- Poc_cover_data2 %>% 
  filter(Taxa=="Pocillopora acuta") %>% 
  group_by(CowTagID) %>% 
  mutate(SpeciesCounts_Poc = SpeciesCounts) %>% 
  dplyr::select(!SpeciesCounts)
 
### joins correlated biotics counts for that CowTag -- only has CowTagIDs with P. acuta, missing data from CowTags without P. acuta  
Poc_cover_data5 <- Poc_cover_data4 %>% 
  left_join(Poc_cover_data3) %>% 
  group_by(CowTagID, Location, SpeciesCounts_Poc,total_count_biotics) %>% 
  dplyr::summarise(percent_cover = (SpeciesCounts_Poc/total_count_biotics)*(100)) 

##########################################
### join together with recruit count env data df and then make sure to add in the zeros for P acuta cowtags 
##########################################
EnvData_wPocCover <- EnvData_wrecruitcounts %>% 
  full_join(Poc_cover_data5) %>% 
  mutate(percent_cover = ifelse(is.na(percent_cover), 0, percent_cover)) %>% ###  ADD IN ZERO FOR COWTAGS WITHOUT P ACUTA COUNTS 
  write_csv(here::here("Data", "EnvData_wPocCover.csv"))

##########################################
## using Poc data WITH hard substrate and sand etc 
###########################################
Poc_cover_data_withhardsub <- Poc_cover_data %>% 
  dplyr::select(!c("Date", "PhotoNum", "Notes")) %>% 
  filter(Location!="Varari_Maya") %>% 
  group_by(CowTagID) 
  
###########################
## creating "full" datatset combining all the necessary info for env parameters, settlement counts, percent cover of adult coral, etc
###########################

#full <- Poc_cover_dataedit %>% 
#  left_join(df_Pacuta) %>% 
#  left_join(df_corals) %>% 
#  filter(!CowTagID %in% c("CSEEP","VSEEP")) %>% 
 # mutate(Pacuta_counts = ifelse(is.na(Pacuta_counts), 0, Pacuta_counts), 
      #   coral_counts = ifelse(is.na(coral_counts), 0, coral_counts)) %>%
 # mutate(pacuta_cover = (Pacuta_counts/total_counts)*100, 
     #    coral_percentcover = (coral_counts/total_counts)*100) %>% 
#  left_join(SettlementDF_Full) %>% ## need to bring/load this dataset at beginning of script!! 
 # write_csv(here::here("Data", "full.csv"))

#full2 <- full %>% 
 # group_by(Location) %>% 
 # summarize(totalPacuta_site = sum(Pacuta_counts), 
        #    totalcoralcover_site = sum(coral_counts))

#full3 <- full2 %>% 
 # group_by(Location) %>% 
 # summarize(totalPacuta_site_percent = (totalPacuta_site/totalcoralcover_site)*100)

```


## Look at recruit totals by how many P acuta at the CowTagID
-  might need to square root the percent cover data bc of the variances (shown in recruits_by_PocCover)
-  faceted plot = non-transformed percent cover

```{r}

recruits_by_PocCover <- EnvData_wPocCover %>% 
  ggplot(aes(x=sqrt(percent_cover), 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              formula="y~poly(x,2)", 
              fill="cornsilk3", 
              alpha=0.2) #+ ## alpha adjusts the transparency 
 #labs(x="% Cover of P. acuta Species found near tile location", 
   #    y="Logged Total of Recruits")

recruits_by_PocCover
ggsave(here::here("Outputs", "RecruitmentTiles", "recruits_by_PocCover.jpg"))

recruits_by_PocCover_facet <- EnvData_wPocCover %>% 
  ggplot(aes(x=percent_cover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm,
              formula="y~poly(x,2)", 
              fill="cornsilk3", 
              alpha=0.2) + ## alpha adjusts the transparency 
  facet_wrap(~Location, scales="free") + 
  labs(x="% Cover of P. acuta Species found near tile location", 
       y="Logged Total of Recruits")

recruits_by_PocCover_facet
ggsave(here::here("Outputs", "RecruitmentTiles", "Results", "Prelim", "recruits_by_PocCover_facet.jpg"))

```

-  run models comparing nonlinear vs linear 
-  **Results**: in the linear model (whether transformed or not), there is a significant effect of cover on the number of settlers but no difference in the average cover or settlers between locations   

```{r}

nonlinear_PacutaCover <- lm(log(sum_total+1) ~ percent_cover + percent_cover^2 , data = EnvData_wPocCover)

anova(nonlinear_PacutaCover)
summary(nonlinear_PacutaCover)
AIC(nonlinear_PacutaCover)

#### something is up with polynomial models (not working for the second degree)

### with sqrt transformation 
linear_PacutaCover_sqrt <- lm(log(sum_total+1) ~ sqrt(percent_cover)*Location, data = EnvData_wPocCover)

anova(linear_PacutaCover_sqrt)
summary(linear_PacutaCover_sqrt)
AIC(linear_PacutaCover_sqrt)

### withOUT sqrt transformation 
linear_PacutaCover <- lm(log(sum_total+1) ~ percent_cover*Location, data = EnvData_wPocCover)

anova(linear_PacutaCover)
summary(linear_PacutaCover)
AIC(linear_PacutaCover)

```


### Transform Varari data 
-  try log, sqrt, cube root, reciprocal, exponential, **arc sine**, z-score, and rank transformations
-  honestly they all look the same other than logit 
-  Ended up going with: 


```{r}
## check normality for Cabral 
EnvData_wPocCover_C <- EnvData_wPocCover %>% 
  filter(Location=="Cabral") 

qqp(EnvData_wPocCover_C$percent_cover, "norm") ## pretty good 

## check normality for Varari 
EnvData_wPocCover_V <- EnvData_wPocCover %>% 
  filter(Location=="Varari") 

qqp(EnvData_wPocCover_V$percent_cover, "norm") ## not as good 

########################################
### transform Varari data 
########################################

### arc sine transformation 
EnvData_wPocCover_Vtransformed <- EnvData_wPocCover %>% 
  filter(Location=="Varari") %>% 
  mutate(percent_cover = asin(sqrt(percent_cover/100)))

recruits_by_PocCover_facet_transformed_V <- EnvData_wPocCover_Vtransformed %>% 
  ggplot(aes(x=transformed_percentcover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm,
              formula="y~poly(x,2)", 
              fill="cornsilk3", 
              alpha=0.2) #+ ## alpha adjusts the transparency 
 # facet_wrap(~Location, scales="free") + 
 # labs(x="Transformed % Cover of P. acuta Species found near tile location", 
      # y="Logged Total of Recruits")

recruits_by_PocCover_facet_transformed_V

qqp(EnvData_wPocCover_Vtransformed$transformed_percentcover, "norm")


### log transformation 
EnvData_wPocCover_Vtransformed2 <- EnvData_wPocCover %>% 
  filter(Location=="Varari") %>% 
  mutate(transformed_percentcover = log(percent_cover + 1))

recruits_by_PocCover_facet_transformed_V2 <- EnvData_wPocCover_Vtransformed2 %>% 
  ggplot(aes(x=transformed_percentcover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm,
              formula="y~poly(x,2)", 
              fill="cornsilk3", 
              alpha=0.2) 

recruits_by_PocCover_facet_transformed_V2

qqp(EnvData_wPocCover_Vtransformed2$transformed_percentcover, "norm")

### square root transformation 
EnvData_wPocCover_Vtransformed3 <- EnvData_wPocCover %>% 
  filter(Location=="Varari") %>% 
  mutate(transformed_percentcover = sqrt(percent_cover + 1))

recruits_by_PocCover_facet_transformed_V3 <- EnvData_wPocCover_Vtransformed3 %>% 
  ggplot(aes(x=transformed_percentcover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm,
              formula="y~poly(x,2)", 
              fill="cornsilk3", 
              alpha=0.2) 

recruits_by_PocCover_facet_transformed_V3

qqp(EnvData_wPocCover_Vtransformed3$transformed_percentcover, "norm")


### logit transformation 
EnvData_wPocCover_Vtransformed4 <- EnvData_wPocCover %>% 
  filter(Location=="Varari") %>% 
  mutate(percent_cover_transformed = (percent_cover + 0.01) / 100,
         transformed_percentcover2 = log(percent_cover_transformed / (1 - percent_cover_transformed)))

recruits_by_PocCover_facet_transformed_V4 <- EnvData_wPocCover_Vtransformed4 %>% 
  ggplot(aes(x=transformed_percentcover2, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm,
              formula="y~poly(x,2)", 
              fill="cornsilk3", 
              alpha=0.2) 

recruits_by_PocCover_facet_transformed_V4


qqp(EnvData_wPocCover_Vtransformed4$transformed_percentcover2, "norm")


```

### join newly transformed data back in with Cabral data 
```{r}

transformedPocdata <- EnvData_wPocCover_Vtransformed %>% 
  bind_rows(EnvData_wPocCover_C)

## plot 
plot_transformedV <- transformedPocdata %>% 
  ggplot(aes(x=percent_cover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm,
              formula="y~poly(x,2)", 
              fill="cornsilk3", 
              alpha=0.2) + 
  facet_wrap(~Location, scales="free_x")

plot_transformedV

```






ignore below this line 
#######################################################################################
#######################################################################################

### create a similiar plot for recruits by just hard substrate 
Currently getting an error (June 1st, 2024) because don't have substrate type in the EnvData_wPocCover df anymore 
Should probably use PercentCover df and group into substrate -- might have code somehwere, currently tired 

```{r}
#EnvData_wPocCover_long <- EnvData_wPocCover %>% 
#  pivot_longer(cols=Sand:HardSubstrate, names_to = "SubstrateType", values_to = "SubstrateValues")

#recruits_by_HardSubstrate <- EnvData_wPocCover_long %>% 
 # ggplot(aes(x=SubstrateValues, 
       #      y=log(sum_total+1), 
        #     color=SubstrateType)) + 
#  geom_point(size=3) + 
#  theme_classic() + 
#  scale_color_manual(values=c("cornflowerblue", "salmon")) +
 # geom_smooth(method=lm, 
           #   fill="cornsilk3", 
            #  alpha=0.2) + 
 # labs(x="Percent Cover of Hard Substrate vs Sand", 
     #  y="Logged Total of Recruits")
#recruits_by_HardSubstrate

#ggsave(here("Outputs", "RecruitmentTiles", "recruits_by_SubstrateType.jpg"))


```

### what about just live coral cover? include Porites, etc 

```{r}
CommComp_recruit_coralcover <- PercentCover3 %>% 
  left_join(RecruitmentCounts2.1, join_by(CowTagID, Location))

CommComp_recruit_coralcover_liveonly <- CommComp_recruit_coralcover %>% 
  filter(comm_type=="LiveCoral", 
         CowTagID!="CSEEP", CowTagID!="VSEEP")

#CommComp_recruit_seasonavg_envdata_wLiveCover <- CommComp_recruit_coralcover_liveonly %>% 
 # left_join(AllChemDataSeasonAvg_calculatedvalues, join_by(CowTagID)) 

#recruits_by_AllLiveCoral <- CommComp_recruit_seasonavg_envdata_wLiveCover %>% 
 # ggplot(aes(x=percent_cover, 
          #   y=log(sum_total+1), 
           #  color=Location)) + 
 # geom_point(size=3) + 
  #theme_classic() + 
#  scale_color_manual(values=c("cornflowerblue", "salmon")) +
#  geom_smooth(method=lm, 
         #     fill="cornsilk3", 
        #      alpha=0.2) + 
 # facet_wrap(~Location) + 
 # labs(x="Percent Live Coral", 
     #  y= "Log of Total Recruits")
#
#recruits_by_AllLiveCoral

#ggsave(here::here("Outputs", "RecruitmentTiles", "recruits_by_AllLiveCoral.jpg"))



```

### run models comparing total live coral cover vs Poc cover at Varari 
-  Results: 
Varari: 
- Poc cover: 0.044
- Live coral: 0.0184
Cabral: 
- Poc cover: 0.54
- Live coral: 0.44
```{r}

######################
## all live coral including both sites 
######################
#model_livecoral <- lm(log(sum_total+1) ~ percent_cover + Location, data= CommComp_recruit_seasonavg_envdata_wLiveCover) 
#summary(model_livecoral)

######################
## just Varari 
######################

#CommComp_recruit_coralcover_liveonly_V <- CommComp_recruit_seasonavg_envdata_wLiveCover %>% 
 # filter(Location=="Varari")

#model_Vlivecoral <- lm(log(sum_total+1) ~ percent_cover, data= CommComp_recruit_coralcover_liveonly_V) 
#summary(model_Vlivecoral)

######################
## just Cabral 
######################

#CommComp_recruit_coralcover_liveonly_C <- CommComp_recruit_coralcover_liveonly %>% 
 # filter(Location=="Cabral")

#model_Clivecoral <- lm(log(sum_total+1) ~ percent_cover, data= CommComp_recruit_coralcover_liveonly_C) 
#summary(model_Clivecoral)

###################################################################
######################
## just Poc cover 
######################
#EnvData_wPocCover_V <- EnvData_wPocCover %>% 
 # filter(Location=="Varari")

#model_VPocCover <- lm(log(sum_total+1) ~ SpeciesCounts, data= EnvData_wPocCover_V) 
#summary(model_VPocCover)


#EnvData_wPocCover_C <- EnvData_wPocCover %>% 
 # filter(Location=="Cabral")

#model_CPocCover <- lm(log(sum_total+1) ~ SpeciesCounts, data= EnvData_wPocCover_C) 
#summary(model_CPocCover)

```



