---
title: "PCA for Environmental Parameters along natural SGD gradient/recruitment tiles"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

### Load libraries 
```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(factoextra)
library(devtools)
library(ggbiplot)
```

### Load Data
```{r}
EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

```

## tidy data for a PCA 
```{r}
EnvData_wrecruitcounts2 <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) ## confirm if this should still be logged 

EnvData_wrecruitcounts2$sum_total <- as.character(EnvData_wrecruitcounts2$sum_total)
```

## start building PCA 
```{r}

biogeochem1 <- EnvData_wrecruitcounts2[,-c(1,2)] # remove first column so it is just biogeochem data to get PCs

biogeochem.scale <- scale(biogeochem1, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel <- princomp(biogeochem.scale, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel)

biogeochem_PCAmodel$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
biogeochem_PC1 <- biogeochem_PCAmodel$scores[,1] # Pulls out the first column of the table 
biogeochem_PC1

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model1 <- lm(biogeochem_PC1~EnvData_wrecruitcounts2$sum_total*EnvData_wrecruitcounts2$Location)
anova(biogeochem_model1)

biplot(biogeochem_PCAmodel, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 

```


## use ggbiplot to make nicer output 
```{r}
# In this biplot code, I'm going to separate the recruits by number (so use non-subsetted data file). use the original file to refer to recruits, but the PCA data for everything else. I'm also going to put confidence intervals around the recruit counts 

biogeochem_PCAplot <- biogeochem_PCAmodel %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1, 
           groups=EnvData_wrecruitcounts2$Location, 
           ellipse=TRUE, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts2$Location)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8))

biogeochem_PCAplot
ggsave(here("Outputs", "RecruitmentTiles", "PCA_plot.jpg")) 


```


#### Everything above included BOTH Varari and Cabral 
## now separate the two - Varari first 
## and only include means and sd
```{r}

EnvData_wrecruitcounts_V <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) %>% 
  filter(Location=="Varari") %>% 
  select(!"Location")

EnvData_wrecruitcounts_V$sum_total <- as.character(EnvData_wrecruitcounts_V$sum_total)

##########################
###########################
biogeochem_V <- EnvData_wrecruitcounts_V[,-1] # remove first and second columns so it is just biogeochem data to get PCs

Vbiogeochem.scale <- scale(biogeochem_V, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel_V <- princomp(Vbiogeochem.scale, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel_V) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel_V, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel_V)

biogeochem_PCAmodel_V$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel_V$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
Vbiogeochem_PC1 <- biogeochem_PCAmodel_V$scores[,1] # Pulls out the first column of the table 
Vbiogeochem_PC1

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model_V <- lm(Vbiogeochem_PC1~EnvData_wrecruitcounts_V$sum_total)
anova(biogeochem_model_V)

biplot(biogeochem_PCAmodel_V, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 

###########################
###########################

biogeochem_PCAplot_V <- biogeochem_PCAmodel_V %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1,
           groups=EnvData_wrecruitcounts_V$sum_total, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts_V$sum_total)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8)) + 
  labs(title="Varari")

biogeochem_PCAplot_V
ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_V.jpg")) 

```

## now Cabral 
```{r}

EnvData_wrecruitcounts_C <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) %>% 
  filter(Location=="Cabral") %>% 
  select(!"Location")

EnvData_wrecruitcounts_C$sum_total <- as.character(EnvData_wrecruitcounts_C$sum_total)

##########################
###########################
biogeochem_C <- EnvData_wrecruitcounts_C[,-1] # remove first and second columns so it is just biogeochem data to get PCs

Cbiogeochem.scale <- scale(biogeochem_C, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel_C <- princomp(Cbiogeochem.scale, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel_C) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel_C, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel_C)

biogeochem_PCAmodel_C$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel_C$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
Cbiogeochem_PC1 <- biogeochem_PCAmodel_C$scores[,1] # Pulls out the first column of the table 
Cbiogeochem_PC1

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model_C <- lm(Cbiogeochem_PC1~EnvData_wrecruitcounts_C$sum_total)
anova(biogeochem_model_C)

biplot(biogeochem_PCAmodel_C, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 

###########################
###########################

biogeochem_PCAplot_C <- biogeochem_PCAmodel_C %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1, 
           groups=EnvData_wrecruitcounts_C$sum_total, 
           ellipse=TRUE, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts_C$sum_total)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8)) + 
  labs(title="Cabral")

biogeochem_PCAplot_C
ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_C.jpg")) 


biogeochem_PCAplot_C + biogeochem_PCAplot_V
ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_bothsites.jpg"), 
       width=12, height=8) 

```

### PCA without temperature 
```{r}
EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

##### Varari 
EnvData_wrecruitcounts3_V <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) %>% ## confirm if this should still be logged 
  select(!c("temp_avg", "temp_sd")) %>% 
  filter(Location=="Varari") %>% 
  select(!"Location")

EnvData_wrecruitcounts3_V$sum_total <- as.character(EnvData_wrecruitcounts3_V$sum_total)

##### Cabral 
EnvData_wrecruitcounts3_C <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) %>% ## confirm if this should still be logged 
  select(!c("temp_avg", "temp_sd")) %>% 
  filter(Location=="Cabral") %>% 
  select(!"Location")

EnvData_wrecruitcounts3_C$sum_total <- as.character(EnvData_wrecruitcounts3_C$sum_total)

######################################################
######################################################

####################################
#### PCA without temp at Varari 
####################################
biogeochem_notempV <- EnvData_wrecruitcounts3_V[,-1] # remove first and second columns so it is just biogeochem data to get PCs

Vbiogeochem.scale_notemp <- scale(biogeochem_notempV, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel_V_notemp <- princomp(Vbiogeochem.scale_notemp, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel_V_notemp) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel_V_notemp, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel_V_notemp)

biogeochem_PCAmodel_V_notemp$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel_V_notemp$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
Vbiogeochem_PC1_notemp <- biogeochem_PCAmodel_V_notemp$scores[,1] # Pulls out the first column of the table 
Vbiogeochem_PC1_notemp

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model_V_notemp <- lm(Vbiogeochem_PC1_notemp~EnvData_wrecruitcounts3_V$sum_total)
anova(biogeochem_model_V_notemp)

biplot(biogeochem_PCAmodel_V_notemp, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 
###########################

biogeochem_PCAplot_V_notemp <- biogeochem_PCAmodel_V_notemp %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1,
           groups=EnvData_wrecruitcounts3_V$sum_total, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts3_V$sum_total)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8)) + 
  labs(title="Varari Without Temperature")

biogeochem_PCAplot_V_notemp
#ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_V.jpg")) 


#########################
## Cabral no temp 
#########################

biogeochem_notempC <- EnvData_wrecruitcounts3_C[,-1] # remove first and second columns so it is just biogeochem data to get PCs

Cbiogeochem.scale_notemp <- scale(biogeochem_notempC, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel_C_notemp <- princomp(Cbiogeochem.scale_notemp, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel_C_notemp) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel_C_notemp, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel_C_notemp)

biogeochem_PCAmodel_C_notemp$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel_C_notemp$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
Cbiogeochem_PC1_notemp <- biogeochem_PCAmodel_C_notemp$scores[,1] # Pulls out the first column of the table 
Cbiogeochem_PC1_notemp

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model_C_notemp <- lm(Cbiogeochem_PC1_notemp~EnvData_wrecruitcounts3_C$sum_total)
anova(biogeochem_model_C_notemp)

biplot(biogeochem_PCAmodel_C_notemp, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 

###########################
###########################

biogeochem_PCAplot_C_notemp <- biogeochem_PCAmodel_C_notemp %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1,
           groups=EnvData_wrecruitcounts3_C$sum_total, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts3_C$sum_total)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8)) + 
  labs(title="Cabral Without Temperature")

biogeochem_PCAplot_C_notemp
#ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_V.jpg")) 


##########################################
##########################################
biogeochem_PCAplot_C_notemp + biogeochem_PCAplot_V_notemp

```
### remove all the sd and just use averages 
```{r}

EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

##### Varari 
EnvData_wrecruitcounts3_V_avgonly <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) %>% ## confirm if this should still be logged 
  select(!c("sal_sd":"ammonia_sd")) %>% 
  filter(Location=="Varari") %>% 
  select(!"Location")

EnvData_wrecruitcounts3_V_avgonly$sum_total <- as.character(EnvData_wrecruitcounts3_V_avgonly$sum_total)

##### Cabral 
EnvData_wrecruitcounts3_C_avgonly <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) %>% ## confirm if this should still be logged 
   select(!c("sal_sd":"ammonia_sd")) %>% 
  filter(Location=="Cabral") %>% 
  select(!"Location")

EnvData_wrecruitcounts3_C_avgonly$sum_total <- as.character(EnvData_wrecruitcounts3_C_avgonly$sum_total)

######################################################
######################################################

####################################
#### PCA without temp at Varari 
####################################
biogeochem_V_avgonly <- EnvData_wrecruitcounts3_V_avgonly[,-1] # remove first and second columns so it is just biogeochem data to get PCs

Vbiogeochem.scale_avgonly <- scale(biogeochem_V_avgonly, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel_V_avgonly <- princomp(Vbiogeochem.scale_avgonly, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel_V_avgonly) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel_V_avgonly, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel_V_avgonly)

biogeochem_PCAmodel_V_avgonly$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel_V_avgonly$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
Vbiogeochem_PC1_avgonly <- biogeochem_PCAmodel_V_avgonly$scores[,1] # Pulls out the first column of the table 
Vbiogeochem_PC1_avgonly

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model_V_avgonly <- lm(Vbiogeochem_PC1_avgonly~EnvData_wrecruitcounts3_V_avgonly$sum_total)
anova(biogeochem_model_V_avgonly)

biplot(biogeochem_PCAmodel_V_avgonly, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 
###########################

biogeochem_PCAplot_V_avgonly <- biogeochem_PCAmodel_V_avgonly %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1,
           groups=EnvData_wrecruitcounts3_V_avgonly$sum_total, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts3_V_avgonly$sum_total)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8)) + 
  labs(title="Varari Average Only")

biogeochem_PCAplot_V_avgonly
#ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_V.jpg")) 


#########################
## Cabral no temp 
#########################

biogeochem_Cavgonly <- EnvData_wrecruitcounts3_C_avgonly[,-1] # remove first and second columns so it is just biogeochem data to get PCs

Cbiogeochem.scale_avgonly <- scale(biogeochem_Cavgonly, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel_C_avgonly <- princomp(Cbiogeochem.scale_avgonly, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel_C_avgonly) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel_C_avgonly, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel_C_avgonly)

biogeochem_PCAmodel_C_avgonly$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel_C_avgonly$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
Cbiogeochem_PC1_avgonly <- biogeochem_PCAmodel_C_avgonly$scores[,1] # Pulls out the first column of the table 
Cbiogeochem_PC1_avgonly

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model_C_avgonly <- lm(Cbiogeochem_PC1_avgonly~EnvData_wrecruitcounts3_C_avgonly$sum_total)
anova(biogeochem_model_C_avgonly)

biplot(biogeochem_PCAmodel_C_avgonly, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 

###########################
###########################

biogeochem_PCAplot_C_avgonly <- biogeochem_PCAmodel_C_avgonly %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1,
           groups=EnvData_wrecruitcounts3_C_avgonly$sum_total, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts3_C_avgonly$sum_total)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8)) + 
  labs(title="Cabral Average Only")

biogeochem_PCAplot_C_avgonly
#ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_V.jpg")) 


##########################################
##########################################
biogeochem_PCAplot_C_avgonly + biogeochem_PCAplot_V_avgonly

```


#### remove all the sd and just use averages  MINUS temperature 
```{r}

EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

##### Varari 
EnvData_wrecruitcounts3_V_avgonly_notemp <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) %>% ## confirm if this should still be logged 
  select(!c("sal_sd":"ammonia_sd", "temp_avg")) %>% 
  filter(Location=="Varari") %>% 
  select(!"Location")

EnvData_wrecruitcounts3_V_avgonly_notemp$sum_total <- as.character(EnvData_wrecruitcounts3_V_avgonly_notemp$sum_total)

##### Cabral 
EnvData_wrecruitcounts3_C_avgonly_notemp <- EnvData_wrecruitcounts %>% 
  select(!c("CowTagID":"HardSubstrate", "sum_total_dead", "sum_total_alive", "sal_max":"ammonia_CV")) %>% 
  drop_na() %>% 
  mutate(sum_total=log(sum_total+1)) %>% ## confirm if this should still be logged 
   select(!c("sal_sd":"ammonia_sd", "temp_avg")) %>% 
  filter(Location=="Cabral") %>% 
  select(!"Location")

EnvData_wrecruitcounts3_C_avgonly_notemp$sum_total <- as.character(EnvData_wrecruitcounts3_C_avgonly_notemp$sum_total)

######################################################
######################################################

####################################
#### PCA without temp at Varari 
####################################
biogeochem_V_avgonly_notemp <- EnvData_wrecruitcounts3_V_avgonly_notemp[,-1] # remove first and second columns so it is just biogeochem data to get PCs

Vbiogeochem.scale_avgonly_notemp <- scale(biogeochem_V_avgonly_notemp, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel_V_avgonly_notemp <- princomp(Vbiogeochem.scale_avgonly_notemp, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel_V_avgonly_notemp) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel_V_avgonly_notemp, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel_V_avgonly_notemp)

biogeochem_PCAmodel_V_avgonly_notemp$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel_V_avgonly_notemp$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
Vbiogeochem_PC1_avgonly_notemp <- biogeochem_PCAmodel_V_avgonly_notemp$scores[,1] # Pulls out the first column of the table 
Vbiogeochem_PC1_avgonly_notemp

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model_V_avgonly_notemp <- lm(Vbiogeochem_PC1_avgonly_notemp~EnvData_wrecruitcounts3_V_avgonly_notemp$sum_total)
anova(biogeochem_model_V_avgonly_notemp)

biplot(biogeochem_PCAmodel_V_avgonly_notemp, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 
###########################

biogeochem_PCAplot_V_avgonly_notemp <- biogeochem_PCAmodel_V_avgonly_notemp %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1,
           groups=EnvData_wrecruitcounts3_V_avgonly_notemp$sum_total, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts3_V_avgonly_notemp$sum_total)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8)) + 
  labs(title="Varari Average Only Without Temperature")

biogeochem_PCAplot_V_avgonly_notemp
#ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_V.jpg")) 


#########################
## Cabral no temp 
#########################

biogeochem_Cavgonly_notemp <- EnvData_wrecruitcounts3_C_avgonly_notemp[,-1] # remove first and second columns so it is just biogeochem data to get PCs

Cbiogeochem.scale_avgonly_notemp <- scale(biogeochem_Cavgonly_notemp, scale=TRUE, center=TRUE) ## convert all data to z-scores 

biogeochem_PCAmodel_C_avgonly_notemp <- princomp(Cbiogeochem.scale_avgonly_notemp, cor=FALSE) # Run the PCA with `princomp()`

summary(biogeochem_PCAmodel_C_avgonly_notemp) ## look at the model
# This shows amount of variance explained by each axis.

plot(biogeochem_PCAmodel_C_avgonly_notemp, type="lines")

## prettier version 
fviz_eig(biogeochem_PCAmodel_C_avgonly_notemp)

biogeochem_PCAmodel_C_avgonly_notemp$loadings  # Shows the loadings of each trait on each PC 

biogeochem_PCAmodel_C_avgonly_notemp$scores # Gives output of the principal components for each individual

# If we want to use the PC scores for something else (maybe we want to use PC1 to run an ANOVA?), then we can pull out that vector.
Cbiogeochem_PC1_avgonly_notemp <- biogeochem_PCAmodel_C_avgonly_notemp$scores[,1] # Pulls out the first column of the table 
Cbiogeochem_PC1_avgonly_notemp

# And we could run a quick ANOVA to see if biogeochem differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)

biogeochem_model_C_avgonly_notemp <- lm(Cbiogeochem_PC1_avgonly_notemp~EnvData_wrecruitcounts3_C_avgonly_notemp$sum_total)
anova(biogeochem_model_C_avgonly_notemp)

biplot(biogeochem_PCAmodel_C_avgonly_notemp, xlab="PC1", ylab="PC2")
# The black numbers are each individual recruit total count. The red lines are the vectors for each biogeochem parameter 

###########################
###########################

biogeochem_PCAplot_C_avgonly_notemp <- biogeochem_PCAmodel_C_avgonly_notemp %>% 
  ggbiplot(obs.scale=1, 
           var.scale=1,
           groups=EnvData_wrecruitcounts3_C_avgonly_notemp$sum_total, 
           varname.size=3, 
           varname.adjust=1.2, 
           circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(EnvData_wrecruitcounts3_C_avgonly_notemp$sum_total)), 
             size = 3) + 
  theme_classic() +
  theme(legend.direction = 'horizontal', 
        legend.position='bottom', 
        legend.text=element_text(size=8)) + 
  labs(title="Cabral Average Only Without Temperature")

biogeochem_PCAplot_C_avgonly_notemp
ggsave(here::here("Outputs", "RecruitmentTiles", "PCA_plot_C_notemp_avgs.jpg")) 


##########################################
##########################################
biogeochem_PCAplot_C_avgonly_notemp + biogeochem_PCAplot_V_avgonly_notemp


```

#### take the scores of PC1 and PC2 and plot as independent vars against log recruitment 

```{r}
### turn the PCA scores into df 
############ for Cabral first 
Cbiogeochem_PC1_avgonly_notemp <- biogeochem_PCAmodel_C_avgonly_notemp$scores[,c(1,2)]

C_PC1_df <- data.frame(Cbiogeochem_PC1_avgonly_notemp)

### now need to include logged recruit counts 

C_PC1_df2 <- C_PC1_df %>% 
  mutate(sum_total=c(3.4011974, 2.4849066, 1.0986123,2.8332133,2.7080502, 2.3978953, 3.7612001, 1.0986123, 0.0000000, 4.9904326, 0.6931472, 1.0986123, 1.9459101, 2.9444390, 2.1972246, 2.3025851, 1.3862944))

C_PC1_df2_long <- C_PC1_df2 %>% 
  pivot_longer(cols=Comp.1:Comp.2, names_to = "Comp1_2", values_to = "PCA_scores")

############ for Varari next  
Vbiogeochem_PC1_avgonly_notemp <- biogeochem_PCAmodel_V_avgonly_notemp$scores[,c(1,2)]

V_PC1_df <- data.frame(Vbiogeochem_PC1_avgonly_notemp)

V_PC1_df2 <- V_PC1_df %>% 
  mutate(sum_total=c(4.543295, 4.143135, 4.077537, 4.882802, 1.098612, 3.526361, 3.135494, 2.079442, 2.639057, 1.791759, 4.574711, 1.609438, 3.332205, 3.806662, 3.496508, 1.098612, 2.995732))

V_PC1_df2_long <- V_PC1_df2 %>% 
  pivot_longer(cols=Comp.1:Comp.2, names_to = "Comp1_2", values_to = "PCA_scores")

#########################################################
#########################################################

########### ggplot ##################
V_PCscores_recruits <- V_PC1_df2_long %>% 
  ggplot(aes(x=PCA_scores, 
             y=sum_total, 
             color=Comp1_2)) + 
  geom_point(size=3) + 
  theme_bw() + 
  facet_wrap(~Comp1_2) + 
  labs(title="Varari") + 
  geom_smooth(method="lm", formula="y~poly(x,2)")
  
V_PCscores_recruits


### for Cabral 
############ ggplot 

C_PCscores_recruits <- C_PC1_df2_long %>% 
  ggplot(aes(x=PCA_scores, 
             y=sum_total, 
             color=Comp1_2)) + 
  geom_point(size=3) + 
  theme_bw() + 
  facet_wrap(~Comp1_2) + 
  labs(title="Cabral") + 
  geom_smooth(method="lm")
  
C_PCscores_recruits
ggsave(here("Outputs", "RecruitmentTiles", "CabralPCscores.jpg"))

C_PCscores_recruits / V_PCscores_recruits

```

## are these models significant? 
```{r}
### Varari 
V_PCAmodel <- lm(sum_total ~ PCA_scores * Comp1_2 ,data = V_PC1_df2_long)
summary(V_PCAmodel)

V_PCAmodel2 <- lm(sum_total ~ poly(Comp.1, degree=2) ,data = V_PC1_df2)
summary(V_PCAmodel2)

## Cabral 
C_PCAmodel2 <- lm(sum_total ~ Comp.1 ,data = C_PC1_df2)
summary(C_PCAmodel2)

```

### BRING IN 3 tiles missing 





