---
title: "Poc Cover Data with Biogeochem Params and Recruit Counts"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

### Load libraries 
```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(factoextra)
library(devtools)
library(ggbiplot)
```


### Load Data
```{r}
EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

Poc_cover_data <- read_csv(here::here("Data", "Species_Composition_2022_Danielle.csv"))

PercentCover <- read_csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Full_Metadata.csv")

PercentCover3 <- read_csv(here::here("Data", "PercentCover3.csv"))

RecruitmentCounts2.1 <- read_csv((here::here("Data", "RecruitmentCounts2.1.csv")))
```

## Tidy data and join in Danielle's Poc cover data 
a) filter out the sand and hard substrate categories and b) sum the totals by cowtagid to get percent cover of the biotics c) calcualte percent cover of P. acuta based on total amount of biotics in the area 
```{r}
Poc_cover_data2 <- Poc_cover_data %>% 
  select(!c("Date", "PhotoNum", "Notes")) %>% 
  filter(Taxa!="Bare Rock",Taxa!="Sand",Taxa!="Rubble", Location!="Varari_Maya") %>% 
  group_by(CowTagID) 

Poc_cover_data3 <- Poc_cover_data2 %>% 
  dplyr::summarise(total_count_biotics = sum(SpeciesCounts)) %>% 
  filter(CowTagID!="VSEEP")

Poc_cover_data4 <- Poc_cover_data2 %>% 
  filter(Taxa=="Pocillopora acuta") %>% 
  group_by(CowTagID) %>% 
  mutate(SpeciesCounts_Poc = SpeciesCounts) %>% 
  dplyr::select(!SpeciesCounts)
 
### joins and drops CowTgs without P acuta data + correlated biotics counts for that CowTag 
Poc_cover_data5 <- Poc_cover_data4 %>% 
  left_join(Poc_cover_data3) %>% 
  group_by(CowTagID, Location, SpeciesCounts_Poc,total_count_biotics ) %>% 
  dplyr::summarise(percent_cover = (SpeciesCounts_Poc/total_count_biotics)*(100))

##########################################
### THIS REMOVES SOME OF THE TILES BC NOT ALL TILES HAD P. ACUTA COUNTS 
##########################################
EnvData_wPocCover <- EnvData_wrecruitcounts %>% 
  full_join(Poc_cover_data5) %>% 
  mutate(percent_cover = ifelse(is.na(percent_cover), 0, percent_cover)) %>% ###  ADD IN ZERO FOR COWTAGS WITHOUT P ACUTA COUNTS 
  write_csv(here::here("Data", "EnvData_wPocCover.csv"))

##########################################
## using Poc data WITH hard substrate and sand etc 
###########################################
Poc_cover_data_withhardsub <- Poc_cover_data %>% 
  select(!c("Date", "PhotoNum", "Notes")) %>% 
  filter(Location!="Varari_Maya") %>% 
  group_by(CowTagID) 
  
  

```

## Look at recruit totals by how many P acuta at the CowTagID
-  might need to square root the percent cover data bc of the variances 

```{r}

recruits_by_PocCover <- EnvData_wPocCover %>% 
  ggplot(aes(x=sqrt(percent_cover), 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              fill="cornsilk3", 
              alpha=0.2) #+ ## alpha adjusts the transparency 
 #labs(x="% Cover of P. acuta Species found near tile location", 
   #    y="Logged Total of Recruits")

recruits_by_PocCover
ggsave(here::here("Outputs", "RecruitmentTiles", "recruits_by_PocCover.jpg"))

recruits_by_PocCover_facet <- EnvData_wPocCover %>% 
  ggplot(aes(x=percent_cover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              fill="cornsilk3", 
              alpha=0.2) + ## alpha adjusts the transparency 
  facet_wrap(~Location, scales="free") + 
  labs(x="% Cover of P. acuta Species found near tile location", 
       y="Logged Total of Recruits")

recruits_by_PocCover_facet
ggsave(here::here("Outputs", "RecruitmentTiles", "Results", "Prelim", "recruits_by_PocCover_facet.jpg"))

```

-  run models comparing nonlinear vs linear 
-  **Results**: in the linear model (whether transformed or not), there is a significant effect of cover on the number of settlers but no difference in the average cover or settlers between locations   

```{r}

nonlinear_PacutaCover <- lm(log(sum_total+1) ~ percent_cover + percent_cover^2 , data = EnvData_wPocCover)

anova(nonlinear_PacutaCover)
summary(nonlinear_PacutaCover)
AIC(nonlinear_PacutaCover)

#### something is up with polynomial models (not working for the second degree)

### with sqrt transformation 
linear_PacutaCover_sqrt <- lm(log(sum_total+1) ~ sqrt(percent_cover)*Location, data = EnvData_wPocCover)

anova(linear_PacutaCover_sqrt)
summary(linear_PacutaCover_sqrt)
AIC(linear_PacutaCover_sqrt)

### withOUT sqrt transformation 
linear_PacutaCover <- lm(log(sum_total+1) ~ percent_cover*Location, data = EnvData_wPocCover)

anova(linear_PacutaCover)
summary(linear_PacutaCover)
AIC(linear_PacutaCover)

```











#######################################################################################
#######################################################################################

### create a similiar plot for recruits by just hard substrate 
Currently getting an error (June 1st, 2024) because don't have substrate type in the EnvData_wPocCover df anymore 
Should probably use PercentCover df and group into substrate -- might have code somehwere, currently tired 

```{r}
#EnvData_wPocCover_long <- EnvData_wPocCover %>% 
#  pivot_longer(cols=Sand:HardSubstrate, names_to = "SubstrateType", values_to = "SubstrateValues")

#recruits_by_HardSubstrate <- EnvData_wPocCover_long %>% 
 # ggplot(aes(x=SubstrateValues, 
       #      y=log(sum_total+1), 
        #     color=SubstrateType)) + 
#  geom_point(size=3) + 
#  theme_classic() + 
#  scale_color_manual(values=c("cornflowerblue", "salmon")) +
 # geom_smooth(method=lm, 
           #   fill="cornsilk3", 
            #  alpha=0.2) + 
 # labs(x="Percent Cover of Hard Substrate vs Sand", 
     #  y="Logged Total of Recruits")
#recruits_by_HardSubstrate

#ggsave(here("Outputs", "RecruitmentTiles", "recruits_by_SubstrateType.jpg"))


```

### what about just live coral cover? include Porites, etc 

```{r}
CommComp_recruit_coralcover <- PercentCover3 %>% 
  left_join(RecruitmentCounts2.1, join_by(CowTagID, Location))

CommComp_recruit_coralcover_liveonly <- CommComp_recruit_coralcover %>% 
  filter(comm_type=="LiveCoral", 
         CowTagID!="CSEEP", CowTagID!="VSEEP")

CommComp_recruit_seasonavg_envdata_wLiveCover <- CommComp_recruit_coralcover_liveonly %>% 
  left_join(AllChemDataSeasonAvg_calculatedvalues, join_by(CowTagID)) 

recruits_by_AllLiveCoral <- CommComp_recruit_seasonavg_envdata_wLiveCover %>% 
  ggplot(aes(x=percent_cover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              fill="cornsilk3", 
              alpha=0.2) + 
  facet_wrap(~Location) + 
  labs(x="Percent Live Coral", 
       y= "Log of Total Recruits")

recruits_by_AllLiveCoral

ggsave(here::here("Outputs", "RecruitmentTiles", "recruits_by_AllLiveCoral.jpg"))



```

### run models comparing total live coral cover vs Poc cover at Varari 
-  Results: 
Varari: 
- Poc cover: 0.044
- Live coral: 0.0184
Cabral: 
- Poc cover: 0.54
- Live coral: 0.44
```{r}

######################
## all live coral including both sites 
######################
model_livecoral <- lm(log(sum_total+1) ~ percent_cover + Location, data= CommComp_recruit_seasonavg_envdata_wLiveCover) 
summary(model_livecoral)

######################
## just Varari 
######################

CommComp_recruit_coralcover_liveonly_V <- CommComp_recruit_seasonavg_envdata_wLiveCover %>% 
  filter(Location=="Varari")

model_Vlivecoral <- lm(log(sum_total+1) ~ percent_cover, data= CommComp_recruit_coralcover_liveonly_V) 
summary(model_Vlivecoral)

######################
## just Cabral 
######################

CommComp_recruit_coralcover_liveonly_C <- CommComp_recruit_coralcover_liveonly %>% 
  filter(Location=="Cabral")

model_Clivecoral <- lm(log(sum_total+1) ~ percent_cover, data= CommComp_recruit_coralcover_liveonly_C) 
summary(model_Clivecoral)

###################################################################
######################
## just Poc cover 
######################
EnvData_wPocCover_V <- EnvData_wPocCover %>% 
  filter(Location=="Varari")

model_VPocCover <- lm(log(sum_total+1) ~ SpeciesCounts, data= EnvData_wPocCover_V) 
summary(model_VPocCover)


EnvData_wPocCover_C <- EnvData_wPocCover %>% 
  filter(Location=="Cabral")

model_CPocCover <- lm(log(sum_total+1) ~ SpeciesCounts, data= EnvData_wPocCover_C) 
summary(model_CPocCover)

```



