---
title: "Poc Cover Data with Biogeochem Params and Recruit Counts"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

### Load libraries 
```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(factoextra)
library(devtools)
library(ggbiplot)
```


### Load Data
```{r}
EnvData_wrecruitcounts <- read_csv(here::here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

Poc_cover_data <- read_csv(here::here("Data", "Species_Composition_2022_Danielle.csv"))

PercentCover <- read_csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Full_Metadata.csv")

PercentCover3 <- read_csv(here::here("Data", "PercentCover3.csv"))

RecruitmentCounts2.1 <- read_csv((here::here("Data", "RecruitmentCounts2.1.csv")))
```

## Tidy data and join in Danielle's Poc cover data 
a) filter out the sand and hard substrate categories and b) sum the totals by cowtagid to get percent cover of the biotics 
```{r}
Poc_cover_data2 <- Poc_cover_data %>% 
  select(!c("Date", "PhotoNum", "Notes")) %>% 
  filter(Taxa!="Bare Rock",Taxa!="Sand",Taxa!="Rubble", Location!="Varari_Maya") %>% 
  group_by(CowTagID) 

Poc_cover_data3 <- Poc_cover_data2 %>% 
  dplyr::summarise(total_count_biotics = sum(SpeciesCounts)) %>% 
  filter(CowTagID!="VSEEP")

Poc_cover_data4 <- Poc_cover_data2 %>% 
   filter(Taxa=="Pocillopora acuta")
 
Poc_cover_data5 <- Poc_cover_data4 %>% 
  left_join(Poc_cover_data3) %>% 
  group_by(CowTagID, Location, SpeciesCounts,total_count_biotics ) %>% 
  dplyr::summarise(percent_cover = (SpeciesCounts/total_count_biotics)*(100))

##########################################
### THIS REMOVES SOME OF THE TILES BC NOT ALL TILES HAD P. ACUTA COUNTS 
##########################################
EnvData_wPocCover <- EnvData_wrecruitcounts %>% 
  full_join(Poc_cover_data5) %>% 
  write_csv(here::here("Data", "EnvData_wPocCover.csv"))

##########################################
## using Poc data WITH hard substrate and sand etc 
###########################################
Poc_cover_data_withhardsub <- Poc_cover_data %>% 
  select(!c("Date", "PhotoNum", "Notes")) %>% 
  filter(Location!="Varari_Maya") %>% 
  group_by(CowTagID) 
  
  

```

## Look at recruit totals by how many P acuta at the CowTagID

```{r}

recruits_by_PocCover <- EnvData_wPocCover %>% 
  ggplot(aes(x=percent_cover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              fill="cornsilk3", 
              alpha=0.2) + ## alpha adjusts the transparency 
  labs(x="% Cover of P. acuta Species found near tile location", 
       y="Logged Total of Recruits")

recruits_by_PocCover
ggsave(here::here("Outputs", "RecruitmentTiles", "recruits_by_PocCover.jpg"))

recruits_by_PocCover_facet <- EnvData_wPocCover %>% 
  ggplot(aes(x=percent_cover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              fill="cornsilk3", 
              alpha=0.2) + ## alpha adjusts the transparency 
  facet_wrap(~Location, scales="free") + 
  labs(x="% Cover of P. acuta Species found near tile location", 
       y="Logged Total of Recruits")

recruits_by_PocCover_facet
ggsave(here::here("Outputs", "RecruitmentTiles", "Results", "Prelim", "recruits_by_PocCover_facet.jpg"))

```

### create a similiar plot for recruits by just hard substrate 
Currently getting an error (June 1st, 2024) because don't have substrate type in the EnvData_wPocCover df anymore 
Should probably use PercentCover df and group into substrate -- might have code somehwere, currently tired 
```{r}
EnvData_wPocCover_long <- EnvData_wPocCover %>% 
  pivot_longer(cols=Sand:HardSubstrate, names_to = "SubstrateType", values_to = "SubstrateValues")

recruits_by_HardSubstrate <- EnvData_wPocCover_long %>% 
  ggplot(aes(x=SubstrateValues, 
             y=log(sum_total+1), 
             color=SubstrateType)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              fill="cornsilk3", 
              alpha=0.2) + 
  labs(x="Percent Cover of Hard Substrate vs Sand", 
       y="Logged Total of Recruits")
recruits_by_HardSubstrate

ggsave(here("Outputs", "RecruitmentTiles", "recruits_by_SubstrateType.jpg"))


```

### what about just live coral cover? include Porites, etc 

```{r}
CommComp_recruit_coralcover <- PercentCover3 %>% 
  left_join(RecruitmentCounts2.1, join_by(CowTagID, Location))

CommComp_recruit_coralcover_liveonly <- CommComp_recruit_coralcover %>% 
  filter(comm_type=="LiveCoral", 
         CowTagID!="CSEEP", CowTagID!="VSEEP")

CommComp_recruit_seasonavg_envdata_wLiveCover <- CommComp_recruit_coralcover_liveonly %>% 
  left_join(AllChemDataSeasonAvg_calculatedvalues, join_by(CowTagID)) 

recruits_by_AllLiveCoral <- CommComp_recruit_seasonavg_envdata_wLiveCover %>% 
  ggplot(aes(x=percent_cover, 
             y=log(sum_total+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_classic() + 
  scale_color_manual(values=c("cornflowerblue", "salmon")) +
  geom_smooth(method=lm, 
              fill="cornsilk3", 
              alpha=0.2) + 
  facet_wrap(~Location) + 
  labs(x="Percent Live Coral", 
       y= "Log of Total Recruits")

recruits_by_AllLiveCoral

ggsave(here::here("Outputs", "RecruitmentTiles", "recruits_by_AllLiveCoral.jpg"))



```

#### run models comparing total live coral cover vs Poc cover at Varari 
```{r}

######################
## all live coral including both sites 
######################
model_livecoral <- lm(log(sum_total+1) ~ percent_cover + Location, data= CommComp_recruit_seasonavg_envdata_wLiveCover) 
summary(model_livecoral)

######################
## just Varari 
######################

CommComp_recruit_coralcover_liveonly_V <- CommComp_recruit_seasonavg_envdata_wLiveCover %>% 
  filter(Location=="Varari")

model_Vlivecoral <- lm(log(sum_total+1) ~ percent_cover, data= CommComp_recruit_coralcover_liveonly_V) 
summary(model_Vlivecoral)

######################
## just Cabral 
######################

CommComp_recruit_coralcover_liveonly_C <- CommComp_recruit_coralcover_liveonly %>% 
  filter(Location=="Cabral")

model_Clivecoral <- lm(log(sum_total+1) ~ percent_cover, data= CommComp_recruit_coralcover_liveonly_C) 
summary(model_Clivecoral)

###################################################################
######################
## just Poc cover 
######################
EnvData_wPocCover_V <- EnvData_wPocCover %>% 
  filter(Location=="Varari")

model_VPocCover <- lm(log(sum_total+1) ~ SpeciesCounts, data= EnvData_wPocCover_V) 
summary(model_VPocCover)


EnvData_wPocCover_C <- EnvData_wPocCover %>% 
  filter(Location=="Cabral")

model_CPocCover <- lm(log(sum_total+1) ~ SpeciesCounts, data= EnvData_wPocCover_C) 
summary(model_CPocCover)

```


### ^^ Results: 
Varari: 
- Poc cover: 0.044
- Live coral: 0.0184
Cabral: 
- Poc cover: 0.54
- Live coral: 0.44




