---
title: "RecruitmentTiles_PrelimAnalysis"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

## Load libraries 
```{r}
library(segmented)
library(plotrix)
library(gridExtra)
library(LoLinR)
library(lubridate)
library(chron)
library(patchwork)
library(tidyverse)
library(here)
library(PNWColors)
library(ggrepel)
library(reshape2)
library(viridis)
library(car)
library(GGally)
library(corrplot)
library(PNWColors)
library(seacarb)
library(broom)
library(calecopal)
library(ggridges)
library(agricolae)
library(lme4)
library(lmerTest)
library(modelsummary)
library(tidymodels)
library(flextable)
library(performance)
library(agricolae)
library(ggeffects)
library(sjPlot)
library(patchwork)
library(broom)
library(purrr)
library(nlstools)
library(stringr)
library(emmeans)
library(MuMIn)
library(forcats)
library(ggmap)
library(maptools)
library(kriging)
library(ggnewscale)
library(wql)
library(pscl)
```

## Bring in datasets
```{r}
RecruitmentCounts <- read_csv(here("Data","Recruitment_Tiles_Counts.csv"))
AllChemData <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/August2021/Allbiogeochemdata_QC.csv")
PercentCover <- read_csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Full_Metadata.csv")
pH <- read_csv(here("Data", "pHSlope.csv"))
```

## join dfs 
```{r}

full_CommunityRecruit <- PercentCover %>% 
  left_join(RecruitmentCounts, join_by(CowTagID, Location))

full_envdataRecruits <- full_CommunityRecruit %>% 
  left_join(AllChemData, join_by(Location, CowTagID, lat, lon))
```

# tidy this now fully complete and joined dataset 
```{r}
full_envdataRecruits <- full_envdataRecruits %>% 
  select(!c("del15N", "adj_CT_depth_cm", "meanRugosity", "Top_Plate_ID", "Bottom_Plate_ID", "Jamie_Plate_ID", "Time", "DateTime", "M_C":"Lignin_Like"))

#full_envdataRecruits <- full_envdataRecruits %>% 
  #select()
# should I get rid of extra rows from seeps and rock wall, etc ??? 

```

# group and average the day/night and high/low tide measurements 
```{r}

full_envdataRecruits2 <- full_envdataRecruits %>% 
  select(!c("Date", "Tide", "Day_Night", "Plate_Seep")) %>% 
  group_by(CowTagID) %>% 
  mutate(avgSal=mean(Salinity, na.rm=TRUE), 
         avgTemp=mean(Temperature, na.rm=TRUE), 
         avgTA=mean(TA, na.rm=TRUE), 
         avgpH=mean(pH, na.rm=TRUE), 
         avgPhosphate=mean(Phosphate_umolL, na.rm=TRUE), 
         avgSilicate=mean(Silicate_umolL, na.rm=TRUE),
         avgNN=mean(NN_umolL, na.rm=TRUE), 
         avgAmmonia=mean(Ammonia_umolL, na.rm=TRUE))

### not sure if I did this correctly??? maybe I could make this tidier? lots of extra rows due to date, tide, and day/night 
  
```

## tidy the df a bit more to start visualizing with plots 
### mainly pivoting the recruitment counts 
```{r}
## pivot dataset and group together 
Recruits_pivoted <- RecruitmentCounts %>% 
  group_by(CowTagID, Location, SGD_level, Side) %>% 
  pivot_longer(cols="Pocilloporidae":"Other", names_to = "recruit_counts") %>% 
  summarize(sum_total=sum(value)) 

### getting incorrect values if you FULL, joined dataframe because too many rows per CowTag bc of other data -- fix this -- for right now, using just the recruitment count df 

totalcounts_tilelocation <- Recruits_pivoted %>% 
  pivot_wider(names_from = Side, values_from = sum_total) %>% 
  mutate(total_count = Bottom + Side + Top) ##R recognizes double quotes as characters, so just add the literal column name 


```

## these plots show rough SGD estimate, taken from silicate kriging plot with total recruit counts per plate 

```{r}
## try to plot basics for Varari 
### this plot 
Recruits_by_SGD_level_V <- totalcounts_tilelocation %>% 
  filter(Location=="Varari") %>% 
  ggplot(aes(x=reorder(CowTagID, -total_count), 
             y=total_count, 
             fill=SGD_level)) + 
  geom_col() + 
  theme_blank()  + 
  scale_fill_manual(values=c("darkseagreen4", "darkgoldenrod2", "coral4"), 
                    limits=c("Low", "Mid", "High")) + 
   labs(title="Varari", 
       x="CowTag_ID", 
       y="Settler Counts")
  
Recruits_by_SGD_level_V

## for Cabral 
Recruits_by_SGD_level_C <- totalcounts_tilelocation %>% 
  filter(Location=="Cabral") %>% 
  ggplot(aes(x=reorder(CowTagID, -total_count), 
             y=total_count, 
             fill=SGD_level)) + 
  geom_col() + 
  theme_blank() + 
  scale_fill_manual(values=c("darkseagreen4", "darkgoldenrod2", "coral4"), 
                    limits=c("Low", "Mid", "High")) + 
  labs(title="Cabral", 
       x="CowTag_ID", 
       y="Settler Counts")
  
Recruits_by_SGD_level_C


Recruits_by_SGD_level_C + Recruits_by_SGD_level_V
ggsave(here("Outputs", "RecruitmentTiles", "RecruitbySGDestimates.jpg"), 
       width = 14, height = 10)


### try with SE bars  ?? 
totalcounts_tilelocation2 <- totalcounts_tilelocation %>% 
  group_by(SGD_level, CowTagID, Location, total_count) %>% 
  summarize(mean_recruit = mean(total_count, na.rm=TRUE), 
            se_recruit = sd(total_count, na.rm=TRUE)/sqrt(n())) %>% 
  ungroup() 
 # geom_errorbar(aes(ymin=mean_settled-se_settled, ymax=mean_settled+se_settled), width=0.1) 


################################
### associated models 
#################################

## trying to run ANOVA to see if there is any significance between treatments and between sites  
Varari_recruitment <- totalcounts_tilelocation %>% 
   filter(Site=="Varari")

anova_recruits_Varari <- aov(total_count ~ SGD_level, data=Varari_recruitment)
summary(anova_recruits_Varari)
emmeans(anova_recruits_Varari, pairwise~"SGD_level", adjust="Tukey")

Cabral_recruitment <- totalcounts_tilelocation %>% 
   filter(Site=="Cabral")

anova_recruits_Cabral <- lm(total_count ~ SGD_level, data=Cabral_recruitment) 
summary(anova_recruits_Cabral)
emmeans(anova_recruits_Cabral, pairwise~"SGD_level", adjust="Tukey")

anova_recruits <- aov(total_count ~ SGD_level + Site, data=totalcounts_tilelocation)
summary(anova_recruits)
emmeans(anova_recruits, pairwise~"SGD_level", adjust="Tukey")
```


## want to look at tile data for each of the env parameters (April 17th, 2024)
## start with percent cover data 
### try with a regular barlplot and then stacked barplot because part out of whole 
```{r}

full_envdataRecruits3 <- full_envdataRecruits2 %>% 
  pivot_longer(cols = LiveCoral:Sand, names_to = "comm_type", values_to = "percent_cover") %>% 
  group_by(comm_type, CowTagID, Location) %>% 
  summarize(percent_total=sum(percent_cover)) 
  
full_envdataRecruits4 <- full_envdataRecruits3 %>% 
  pivot_wider(names_from = CowTagID, values_from = percent_total) 

full_envdataRecruits4 <- full_envdataRecruits4 %>% 
  select(!starts_with(c("Reef","RockWall","Sand"))) %>% ## gets rid of the extra data that I don't need - columns that start with Reef, Rockwall, or Sand 
  pivot_longer(cols = C1:VSEEP, names_to = "CowTagID", values_to = "percent_cover") %>% 
  drop_na(percent_cover)

#### now plot 
commcomp_plot_V <- full_envdataRecruits4 %>% 
  filter(Location=="Varari") %>% 
  ggplot(aes(x=factor(CowTagID, level = c("V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14", "V15", "V16", "V17", "V18", "V19", "V20", "VSEEP")),
             y=percent_cover,
             fill=comm_type)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_blank() + 
  scale_fill_manual(values=c("darkolivegreen", "coral", "azure4", "burlywood3"),
                    limits=c("DeadCoral", "LiveCoral", "Rubble", "Sand")) + 
  labs(title="Varari", 
       x="CowTag_ID", 
       y="% cover of Comm Types") + 
  theme(axis.text.x = element_text(angle = 45))
  
  
commcomp_plot_V

## for Cabral 
commcomp_plot_C <- full_envdataRecruits4 %>% 
  filter(Location=="Cabral") %>% 
  ggplot(aes(x=factor(CowTagID, level = c("C1", "C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10", "C11", "C12", "C13", "C14", "C15", "C16", "C17", "C18", "C19", "C20", "CSEEP")),
             y=percent_cover,
             fill=comm_type)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_blank() + 
  scale_fill_manual(values=c("darkolivegreen", "coral", "azure4", "burlywood3"),
                    limits=c("DeadCoral", "LiveCoral", "Rubble", "Sand")) + 
  labs(title="Cabral", 
       x="CowTag_ID", 
       y="% cover of Comm Types") + 
  theme(axis.text.x = element_text(angle = 45))
  
commcomp_plot_C




commcomp_plot_C + commcomp_plot_V


```

## ^^ is way too high for total percentage so trying to fix that here (April 18th, 2024)
```{r}
PercentCover <- PercentCover %>% 
  select(!c("del15N":"N_percent", "adj_CT_depth_cm", "meanRugosity"))

PercentCover2 <- PercentCover %>% 
  pivot_longer(cols = LiveCoral:Sand, names_to = "comm_type", values_to = "percent_cover") %>% 
  group_by(comm_type, CowTagID, Location) %>% 
  summarize(percent_total=sum(percent_cover)) 

PercentCover3 <- PercentCover2 %>% 
  pivot_wider(names_from = CowTagID, values_from = percent_total) 

PercentCover3 <- PercentCover3 %>% 
  select(!starts_with(c("Reef","RockWall","Sand"))) %>% ## gets rid of the extra data that I don't need - columns that start with Reef, Rockwall, or Sand 
  pivot_longer(cols = C1:VSEEP, names_to = "CowTagID", values_to = "percent_cover") %>% 
  drop_na(percent_cover)


```
 
# was successful - now plot normally (April 18th, 2024)

```{r}
PercentCover_plot_V <- PercentCover3 %>% 
  filter(Location=="Varari") %>% 
  ggplot(aes(x=factor(CowTagID, level = c("V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14", "V15", "V16", "V17", "V18", "V19", "V20", "VSEEP")),
             y=percent_cover,
             fill=comm_type)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_blank() + 
  scale_fill_manual(values=c("darkolivegreen", "coral", "azure4", "burlywood3"),
                    limits=c("DeadCoral", "LiveCoral", "Rubble", "Sand")) + 
  labs(title="Varari", 
       x="CowTag_ID", 
       y="% cover of Comm Types") + 
  theme(axis.text.x = element_text(angle = 45))

PercentCover_plot_V

## for Cabral 
PercentCover_plot_C <- PercentCover3 %>% 
  filter(Location=="Cabral") %>% 
  ggplot(aes(x=factor(CowTagID, level = c("C1", "C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10", "C11", "C12", "C13", "C14", "C15", "C16", "C17", "C18", "C19", "C20", "CSEEP")),
             y=percent_cover,
             fill=comm_type)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_blank() + 
  scale_fill_manual(values=c("darkolivegreen", "coral", "azure4", "burlywood3"),
                    limits=c("DeadCoral", "LiveCoral", "Rubble", "Sand")) + 
  labs(title="Cabral", 
       x="CowTag_ID", 
       y="% cover of Comm Types") + 
  theme(axis.text.x = element_text(angle = 45))
  
PercentCover_plot_C


PercentCover_plot_C + PercentCover_plot_V
ggsave(here("Outputs", "RecruitmentTiles", "PercentCommunityCover.jpg"), 
       width = 14, height = 10)

```

## create a scatterplot to show comm comp data vs total number of recruits on tiles at each of these sites 

```{r}

## need to summarize recruit totals properly and then join with percent cover data and plot 
RecruitmentCounts2 <- RecruitmentCounts %>% 
  pivot_longer(cols = Pocilloporidae:Other, names_to = "coral_spp", values_to = "recruit_counts")

RecruitmentCounts3 <- RecruitmentCounts2 %>% 
  group_by(CowTagID, Location, SGD_level) %>% 
  summarize(recruit_total= sum(recruit_counts))

CommComp_recruit <- PercentCover3 %>% 
  left_join(RecruitmentCounts3, join_by(CowTagID, Location))


CommComp_recruit_plot <- CommComp_recruit %>%
  ggplot(aes(x=recruit_total, 
             y=percent_cover, 
             fill=factor(comm_type))) +
  geom_point(size=4, shape=21, color="black") +
  scale_fill_manual(values=c("darkolivegreen", "coral", "azure4", "burlywood3"),
                    limits=c("DeadCoral", "LiveCoral", "Rubble", "Sand")) + 
  facet_wrap(~Location) + 
  theme_bw() + 
  labs(x="recruit counts", 
       y="percent cover of community types", 
       fill="community types")
  

CommComp_recruit_plot
ggsave(here("Outputs", "RecruitmentTiles", "RecruitCounts_by_PercentCover.jpg"), 
       width = 18, height = 10)

###############################
## also facet by each CowTag so can isolate specific correlations if there are any  
################################

CommComp_recruit_plot2V <- CommComp_recruit %>%
  filter(Location=="Varari") %>% 
  ggplot(aes(x=recruit_total, 
             y=percent_cover, 
             fill=factor(comm_type))) +
  geom_point(size=4, shape=21, color="black") +
  scale_fill_manual(values=c("darkolivegreen", "coral", "azure4", "burlywood3"),
                    limits=c("DeadCoral", "LiveCoral", "Rubble", "Sand")) + 
  facet_wrap(~CowTagID) + 
  theme_bw() + 
  labs(x="recruit counts", 
       y="percent cover of community types", 
       fill="community types")
  

CommComp_recruit_plot2V

###############################
CommComp_recruit_plot2C <- CommComp_recruit %>%
  filter(Location=="Cabral") %>% 
  ggplot(aes(x=recruit_total, 
             y=percent_cover, 
             fill=factor(comm_type))) +
  geom_point(size=4, shape=21, color="black") +
  scale_fill_manual(values=c("darkolivegreen", "coral", "azure4", "burlywood3"),
                    limits=c("DeadCoral", "LiveCoral", "Rubble", "Sand")) + 
  facet_wrap(~CowTagID) + 
  theme_bw() + 
  labs(x="recruit counts", 
       y="percent cover of community types", 
       fill="community types")
  

CommComp_recruit_plot2C

```


## create a plot to show just total recruit counts on the y axis by CowTag 
```{r}
## for Cabral 
recruit_plot_C <- RecruitmentCounts3 %>% 
  filter(Location=="Cabral") %>% 
  ggplot(aes(x=factor(CowTagID, level = c("C1", "C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10", "C11", "C12", "C13", "C14", "C15", "C16", "C17", "C18", "C19", "C20")), 
             y=recruit_total, 
             fill=CowTagID)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_blank() + 
  labs(title="Cabral", 
       x="CowTag_ID", 
       y="Total Recruits per Plate") + 
  scale_fill_manual(values=rev(pnw_palette("Starfish", n=20)), 
                     guide="none") +
  theme(axis.text.x = element_text(angle = 45))

recruit_plot_C

## for Varari 
recruit_plot_V <- RecruitmentCounts3 %>% 
  filter(Location=="Varari") %>% 
  ggplot(aes(x=factor(CowTagID, level = c("V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14", "V15", "V16", "V17", "V18", "V19", "V20")), 
             y=recruit_total, 
             fill=CowTagID)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_blank() + 
  labs(title="Varari", 
       x="CowTag_ID", 
       y="Total Recruits per Plate") + 
  scale_fill_manual(values=rev(pnw_palette("Starfish", n=20)), 
                     guide="none") +
  theme(axis.text.x = element_text(angle = 45))

recruit_plot_V

recruit_plot_V + recruit_plot_C 
ggsave(here("Outputs", "RecruitmentTiles", "RecruitCounts.jpg"), 
       width = 18, height = 10)


```


## create a new plot where recruit counts are colored by estimated SGD 
```{r}
recruit_plot_C2 <- RecruitmentCounts3 %>% 
  filter(Location=="Cabral") %>% 
  ggplot(aes(x=factor(CowTagID, level = c("C1", "C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10", "C11", "C12", "C13", "C14", "C15", "C16", "C17", "C18", "C19", "C20")), 
             y=recruit_total, 
             fill=SGD_level)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_blank() + 
  scale_fill_manual(values=c("darkseagreen4", "darkgoldenrod2", "coral4"), 
                    limits=c("Low", "Mid", "High")) +
  labs(title="Cabral", 
       x="CowTag_ID", 
       y="Total Recruits per Plate") 
#  theme(axis.text.x = element_text(angle = 45))

recruit_plot_C2

## for Varari 
recruit_plot_V2 <- RecruitmentCounts3 %>% 
  filter(Location=="Varari") %>% 
  ggplot(aes(x=factor(CowTagID, level = c("V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14", "V15", "V16", "V17", "V18", "V19", "V20")), 
             y=recruit_total, 
             fill=SGD_level)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_blank() + 
  labs(title="Varari", 
       x="CowTag_ID", 
       y="Total Recruits per Plate") + 
  scale_fill_manual(values=c("darkseagreen4", "darkgoldenrod2", "coral4"), 
                    limits=c("Low", "Mid", "High"))
 # theme(axis.text.x = element_text(angle = 45))

recruit_plot_V2

recruit_plot_V2 + recruit_plot_C2
ggsave(here("Outputs", "RecruitmentTiles", "RecruitCounts_bySGDest.jpg"), 
       width = 18, height = 10)

```



## now try to make a plot that incorporates env data with recruit counts
```{r}
## first join necessary dfs

AllChemData_edit <- AllChemData %>% 
  select(!c("Top_Plate_ID":"Plate_Seep", "M_C":"Lignin_Like")) 
  
AllChemData_edit_avg <- AllChemData_edit %>% 
  group_by(CowTagID) %>% 
  summarize(sal_avg = mean(Salinity), 
            temp_avg = mean(Temperature), 
            TA_avg = mean(TA), 
            pH_avg = mean(pH), 
            phos_avg = mean(Phosphate_umolL), 
            NN_avg = mean(NN_umolL), 
            silicate_avg= mean(Silicate_umolL), 
            ammonia_avg = mean(Ammonia_umolL))


CommComp_recruit_envdata <- CommComp_recruit %>% 
  left_join(AllChemData_edit_avg, join_by(CowTagID))

CommComp_recruit_envdata_long <- CommComp_recruit_envdata %>% 
  pivot_longer(cols="sal_avg":"ammonia_avg", names_to = "env_params")

######## now plot 
##### REMOVE SEEP DATA #####
CommComp_recruit_envdata_plot_C <- CommComp_recruit_envdata_long %>%
  filter(Location=="Cabral") %>% 
   drop_na(recruit_total) %>% 
  ggplot(aes(x=value, 
             y=log(recruit_total+1))) +
  geom_point(size=3, shape=21) + 
  facet_wrap(~env_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts")

CommComp_recruit_envdata_plot_C

## for Varari 
CommComp_recruit_envdata_plot_V <- CommComp_recruit_envdata_long %>%
  #filter(Location=="Varari") %>% 
   drop_na(recruit_total) %>% 
  ggplot(aes(x=value, 
             y=log(recruit_total+1), 
             color=Location)) +
  geom_point(size=3, shape=21) + 
  facet_wrap(Location~env_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts")

CommComp_recruit_envdata_plot_V


## env params just silicate 
CommComp_recruit_envdata_plot_V2 <- CommComp_recruit_envdata_long %>%
  filter(env_params=="silicate_avg") %>% 
  filter(Location=="Varari") %>% 
   drop_na(recruit_total) %>% 
  ggplot(aes(x=value, 
             y=log(recruit_total+1), 
             color=Location)) +
  geom_point(size=3, shape=21) + 
  geom_smooth(method=lm, formula="y~poly(x,2)") + 
  facet_wrap(~env_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts")
CommComp_recruit_envdata_plot_V2

## run a model 
test_silicatemodelV <- anova(lm(data=CommComp_recruit_envdata_long %>%
  filter(env_params=="silicate_avg") %>% 
  filter(Location=="Varari") %>% 
    drop_na(recruit_total, value),
  log(recruit_total+1)~poly(value,2)))

check_model(test_silicatemodelV) 



### trying a bar plot 
CommComp_recruit_envdata_plot_C2 <- CommComp_recruit_envdata_long %>%
  filter(Location=="Cabral") %>% 
  group_by(env_params) %>% 
  ggplot(aes(x=env_params, 
             y=recruit_total, 
             fill=CowTagID)) +
  geom_bar(position="stack", stat="identity") +
 # scale_fill_viridis_b() + 
  theme_bw() + 
  labs(x="CowTag", 
       y="recruit counts")

CommComp_recruit_envdata_plot_C2



```













## try to make a PCA for the env data? 

```{r}

# Before we can use this data file, R wants ONLY trait data in order to do the PCA. 
# (just to get PCA scores).
AllChemData_edit_avg_PCA <- AllChemData_edit_avg %>% 
  select(!"CowTagID") %>% 
  drop_na()

# First, to get all of our traits on the same scale, we'll convert them to z-scores. You don't have to do this, but it is fairly common, and it can be helpful down the line:
envdata.scale <- scale(AllChemData_edit_avg_PCA, scale=TRUE, center=TRUE)

# Run the PCA with `princomp()`
PCAmodel <- princomp(envdata.scale, cor=FALSE)

# In princomp, the default is to use the covariance matrix. If you want to use the correlation matrix, then use cor=TRUE Here we converted to z-scores first, so all variables are on the same scale and we can use the covariancematrix. (The correlation matrix is the scaled covariance matrix, recall.)

summary(PCAmodel) 

# This shows amount of variance explained by each axis. It's actually a big wide
# table, with all 38 principle components. It tells us the amount of variance 
# explained by each axis, and the cumulative proportion of variance explained
# Axis 1 will always explain the most variation, Axis 2 the second most, etc.

# We can plot that data easily in a scree plot
plot(PCAmodel, type="lines")

# Or even prettier:
library(factoextra)
fviz_eig(PCAmodel)

PCAmodel$loadings  # Shows the loadings of each trait on each PC 
# --- make sure you scroll up, this is a long readout. 

PCAmodel$scores # Gives output of the principal components for each individual 
# on each PC --- another 

# If we want to use the PC scores for something else (maybe we want to use 
# PC1 to run an ANOVA?), then we can pull out that vector.

PC1 <- PCAmodel$scores[,1] # Pulls out the first column of the table 
PC1

# And we could run a quick ANOVA to see if guilds differ in PC1 scores. 
# (This works because the ordering of rows in the PC scores table is the same 
# as the order in the original data table.)
AllChemData_edit_avg <- AllChemData_edit_avg %>% 
  drop_na()

model1 <- lm(PC1~AllChemData_edit_avg$CowTagID)
anova(model1)


# Another common thing to do with PCA is to make a biplot
biplot(PCAmodel, xlab="PC1", ylab="PC2")
# The black numbers are each individual spider. The red lines are the vectors 
# for each trait.

# There are tons of options for playing around with the biplot. I'll just do 
# this to make this look a bit better so we can see differences among guilds

# I like using ggbiplot
library(devtools)
install_github("vqv/ggbiplot") #You only need to download this package once

# In this biplot code, I'm going to separate the spiders by guild (so I need to
# use my non-subsetted data file). So I use the original file to refer to 
# guilds, but the PCA data for everything else. I'm also going to put confidence 
# intervals around the individuals in each guild.

library(ggbiplot)

ggbiplot(PCAmodel, groups=AllChemData_edit_avg$CowTagID, ellipse=TRUE, circle=FALSE) +
  scale_color_discrete(name='') +
  geom_point(aes(colour=factor(AllChemData_edit_avg$CowTagID)), size = 1) +
  theme(legend.direction = 'horizontal', legend.position='bottom', legend.text=element_text(size=8))



```

















