---
title: "SettlementExp_PrelimAnalysis"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

### Load libraries 
```{r}
library(patchwork)
library(tidyverse)
library(here)
library(PNWColors)
library(viridis)
library(car)
library(lubridate)
library(lme4)
library(lmerTest)
library(modelsummary)
library(tidymodels)
library(flextable)
library(performance)
library(agricolae)
library(broom)
library(stringr)
library(emmeans)

```

### rework with silicate data - August 13th, 2024 
-  Here I fixed join issue by manually adding in water chem data to excel/csv and re-loading as new df called "SettlementCount_updated" 
```{r}
## just bring in Settlements with MANUALLY ADDED nutrient data 

SettlementCount_updated <- read_csv(here::here("Data", "SettlementCounts_edited.csv"))


  SettlementCount_updated <- SettlementCount_updated %>% 
  filter(settled!=0) #%>% ##dropped one outlier in data where there were 0 counts for everything, so I think no larvae were actually added
  #write_csv(here::here("Data", "SettlementCounts_edited.csv"))
```


### Bring in datasets
-  checked to make sure that nutrient data is correctly assigned to each treatment (August 23rd, 2024) for settlement exp 
-  so can use the SettlementCounts_updated df 
```{r}
SettlementCounts <- read_csv(here::here("Data","2024_SettlementCounts.csv"))
## raw settlement counts from the experiment 

pH <- read_csv(here::here("Data", "pHSlope.csv"))
## calcualted pH from the temp and mV data within each of the treatment bins 

WaterChem<-read_csv(here::here("Data", "2024_WaterChemMeasurements.csv"))
## raw water chem measurements from each of the treatment bins throughout the experiment 

#SettlementCounts_updated <- read_csv(here::here("Data","SettlementCounts_edited.csv"))
```

### tidy the data and start visualizing 
-  tidied pH and water chem dfs and then joined them together 
-  then joined with settlement counts 
-  as of May 12th --> not properly joined --> joined by hand in August and created new script so no worries 

-  also created SettlementCounts2 df to include counts of total settled vs dead 
-  can ignore this ^ and comment out bc included in SettlementCounts_updated df 
```{r}
pH <- pH %>% 
  mutate(TrisCalDate = ymd(TrisCalDate))

WaterChem <- WaterChem %>% 
  mutate(TrisCalDate = ymd(TrisCalDate), 
         date=mdy(date)) 

waterchemdata <- pH %>% 
  full_join(WaterChem, join_by(treatmentID_rep, experiment_number, sampling_period, TrisCalDate, time, date, orion_pH, mV, temp, salinity, light, note, TA, NN, phosphate, ammonia, silicate)) 

#SettlementCounts <- SettlementCounts %>% 
#  mutate(date = mdy(date))

#fulldata <- waterchemdata %>% 
 # full_join(SettlementCounts, join_by(treatmentID_rep, treatment_level, date)) #### not working perfecetly as of May 12th 2024 ---- not all water chem data is showing up 

###########################
## tidy data 
###########################

#SettlementCounts2 <- SettlementCounts %>% 
#  mutate(dead=20-(swimming+plastic_glue+mesh+top+bottom+side)) %>% 
 # mutate(settled=(plastic_glue+mesh+top+bottom+side)) %>% 
#  mutate(treatment_level=as.factor(treatment_level), 
    #     spawning_day=as.factor(spawning_day)) %>% 
#  mutate(dead = ifelse(dead < 0, 0, dead)) %>% 
#  mutate(settled = ifelse(settled > 20, 20, settled)) %>% 
#  mutate(countcheck = (dead+settled)) #%>% 
 # write_csv(here::here("Data", "SettlementCounts_edited.csv"))

```

### quick visualizations - updated May 12th, 2024
-  preliminary look at settlement counts throughout the experiment and treatments 
-  not seeing too much of a difference 
-  group by swimming vs settled, settled per day 
```{r}

###############################################
#### quick visualizations to plot patterns of settlement ####
###############################################
# first need to tidy and pivot the df to make count data and sampleID 

settlement_counts_sum <- SettlementCounts %>%
  pivot_longer(cols = swimming:side, names_to = "settlement_surface" , values_to = "count")
##list which columns you want, colon includes "a" through "z", if just want one or two can use "|" 
#to pivot, then rename the new column and the values that you are trying to get as the output

# group by treatment type and then summarize counts from there 
#settlement_counts_sum <- settlement_counts_sum %>% 
# group_by(treatment_level) %>% 
#  summarize(sum_total=sum(count)) 

##calculating standard deviation 
#settlement_counts_sum <- settlement_counts_sum %>% 
# sd(count)

## separated by site 
settlement_plot <- settlement_counts_sum %>% 
  drop_na(site) %>% 
  ggplot(aes(x=treatment_level,  
             y=count,
             fill=settlement_surface)) + 
  geom_col() + 
  facet_wrap(~site) + 
  theme_bw() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High")) 

settlement_plot
ggsave(here("Outputs", "SettlementExperiment", "broadsettlement.jpg"), 
       width=10, height=8)
#geom_errorbar(aes(x=treatment_level, ymin=count-sd, ymax=count+sd), width=0.4)

## separated by day and site to see if there are any differences across day 
settlement_plot2 <- settlement_counts_sum %>% 
  drop_na(spawning_day) %>% 
  ggplot(aes(x=treatment_level,  
             y=count,
             fill=settlement_surface)) + 
  geom_col() + 
  facet_wrap(site~spawning_day) + 
  theme_bw() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High"))
settlement_plot2
ggsave(here("Outputs", "SettlementExperiment", "settlement_byday.jpg"), 
       width=10, height=8)

# group by swimming vs any settled 
settlement_counts_mutated <- SettlementCounts %>% 
  mutate(settled = plastic_glue+mesh+top+bottom+side) %>%
  pivot_longer(cols = swimming|settled, names_to = "swim_noswim" , values_to = "count")

settlement_plot3 <- settlement_counts_mutated %>% 
  drop_na(site) %>% 
  ggplot(aes(x=treatment_level,  
             y=count,
             fill=swim_noswim)) + 
  geom_col() + 
  facet_wrap(~site) + 
  theme_bw() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High"))

settlement_plot3
ggsave(here("Outputs", "SettlementExperiment", "settlement_SwimnoSwim.jpg"), 
       width=10, height=8)

settlement_plot4 <- settlement_counts_mutated %>% 
  drop_na(spawning_day, site) %>% 
  ggplot(aes(x=treatment_level,  
             y=count,
             fill=swim_noswim)) + 
  geom_col() + 
  facet_wrap(site~spawning_day) + 
  theme_bw() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High"))

settlement_plot4
ggsave(here("Outputs", "SettlementExperiment", "settlement_SwimnoSwim_perday.jpg"), 
       width=10, height=8)

```


### More plots and some models? data exploration 
-  Ran ANOVAs to see if there is a difference in settlement across day (did not find signficance)
-  created df SettlementCounts_SEMean to include mean and standard errors for settled vs dead, then made bar plot 
```{r}
#### try making an ANOVA ####

SettlementCounts1 <- SettlementCount_updated %>% 
  filter(spawning_day==2)

SettlementCounts2 <- SettlementCount_updated %>% 
  filter(spawning_day==3)

anova_settlement <- aov(settled ~ treatment_level, data=SettlementCount_updated)
summary(anova_settlement) 

### if do lmer, use anova rather than summary 

## try to isolate difference between just ambient and low or med and high treatments for day 3 


### not sure what is happening here 
plot <- SettlementCount_updated %>%
  ggplot(aes(x=treatment_level, y=settled, color=treatmentID_rep)) +
  geom_point() + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High"))  

SettlementCount_SEMean <- SettlementCount_updated %>% 
  drop_na(site) %>% 
  group_by(spawning_day, treatment_level, site) %>% 
  summarize(mean_settled = mean(settled, na.rm=TRUE), 
            mean_dead = mean(dead, na.rm=TRUE), 
            mean_swimming = mean(swimming, na.rm=TRUE), 
            se_settled = sd(settled, na.rm=TRUE)/sqrt(n()), 
            se_dead = sd(dead, na.rm=TRUE)/sqrt(n()), 
            se_swimming = sd(swimming, na.rm=TRUE)/sqrt(n()))

SettlementCounts_avgs <- SettlementCount_SEMean %>% 
  ggplot(aes(x=spawning_day, 
             y=mean_settled, 
             color=treatment_level)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=mean_settled-se_settled, ymax=mean_settled+se_settled), width=0.1) + 
  scale_color_discrete(limits=c("Ambient","Low","Medium", "High"))  + 
  facet_wrap(site~treatment_level)

SettlementCounts_avgs

ggsave(here("Outputs", "SettlementExperiment", "avg_counts_by_site_and_day.jpg"), 
      width=10, height=8)



ambientonly_meansettlement <- SettlementCount_SEMean %>% 
  filter(treatment_level== "Ambient") %>% 
  ggplot(aes(x=spawning_day, y=mean_settled)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=mean_settled-se_settled, ymax=mean_settled+se_settled), width=0.1)  + 
  facet_wrap(~site)

ambientonly_meansettlement
ggsave(here("Outputs", "SettlementExperiment", "ambientonly_meansettlement.jpg"), 
       width=10, height=8)
# scale_color_discrete(limits=c("Ambient","Low","Medium", "High"))
#facet_wrap(~treatment_level)

## errors and percent averages across replicates 
## box plot of total settlers 
settlement_counts_mutated <- SettlementCount_updated %>% 
  mutate(settled = plastic_glue+mesh+top+bottom+side) %>%
  mutate(dead=20-(swimming+settled)) %>% 
  pivot_longer(cols = swimming|dead|settled, names_to = "swim_settled_dead" , values_to = "count")

settlement_counts_mutated1 <- settlement_counts_mutated %>%
  drop_na(spawning_day) %>% 
  ggplot(aes(x=treatment_level, 
             y=count, 
             fill=swim_settled_dead)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  ggtitle("Settlers vs Swimmers vs Presumed Dead") + 
  facet_wrap(spawning_day~site, scales ="free_y") + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High")) 

settlement_counts_mutated1
ggsave(here("Outputs", "SettlementExperiment", "boxplots.jpg"), 
       width=10, height=8)

# just the settlers 
settlement_counts_mutated2 <- SettlementCount_updated %>% 
  mutate(settled = plastic_glue+mesh+top+bottom+side) %>% 
  pivot_longer(cols = settled, names_to = "settled" , values_to = "count") %>% 
  drop_na(site)

settlement_counts_mutated_plot2 <- settlement_counts_mutated2 %>%
  ggplot(aes(x=treatment_level, y=count, fill=settled)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  ggtitle("Settlers") + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High")) + 
  facet_wrap(~site)

settlement_counts_mutated_plot2
ggsave(here("Outputs", "SettlementExperiment", "boxplots2.jpg"), 
       width=10, height=8)

settlement_by_day <- settlement_counts_mutated2 %>%
  ggplot(aes(x=treatment_level, 
             y=count, 
             fill=settled)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  facet_wrap(site~spawning_day) +
  ggtitle("Settlers") + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High")) 
settlement_by_day

ggsave(here("Outputs", "SettlementExperiment", "boxplots3_byday.jpg"), 
       width=10, height=8)


```

###############################################################
###############################################################

# AUGUST 2024 - THESIS REVISIT for Chapter TWO 


### For time being, forget about treatment blocks and days and combine all results together for settlement experiment 
- Goal: to test silicate on the x axis and _expected proportions_ of settlers on the y 

```{r}
###################################
### first separate out by sites to add in total possible settlers column 
###################################

SettlementCounts_editV <- SettlementCount_updated %>% 
  filter(site=="Varari") %>% 
  dplyr::select(-c("notes", "swimming":"spawning_day", "rep_number", "treatmentID_rep", "countcheck")) %>% 
  group_by(treatment_level, site) %>% 
  dplyr::summarize(settled2 = sum(settled)) %>% 
  dplyr::mutate(totalpossible = 20 * 11) %>% 
  dplyr::mutate(proportionsettled = settled2 / totalpossible)

SettlementCount_editC <- SettlementCount_updated %>% 
  filter(site=="Cabral") %>% 
  dplyr::select(-c("notes", "swimming":"spawning_day", "rep_number", "treatmentID_rep", "countcheck")) %>% 
  group_by(treatment_level, site) %>%
  dplyr::summarize(settled2 = sum(settled),
                   totalpossible=20*30,
                   proportionsettled=settled2/totalpossible, 
                   avgproportion = mean(proportionsettled),
                   seproportion = sd(proportionsettled) / sqrt(n()))

###################################
### rejoin new dfs 
###################################

SettlementCounts_joined <- SettlementCount_editC %>% 
  full_join(SettlementCounts_editV)
  

###################################
### plot
###################################

SettlementCounts_plotted <- SettlementCounts_joined %>% 
  ggplot(aes(x=treatment_level, 
             y=proportionsettled, 
             fill=site)) +
  geom_col(position = position_dodge()) + ##how to get fill variable next to each other rather than stacked 
  scale_x_discrete(limits = c("Ambient", "Low", "Medium", "High")) + 
  scale_fill_manual(values=c("burlywood3", "coral4"), 
                    limits=c("Cabral", "Varari")) + 
  theme_classic() + 
  theme(axis.text.x=element_text(size=25), 
        axis.text.y=element_text(size=25), 
        axis.title.x=element_text(size=25), 
        axis.title.y=element_text(size=25))
  
SettlementCounts_plotted
ggsave(here("Outputs", "SettlementExperiment", "Prelim", "SettlementCounts_proportion1.jpg"), 
       width=10, height=8)

SettlementCounts_plotted2 <- SettlementCounts_joined %>% 
  ggplot(aes(x=treatment_level, 
             y=proportionsettled, 
             fill=site)) +
  geom_col() + 
  scale_x_discrete(limits = c("Ambient", "Low", "Medium", "High")) + 
  scale_fill_manual(values=c("burlywood3", "coral4"), 
                    limits=c("Cabral", "Varari")) + 
  facet_wrap(~site) +
  theme_classic() + 
  theme(axis.text.x=element_text(size=25), 
        axis.text.y=element_text(size=25), 
        axis.title.x=element_text(size=25), 
        axis.title.y=element_text(size=25))
  
SettlementCounts_plotted2
ggsave(here("Outputs", "SettlementExperiment", "Prelim", "SettlementCounts_proportion2.jpg"), 
       width=10, height=8)


###################################
### Run a chi-squared test to see if those are real differences
###################################

# Create a contingency table of counts for each combination of treatment_level and site
SettlementCounts_joined_V <- SettlementCounts_joined %>% 
  filter(site=="Varari")

contingency_table_V <- table(SettlementCounts_joined_V$treatment_level, SettlementCounts_joined_V$proportionsettled)

# Perform the Chi-Square test
chi_sq_test <- chisq.test(contingency_table_V)

chi_sq_test


#####################################
## what would above plot look like as the average proportion with error bars instead of the total proportion across everything?
#####################################

### edit dfs 

#SettlementCounts_editV2 <- SettlementCounts %>% 
 # filter(site=="Varari") %>% 
 # dplyr::select(-c("notes", "swimming":"side", "countcheck")) %>% 
#  group_by(treatment_level, site) %>% 
 # dplyr::summarize(mean_settled = mean(settled)) %>% 
#  dplyr::mutate(total_possible = if_else(spawning_day == 2 & treatment_level == "High", 20 * 6, 
 #                                        if_else(spawning_day == 1, 20 * 6, 20 * 5))) %>% 
 # dplyr::mutate(mean_proportionsettled = mean_settled / total_possible) %>% 
 # ungroup() %>%
 # dplyr::mutate(se = sd(mean_proportionsettled) / sqrt(n()))
##################

###################################
### need avg proportion per treatment  
### calculate the proportion BEFORE grouping so that have individual prop for each data point, and THEN average  
###################################

SettlementCounts_bothsites <- SettlementCount_updated %>% 
  dplyr::select(-c("notes", "countcheck")) %>% 
  dplyr::mutate(proportion_settled = settled/20) %>% 
  group_by(treatment_level, site) %>%
  dplyr::summarize(mean_proportionsettled=mean(proportion_settled), 
                se = sd(proportion_settled) / sqrt(n())) %>% 
  ungroup() 
  
###################################
### now plot with SE bars 
###################################
SettlementCounts_plotted3 <- SettlementCounts_bothsites %>% 
  ggplot(aes(x=treatment_level, 
             y=mean_proportionsettled, 
             fill=site)) +
  geom_col(position = position_dodge(), width = 0.7) + 
  geom_errorbar(aes(ymin = mean_proportionsettled - se, 
                    ymax = mean_proportionsettled + se), 
                 width = 0.2, 
                position = position_dodge(0.7)) + 
  scale_x_discrete(limits = c("Ambient", "Low", "Medium", "High")) + 
  scale_fill_manual(values=c("burlywood3", "coral4"), 
                    limits=c("Cabral", "Varari")) + 
  facet_wrap(~site) +
  theme_classic() + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18), 
        axis.title.y=element_text(size=18))

SettlementCounts_plotted3
ggsave(here("Outputs", "SettlementExperiment", "Prelim", "SettlementCounts_avgproportion.jpg"), 
       width=10, height=8)

###################################
### plot with silicate
###################################

SettlementCounts_bothsites2 <- SettlementCount_updated %>% 
  dplyr::select(-c("notes", "swimming":"side", "countcheck")) %>% 
  dplyr::mutate(proportion_settled = settled/20) %>% 
  group_by(treatment_level, site) %>%
  dplyr::summarize(mean_silicate = mean(silicate), 
                   mean_proportionsettled=mean(proportion_settled), 
                   se = sd(proportion_settled) / sqrt(n())) %>% 
  ungroup() 
  

SettlementCounts_plotted4 <- SettlementCounts_bothsites2 %>% 
  ggplot(aes(x=mean_silicate, 
             y=mean_proportionsettled, 
             color=treatment_level)) +
  geom_point(size=3) + 
  scale_fill_manual(values=c("burlywood3", "coral4"), 
                    limits=c("Cabral", "Varari")) + 
  facet_wrap(~site, scales="free_y") +
  theme_bw() + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18), 
        axis.title.y=element_text(size=18))

SettlementCounts_plotted4

```

### Plots using silicate as the main x-axis 
-  different versions of plots using silicate data on the x and settled counts on the y 
-  silicate_settlement2 plot shows faceted by day 
-  silicate_settlement4 plot shows faceted by day but with AVERAGED silicate across treatments (treating individual bins as if they were identical, grouping those reps together)
-  also check the raw data and made sure that there were no copies of data at Varari (August 23rd, 2024)
```{r}
###########
## by treatment level 
silicate_settlement1 <- SettlementCount_updated %>% 
  filter(site=="Varari") %>% 
  ggplot(aes(x=silicate, 
             y=settled, 
             color=treatment_level,
             shape = as.factor(rep_number), 
             size=as.factor(spawning_day))) + 
  geom_point(alpha=0.5)  + 
  facet_wrap(~site, scales="free_x") + 
  theme_bw()
  
silicate_settlement1


##########
## by spawning day  
silicate_settlement2 <- SettlementCount_updated %>% 
  ggplot(aes(x=silicate, 
             y=settled, 
             color=treatment_level, 
             shape = as.factor(rep_number))) + 
  geom_point(size=3)  + 
  facet_wrap(spawning_day~site, scales="free_x") + 
  theme_bw()
  

silicate_settlement2

##########
## general 
silicate_settlement3 <- SettlementCount_updated %>% 
  ggplot(aes(x=silicate, 
             y=settled, 
             color=treatment_level)) + 
  geom_point(size=2)  + 
  scale_color_manual(values=c("deepskyblue", "darkseagreen", "darkgoldenrod3", "coral3"), 
                    limits=c("Ambient", "Low", "Medium", "High")) + 
  facet_wrap(~site, scales="free_x") + 
  theme_bw() 


silicate_settlement3
ggsave(here("Outputs", "SettlementExperiment", "Prelim", "silicate_settled.jpg"), 
       width=10, height=8)

########################################
## take avg silciate per treatment 
########################################

SettlementCount_silicateavg <- SettlementCount_updated %>% 
  group_by(treatment_level) %>% 
  mutate(avg_silicate= mean(silicate))

silicate_settlement4 <- SettlementCount_silicateavg %>% 
  ggplot(aes(x=avg_silicate, 
             y=settled, 
             color=treatment_level)) + 
  geom_point(size=2)  + 
  scale_color_manual(values=c("deepskyblue", "darkseagreen", "darkgoldenrod3", "coral3"), 
                    limits=c("Ambient", "Low", "Medium", "High")) + 
  facet_wrap(~site, scales="free_x") + 
  theme_bw() 

silicate_settlement4

```

### Create plots for other nutrients as well (NN and P)
- replaced silicate with NN and P 

```{r}
nitrate_settlement <- SettlementCount_updated %>% 
  ggplot(aes(x=NN, 
             y=settled, 
             color=treatment_level)) + 
  geom_point(size=2)  + 
  facet_wrap(~site, scales="free_x") +
  scale_color_manual(values=c("deepskyblue", "darkseagreen", "darkgoldenrod3", "coral3"), 
                    limits=c("Ambient", "Low", "Medium", "High")) + 
  guides(color = "none") +  
  theme_bw()
  
nitrate_settlement

ggsave(here("Outputs", "SettlementExperiment", "Prelim", "NN_settled.jpg"), 
       width=10, height=8)


phosphate_settlement <- SettlementCount_updated %>% 
  ggplot(aes(x=phosphate, 
             y=settled, 
             color=treatment_level)) + 
  geom_point(size=2)  + 
  facet_wrap(~site, scales="free_x") +
  scale_color_manual(values=c("deepskyblue", "darkseagreen", "darkgoldenrod3", "coral3"), 
                    limits=c("Ambient", "Low", "Medium", "High")) + 
  guides(color = "none") +  
  theme_bw()
  
phosphate_settlement

ggsave(here("Outputs", "SettlementExperiment", "Prelim", "PO4_settled.jpg"), 
       width=10, height=8)

########################################
## patch all 3 env param plots together 
########################################

phosphate_settlement / nitrate_settlement / silicate_settlement3
ggsave(here("Outputs", "SettlementExperiment", "Prelim", "EnvParamsPlot_vs_settled.jpg"), 
       width=10, height=15)

```

### plot day on x and settled on y, color by treatment 
-  some of the points are hidden behind others (like the Medium counts for Day 1 at Cabral for example)

```{r}
perday_settlement <- SettlementCount_updated %>% 
  ggplot(aes(x=as.factor(spawning_day), 
             y=settled, 
             color=treatment_level)) + 
  geom_point(size=2, alpha=0.5)  + 
  facet_wrap(~site, scales="free_x") +
  scale_color_manual(values=c("deepskyblue", "darkseagreen", "darkgoldenrod3", "coral3"), 
                    limits=c("Ambient", "Low", "Medium", "High")) + 
  #guides(color = "none") +  
  theme_bw()
  
perday_settlement


```

### plot day on x and settled on y, color by treatment AND facet by treatment 
-  analysis: ambient day 1 and 6 are more variable than 2-4 BUT opposite in Medium at Cabral, High and low kinda have the same variance across 
-  Varari has lower settlement in day 2 compared to day 1 in ambient and low treatments but not necessarily in medium and high 
-  
```{r}
######### 
## facet by treatment 

perday_settlement_treatmentfacet <- SettlementCount_updated %>% 
  ggplot(aes(x=as.factor(spawning_day), 
             y=settled, 
             color=treatment_level)) + 
  geom_point(size=2)  + 
  facet_wrap(treatment_level~site, scales="free_x") +
  scale_color_manual(values=c("deepskyblue", "darkseagreen", "darkgoldenrod3", "coral3"), 
                    limits=c("Ambient", "Low", "Medium", "High")) + 
  #guides(color = "none") +  
  theme_bw()
  
perday_settlement_treatmentfacet
```

### plot silicate on x, settled 





### Plot based on proportion settled by treatment 
- Notes from Craig during meeting (Aug 7th) - focus on Cabral first since 6 days vs only 2 days at Varari 

```{r}
SettlementCounts_Cabral <- SettlementCount_updated %>% 
  filter(site=="Cabral") %>% 
  dplyr::select(-c("notes", "rep_number", "treatmentID_rep", "countcheck")) %>% 
  group_by(treatment_level, site, spawning_day) %>%
  dplyr::summarize(settled2 = sum(settled),
                   totalpossible=20*5,
                   proportionsettled=settled2/totalpossible)

## model = 
Cabralsettlement_model <- lmer(proportionsettled~treatment_level + (1|spawning_day), data =SettlementCounts_Cabral)

summary(Cabralsettlement_model)


## significant difference from ambient ?? 

Cabralsettlementplot <- SettlementCounts_Cabral %>% 
  ggplot(aes(x=treatment_level, 
             y=proportionsettled, 
             fill=as.factor(spawning_day))) + 
  geom_col(position = position_dodge()) + 
  scale_x_discrete(limits = c("Ambient", "Low", "Medium", "High")) + 
  scale_fill_manual(values=pnw_palette("Bay", n=6)) 


Cabralsettlementplot
```










#############################################################
#############################################################
#### EXTRA: 
- not using as not well coded but saved for back-up 

```{r}

#############
## Varari 

SettlementCount_avgproportions_V<- SettlementCount_updated %>% 
  filter(site=="Varari") %>% 
  dplyr::select(-c("notes", "countcheck")) %>% 
  group_by(treatment_level, site) %>% 
  dplyr::summarize(total_settled = sum(settled), 
                   mean_settled = mean(settled), 
                   totalpossible_perbin = 20) %>% 
  dplyr::mutate(total_possible= if_else(treatment_level == "High", 20 * 12, 20 * 11),
                proportion_settled = total_settled/total_possible, 
                mean_proportionsettled=mean_settled/totalpossible_perbin) %>% #, 
               # se = sd(settled / total_possible) / sqrt(n())) %>% 
  ungroup() 

#############
## Cabral  

low_treatment_count <- SettlementCount_updated %>% 
  filter(site=="Cabral") %>% 
  filter(treatment_level == "Medium") %>% 
  nrow()
low_treatment_count

SettlementCounts_editC2 <- SettlementCount_updated %>% 
  filter(site=="Cabral") %>% 
  dplyr::select(-c("notes", "countcheck")) %>% 
  dplyr::mutate(proportion_settled = settled/20) 


 # group_by(treatment_level, site) %>% 
 # dplyr::mutate(se = sd(mean_proportionsettled) / sqrt(n()))
  
  
  SettlementCount_avgproportions2 <- SettlementCount_updated %>% 
  group_by(treatment_level, site) %>% 
  dplyr::summarize(mean_settled = mean(settled), 
                   mean_silicate = mean(silicate)) %>% 
  dplyr::mutate(
    total_possible = if_else(site == "Varari" & treatment_level == "High", 20 * 6, 20 * 5),
    mean_proportionsettled = mean_settled / total_possible,
    se = sd(mean_settled / total_possible) / sqrt(n())
  ) %>%
  ungroup()
```


### Notes for what to work on (March 2024)
```{r}

# add error bars 
# load env data 
# include random effect of tank and maybe table?
# make plots for salinity and count data 
# figure out Tris calibrations to get pH calculations 
# make plots for pH and count data 
# add days 4 and 5 and hopefully 6 to see if there is a stronger pattern 
# think about hourly effect of day 1-3 and also day of larval release 
# plug in hypothetical, approximate nutrient values from respo data to see if pattern with nutrients 

```

