---
title: "SettlementExp_PrelimAnalysis"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

## Load libraries 
```{r}
library(patchwork)
library(tidyverse)
library(here)
library(PNWColors)
library(viridis)
library(car)
library(lubridate)
library(lme4)
library(lmerTest)
library(modelsummary)
library(tidymodels)
library(flextable)
library(performance)
library(agricolae)
library(broom)
library(stringr)
library(emmeans)

```

## Bring in datasets
```{r}
SettlementCounts <- read_csv(here::here("Data","2024_SettlementCounts.csv"))
pH <- read_csv(here::here("Data", "pHSlope.csv"))
WaterChem<-read_csv(here::here("Data", "2024_WaterChemMeasurements.csv"))
```

## tidy the data and start visualizing 
```{r}
pH <- pH %>% 
  mutate(TrisCalDate = ymd(TrisCalDate))

WaterChem <- WaterChem %>% 
  mutate(TrisCalDate = ymd(TrisCalDate), 
         date=mdy(date)) 

waterchemdata <- pH %>% 
  full_join(WaterChem, join_by(treatmentID_rep, experiment_number, sampling_period, TrisCalDate, time, date, orion_pH, mV, temp, salinity, light, note, TA, NN, phosphate, ammonia, silicate)) 

SettlementCounts <- SettlementCounts %>% 
  mutate(date = mdy(date))

fulldata <- waterchemdata %>% 
  full_join(SettlementCounts, join_by(treatmentID_rep, treatment_level, date)) #### not working perfecetly as of May 12th 2024 ---- not all water chem data is showing up 

###########################
## tidy data 
###########################

SettlementCounts <- SettlementCounts %>% 
  mutate(dead=20-(swimming+plastic_glue+mesh+top+bottom+side)) %>% 
  mutate(settled=(plastic_glue+mesh+top+bottom+side)) %>% 
  mutate(treatment_level=as.factor(treatment_level), 
         spawning_day=as.factor(spawning_day)) %>% 
  mutate(dead = ifelse(dead < 0, 0, dead)) %>% 
  mutate(settled = ifelse(settled > 20, 20, settled)) %>% 
  mutate(countcheck = (dead+settled)) 

```

#### quick visualizations - updated May 12th, 2024
```{r}

###############################################
#### quick visualizations to plot patterns of settlement ####
###############################################
# first need to tidy and pivot the df to make count data and sampleID 

settlement_counts_sum <- SettlementCounts %>%
  pivot_longer(cols = swimming:side, names_to = "settlement_surface" , values_to = "count")
##list which columns you want, colon includes "a" through "z", if just want one or two can use "|" 
#to pivot, then rename the new column and the values that you are trying to get as the output

# group by treatment type and then summarize counts from there 
#settlement_counts_sum <- settlement_counts_sum %>% 
# group_by(treatment_level) %>% 
#  summarize(sum_total=sum(count)) 

##calculating standard deviation 
#settlement_counts_sum <- settlement_counts_sum %>% 
# sd(count)

## separated by site 
settlement_plot <- settlement_counts_sum %>% 
  drop_na(site) %>% 
  ggplot(aes(x=treatment_level,  
             y=count,
             fill=settlement_surface)) + 
  geom_col() + 
  facet_wrap(~site) + 
  theme_bw() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High")) 

settlement_plot
ggsave(here("Outputs", "SettlementExperiment", "broadsettlement.jpg"), 
       width=10, height=8)
#geom_errorbar(aes(x=treatment_level, ymin=count-sd, ymax=count+sd), width=0.4)

## separated by day and site to see if there are any differences across day 
settlement_plot2 <- settlement_counts_sum %>% 
  drop_na(spawning_day) %>% 
  ggplot(aes(x=treatment_level,  
             y=count,
             fill=settlement_surface)) + 
  geom_col() + 
  facet_wrap(site~spawning_day) + 
  theme_bw() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High"))
settlement_plot2
ggsave(here("Outputs", "SettlementExperiment", "settlement_byday.jpg"), 
       width=10, height=8)

# group by swimming vs any settled 
settlement_counts_mutated <- SettlementCounts %>% 
  mutate(settled = plastic_glue+mesh+top+bottom+side) %>%
  pivot_longer(cols = swimming|settled, names_to = "swim_noswim" , values_to = "count")

settlement_plot3 <- settlement_counts_mutated %>% 
  drop_na(site) %>% 
  ggplot(aes(x=treatment_level,  
             y=count,
             fill=swim_noswim)) + 
  geom_col() + 
  facet_wrap(~site) + 
  theme_bw() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High"))

settlement_plot3
ggsave(here("Outputs", "SettlementExperiment", "settlement_SwimnoSwim.jpg"), 
       width=10, height=8)

settlement_plot4 <- settlement_counts_mutated %>% 
  drop_na(spawning_day, site) %>% 
  ggplot(aes(x=treatment_level,  
             y=count,
             fill=swim_noswim)) + 
  geom_col() + 
  facet_wrap(site~spawning_day) + 
  theme_bw() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High"))

settlement_plot4
ggsave(here("Outputs", "SettlementExperiment", "settlement_SwimnoSwim_perday.jpg"), 
       width=10, height=8)

```




## THINGS TO DO BEFORE SHOWING PETE AND NYSSA --- March 2024
```{r}


# add error bars 
# load env data 
# include random effect of tank and maybe table?
# make plots for salinity and count data 
# figure out Tris calibrations to get pH calculations 
# make plots for pH and count data 
# add days 4 and 5 and hopefully 6 to see if there is a stronger pattern 
# think about hourly effect of day 1-3 and also day of larval release 
# plug in hypothetical, approximate nutrient values from respo data to see if pattern with nutrients 

```

## More plots and some models? 
```{r}
#### try making an ANOVA ####

SettlementCounts1 <- SettlementCounts %>% 
  filter(spawning_day==2)

SettlementCounts2 <- SettlementCounts %>% 
  filter(spawning_day==3)

anova_settlement <- aov(settled ~ treatment_level, data=SettlementCounts)
summary(anova_settlement) 

### if do lmer, use anova rather than summary 

## try to isolate difference between just ambient and low or med and high treatments for day 3 


### not sure what is happening here 
#plot <- SettlementCounts %>%
 # ggplot(aes(x=treatment_level, y=settled, color=treatmentID_rep)) +
 # geom_point() + 
 # scale_x_discrete(limits=c("Ambient","Low","Medium", "High"))  

SettlementCounts_SEMean <- SettlementCounts %>% 
  drop_na(site) %>% 
  group_by(spawning_day, treatment_level, site) %>% 
  summarize(mean_settled = mean(settled, na.rm=TRUE), 
            mean_dead = mean(dead, na.rm=TRUE), 
            mean_swimming = mean(swimming, na.rm=TRUE), 
            se_settled = sd(settled, na.rm=TRUE)/sqrt(n()), 
            se_dead = sd(dead, na.rm=TRUE)/sqrt(n()), 
            se_swimming = sd(swimming, na.rm=TRUE)/sqrt(n()))

SettlementCounts_avgs <- SettlementCounts_SEMean %>% 
  ggplot(aes(x=spawning_day, 
             y=mean_settled, 
             color=treatment_level)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=mean_settled-se_settled, ymax=mean_settled+se_settled), width=0.1) + 
  scale_color_discrete(limits=c("Ambient","Low","Medium", "High"))  + 
  facet_wrap(site~treatment_level)

SettlementCounts_avgs

ggsave(here("Outputs", "SettlementExperiment", "avg_counts_by_site_and_day.jpg"), 
       width=10, height=8)



ambientonly_meansettlement < - SettlementCounts_SEMean %>% 
  filter(treatment_level== "Ambient") %>% 
  ggplot(aes(x=spawning_day, y=mean_settled)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=mean_settled-se_settled, ymax=mean_settled+se_settled), width=0.1)  + 
  facet_wrap(~site)

ambientonly_meansettlement
ggsave(here("Outputs", "SettlementExperiment", "ambientonly_meansettlement.jpg"), 
       width=10, height=8)
# scale_color_discrete(limits=c("Ambient","Low","Medium", "High"))
#facet_wrap(~treatment_level)

## errors and percent averages across replicates 
## box plot of total settlers 
settlement_counts_mutated <- SettlementCounts %>% 
  mutate(settled = plastic_glue+mesh+top+bottom+side) %>%
  mutate(dead=20-(swimming+settled)) %>% 
  pivot_longer(cols = swimming|dead|settled, names_to = "swim_settled_dead" , values_to = "count")

settlement_counts_mutated1 <- settlement_counts_mutated %>%
  drop_na(spawning_day) %>% 
  ggplot(aes(x=treatment_level, 
             y=count, 
             fill=swim_settled_dead)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  ggtitle("Settlers vs Swimmers vs Presumed Dead") + 
  facet_wrap(spawning_day~site, scales ="free_y") + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High")) 

settlement_counts_mutated1
ggsave(here("Outputs", "SettlementExperiment", "boxplots.jpg"), 
       width=10, height=8)

# just the settlers 
settlement_counts_mutated2 <- SettlementCounts %>% 
  mutate(settled = plastic_glue+mesh+top+bottom+side) %>% 
  pivot_longer(cols = settled, names_to = "settled" , values_to = "count") %>% 
  drop_na(site)

settlement_counts_mutated_plot2 <- settlement_counts_mutated2 %>%
  ggplot(aes(x=treatment_level, y=count, fill=settled)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  ggtitle("Settlers") + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High")) + 
  facet_wrap(~site)

settlement_counts_mutated_plot2
ggsave(here("Outputs", "SettlementExperiment", "boxplots2.jpg"), 
       width=10, height=8)

settlement_by_day <- settlement_counts_mutated2 %>%
  ggplot(aes(x=treatment_level, 
             y=count, 
             fill=settled)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  facet_wrap(site~spawning_day) +
  ggtitle("Settlers") + 
  scale_x_discrete(limits=c("Ambient","Low","Medium", "High")) 
settlement_by_day

ggsave(here("Outputs", "SettlementExperiment", "boxplots3_byday.jpg"), 
       width=10, height=8)


```

### For time being, forget about blocks and days and combine all results together for setttlement experiment 
- to test silicate on the x axis and expected proportions of settlers on the y 

```{r}
###################################
### first separate out by sites to add in total possible settlers column 
###################################

SettlementCounts_editV <- SettlementCounts %>% 
  filter(site=="Varari") %>% 
  dplyr::select(-c("notes", "swimming":"spawning_day", "rep_number", "treatmentID_rep", "countcheck")) %>% 
  group_by(treatment_level, site) %>% 
  dplyr::summarize(settled2 = sum(settled)) %>% 
  dplyr::mutate(totalpossible = if_else(treatment_level == "High", 20 * 12, 20 * 11)) %>% 
  dplyr::mutate(proportionsettled = settled2 / totalpossible)

SettlementCounts_editC <- SettlementCounts %>% 
  filter(site=="Cabral") %>% 
  dplyr::select(-c("notes", "swimming":"spawning_day", "rep_number", "treatmentID_rep", "countcheck")) %>% 
  group_by(treatment_level, site) %>%
  dplyr::summarize(settled2 = sum(settled),
                   totalpossible=20*30,
                   proportionsettled=settled2/totalpossible)

###################################
### rejoin new dfs 
###################################

SettlementCounts_joined <- SettlementCounts_editC %>% 
  full_join(SettlementCounts_editV)
  

###################################
### plot
###################################

SettlementCounts_plotted <- SettlementCounts_joined %>% 
  ggplot(aes(x=treatment_level, 
             y=proportionsettled, 
             fill=site)) +
  geom_col(position = position_dodge()) + ##how to get fill variable next to each other rather than stacked 
  scale_x_discrete(limits = c("Ambient", "Low", "Medium", "High")) + 
  scale_fill_manual(values=c("burlywood3", "coral4"), 
                    limits=c("Cabral", "Varari")) + 
  theme_classic() + 
  theme(axis.text.x=element_text(size=25), 
        axis.text.y=element_text(size=25), 
        axis.title.x=element_text(size=25), 
        axis.title.y=element_text(size=25))
  
SettlementCounts_plotted
ggsave(here("Outputs", "SettlementExperiment", "Prelim", "SettlementCounts_proportion1.jpg"), 
       width=10, height=8)

SettlementCounts_plotted2 <- SettlementCounts_joined %>% 
  ggplot(aes(x=treatment_level, 
             y=proportionsettled, 
             fill=site)) +
  geom_col() + 
  scale_x_discrete(limits = c("Ambient", "Low", "Medium", "High")) + 
  scale_fill_manual(values=c("burlywood3", "coral4"), 
                    limits=c("Cabral", "Varari")) + 
  facet_wrap(~site) +
  theme_classic() + 
  theme(axis.text.x=element_text(size=25), 
        axis.text.y=element_text(size=25), 
        axis.title.x=element_text(size=25), 
        axis.title.y=element_text(size=25))
  
SettlementCounts_plotted2
ggsave(here("Outputs", "SettlementExperiment", "Prelim", "SettlementCounts_proportion2.jpg"), 
       width=10, height=8)


###################################
### Run a chi-squared test to see if those are real differences
###################################

# Create a contingency table of counts for each combination of treatment_level and site
SettlementCounts_joined_V <- SettlementCounts_joined %>% 
  filter(site=="Varari")

contingency_table_V <- table(SettlementCounts_joined_V$treatment_level, SettlementCounts_joined_V$proportionsettled)

# Perform the Chi-Square test
chi_sq_test <- chisq.test(contingency_table_V)

chi_sq_test


#####################################
## above plot look like as the average proportion with error bars instead of the total proportion across everything?
#####################################

### edit dfs 

SettlementCounts_editV2 <- SettlementCounts %>% 
  filter(site=="Varari") %>% 
  dplyr::select(-c("notes", "swimming":"side", "countcheck")) %>% 
  group_by(treatment_level, site, spawning_day) %>% 
  dplyr::summarize(mean_settled = mean(settled)) %>% 
  dplyr::mutate(total_possible = if_else(spawning_day == 2 & treatment_level == "High", 20 * 6, 
                                         if_else(spawning_day == 1, 20 * 6, 20 * 5))) %>% 
  dplyr::mutate(mean_proportionsettled = mean_settled / total_possible) %>% 
  ungroup() %>%
  dplyr::mutate(se = sd(mean_proportionsettled) / sqrt(n()))


SettlementCounts_editC2 <- SettlementCounts %>% 
  filter(site=="Cabral") %>% 
  dplyr::select(-c("notes", "swimming":"side", "countcheck")) %>% 
  group_by(treatment_level, site, spawning_day) %>%
  dplyr::summarize(mean_settled = mean(settled),
                   total_possible=20*5,
                   mean_proportionsettled=mean_settled/total_possible) %>% 
  ungroup() %>%
  group_by(treatment_level, site) %>% 
  dplyr::mutate(se = sd(mean_proportionsettled) / sqrt(n()))

###################################
### rejoin new dfs 
###################################

SettlementCounts_joined2 <- SettlementCounts_editC2 %>% 
  full_join(SettlementCounts_editV2)

###################################
### now plot with SE bars 
###################################
SettlementCounts_plotted3 <- SettlementCounts_joined2 %>% 
  ggplot(aes(x=treatment_level, 
             y=mean_proportionsettled, 
             fill=site)) +
  geom_col(position = position_dodge(), width = 0.7) + 
  geom_errorbar(aes(ymin = mean_proportionsettled - se, 
                    ymax = mean_proportionsettled + se), 
                 width = 0.2, 
                position = position_dodge(0.7)) + 
  scale_x_discrete(limits = c("Ambient", "Low", "Medium", "High")) + 
  scale_fill_manual(values=c("burlywood3", "coral4"), 
                    limits=c("Cabral", "Varari")) + 
  facet_wrap(~site) +
  theme_classic() + 
  theme(axis.text.x=element_text(size=25), 
        axis.text.y=element_text(size=25), 
        axis.title.x=element_text(size=25), 
        axis.title.y=element_text(size=25))

SettlementCounts_plotted3

```

### rework with silicate data - August 7th, 2024 
```{r}
WaterChem <- WaterChem %>% 
  mutate(TrisCalDate = ymd(TrisCalDate), 
         date=mdy(date)) 

waterchemdata <- pH %>% 
  full_join(WaterChem, join_by(treatmentID_rep, experiment_number, sampling_period, TrisCalDate, time, date, orion_pH, mV, temp, salinity, light, note, TA, NN, phosphate, ammonia, silicate)) 

SettlementCounts <- SettlementCounts %>% 
  mutate(date = mdy(date))

fulldata <- waterchemdata %>% 
  full_join(SettlementCounts, join_by(treatmentID_rep, treatment_level, date)) #### not working perfecetly as of May 12th 2024 ---- not all water chem data is showing up 


WaterChem_edit <- WaterChem %>% 
  
  

```




#### Craig meeting - focus on Cabral days first 

```{r}
SettlementCounts_Cabral <- SettlementCounts %>% 
  filter(site=="Cabral") %>% 
  dplyr::select(-c("notes", "swimming":"side", "rep_number", "treatmentID_rep", "countcheck")) %>% 
  group_by(treatment_level, site, spawning_day) %>%
  dplyr::summarize(settled2 = sum(settled),
                   totalpossible=20*5,
                   proportionsettled=settled2/totalpossible)



## model = 
Cabralsettlement_model <- lmer(proportionsettled~treatment_level + (1|spawning_day), data =SettlementCounts_Cabral)

summary(Cabralsettlement_model)


## significant difference from ambient ?? 



Cabralsettlementplot <- SettlementCounts_Cabral %>% 
  ggplot(aes(x=treatment_level, 
             y=proportionsettled, 
             fill=spawning_day)) + 
  geom_col(position = position_dodge()) + 
  scale_x_discrete(limits = c("Ambient", "Low", "Medium", "High")) + 
  scale_fill_manual(values=pnw_palette("Bay", n=6)) 


Cabralsettlementplot
```



