---
title: "Model selection on linear and nonlinear regressions of recruits on plates and SGD parameters"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

# LOAD LIBRARIES
```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)

```

### Load and tidy data 
```{r}
CommComp_recruit_seasonavg_envdata <- read_csv(here("Data", "CommComp_recruit_seasonavg_envdata.csv"))

CommComp_recruit_seasonavg_allenvdata <- CommComp_recruit_seasonavg_envdata %>% 
  select(!c("Sand", "HardSubstrate", "sum_total_dead", "sum_total_alive")) %>%  
  drop_na(sum_total)

CommComp_recruit_seasonavg_allenvdata_C <- CommComp_recruit_seasonavg_allenvdata %>% 
  filter(Location=="Cabral") %>% 
  mutate(sum_total=log(sum_total+1)) %>% 
  drop_na(sal_avg, phos_avg)

CommComp_recruit_seasonavg_allenvdata_V <- CommComp_recruit_seasonavg_allenvdata %>% 
  filter(Location=="Varari") %>% 
  mutate(sum_total=log(sum_total+1)) %>% 
  drop_na()


CommComp_recruit_seasonavg_longpredictorvars_C <- CommComp_recruit_seasonavg_allenvdata %>% 
  pivot_longer(cols = sal_avg:ammonia_CV, values_to = "Predictor_Values") %>% 
  mutate(Predictor_Variable=name, 
         Response_Variable=log(sum_total+1)) %>% 
  select(!c(name, sum_total)) %>% 
  filter(Location=="Cabral") %>% 
  drop_na()

CommComp_recruit_seasonavg_longpredictorvars_V <- CommComp_recruit_seasonavg_allenvdata %>% 
  pivot_longer(cols = sal_avg:ammonia_CV, values_to = "Predictor_Values") %>% 
  mutate(Predictor_Variable=name, 
         Response_Variable=log(sum_total+1)) %>% 
  select(!c(name, sum_total)) %>% 
  filter(Location=="Varari") %>% 
  drop_na()

```



### Danielle's script 
```{r}
modTable <- tibble(Y = as.character(),
                     Parameter = as.character(),
                     Reg_Type = as.character(), # linear or nonlinear
                     AICc = as.numeric(),
                     R2 = as.numeric(),
                     pVal.P = as.numeric(),
                     #pVal.Rug = as.numeric()
                   )

myDep <- colnames(resFric %>% select(NbSpP, NbFEsP, Vol8D,
                                     resSpp, resFEp, resVol))
mydata <- resFric %>%
  select(CowTagID,NbSpP, NbFEsP, Vol8D,
         resSpp, resFEp, resVol,
         Salinity, Temperature, pH, Phosphate_umolL,
         Silicate_umolL, NN_umolL, meanRugosity)


for(i in myDep){
  Y <- as.character(i)
  k <- which(myDep == i) # use as multiplier for list

  for(j in 8:ncol(mydata)){
    Parameter <- colnames(mydata[j])

    # with rugosity as covariate
    model1R <- lm(paste0(Y, "~", Parameter, ""), data = mydata)
    subdata1R <- as_tibble(cbind(Y,Parameter)) %>%
      mutate(Parameter = paste0(Parameter,"")) %>%
      mutate(Reg_Type = "Linear",
             AICc = AICc(model1R),
             R2 = summary(model1R)$r.squared,
             pVal.P = summary(model1R)$coefficients[8]
             )

    # with rugosity as covariate
    model2R <- lm(paste0(Y, "~ poly(", Parameter, ",2)"), data = mydata)
    subdata2R <- as_tibble(cbind(Y,Parameter)) %>%
      mutate(Parameter = paste0(Parameter,"")) %>%
      mutate(Reg_Type = "Polynomial",
             AICc = AICc(model2R),
             R2 = summary(model2R)$r.squared,
             pVal.P = summary(model2R)$coefficients[12]
             )


    modTable <- modTable %>%
      rbind(subdata1R) %>%
      rbind(subdata2R)

  }
}


modelTable <- modTable %>%
  group_by(Y) %>%
  mutate(minAIC = min(AICc)) %>%
  mutate(delAICc = AICc - minAIC) %>%
  select(-c(minAIC)) %>%
  mutate(Parameter = if_else(Parameter == "NN_umolL", "Nitrate+Nitrite",
                     if_else(Parameter == "Phosphate_umolL", "Phosphate",
                     if_else(Parameter == "Silicate_umolL", "Silicate",
                     if_else(Parameter == "meanRugosity", "Rugosity", Parameter)))))


# create AIC table for paper
write_csv(modelTable, here("Output", "PaperFigures","Model_Selection_Table.csv"))
#write_csv(modelTable_rug, here("Output", "PaperFigures","Model_Selection_Table_rugosity.csv"))

modelTable <- read_csv(here("Output", "PaperFigures","Model_Selection_Table.csv"))
#modelTable_rug <- read_csv(here("Output", "PaperFigures","Model_Selection_Table_rugosity.csv"))

mypal <- pnw_palette("Starfish", n = 2)

modelTableData <- modelTable %>%
  group_by(Y) %>%
  mutate(Y = if_else(Y == "resSpp", "% Taxon Richness (res)",
             if_else(Y == "resFEp", "% FE Richness (res)",
             if_else(Y == "resVol", "% FE Volume (res)",
             if_else(Y == "NbSpP", "% Taxon Richness",
             if_else(Y == "NbFEsP", "% FE Richness",
             if_else(Y == "Vol8D", "% FE Volume", Y)))))),
         Y = factor(Y, levels = c('% Taxon Richness', '% FE Richness', '% FE Volume',
                                  '% Taxon Richness (res)', '% FE Richness (res)', '% FE Volume (res)')),
         #Parameter = if_else(Parameter == "meanRugosity", "Rugosity", Parameter),
         Reg_Type = factor(Reg_Type, levels = c("Polynomial", "Linear")),
         Parameter = factor(Parameter,
                            levels = c("Rugosity", "Phosphate", "Nitrate+Nitrite",
                                       "pH", "Salinity", "Silicate", "Temperature",
                                       "Phosphate + Rugosity", "Nitrate+Nitrite + Rugosity",
                                       "pH + Rugosity", "Salinity + Rugosity",
                                       "Silicate + Rugosity", "Temperature + Rugosity"))) %>%
  mutate(minAICc = min(AICc),
         deltaAICc = AICc - minAICc)

AICplot <- modelTableData %>%
  filter(Y== "% Taxon Richness" | Y == "% FE Richness" | Y== "% FE Volume") %>%
  ggplot(aes(x = deltaAICc, y = fct_reorder(Parameter, desc(Parameter)), fill = Reg_Type)) +
  geom_col(position = "dodge", color = "black") +
  facet_wrap(~Y) +
  labs(x = expression(Delta*"AICc"),
       y= "Parameters",
       fill = "Regression") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 10),
        legend.position = "top",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        panel.grid = element_blank(),
        axis.text = element_text(size = 10),
        axis.title.y = element_blank(),
        axis.title.x = element_blank()) +
  scale_fill_manual(values = mypal) +
  geom_vline(xintercept = 2, linetype = "dashed", size = 0.7)
```

## Jamie's script 
```{r}
gp_nn_ms <- GP_long_control %>%
ungroup() %>% # ungroup the data
  nest(data = -c(Response_measurement, Predictor_measurement)) %>% # nest all the data measurement type
  mutate(fit = map(data, ~lm(Final~ Predictor_scaled, data = .))) #create models
gp_nn_ms

gp_nn_models <- gp_nn_ms$fit # shows you each of the models

gp_nn_results <- gp_nn_ms %>%
  mutate(coeff = map(fit, tidy),
         aic =  map(fit, glance)) %>% # R2 and others
  select(Predictor_measurement, coeff, aic) %>% # only keep the results
  unnest() %>%
  write_csv(here("Output", "Coefficients", "gp_coefficients.csv"))

gp_nn_results_cleannames <- gp_nn_results %>%
  mutate(Predictor = dplyr::recode(Predictor_measurement,
                            "max_pH" = "Max pH",
                            "max_Salinity" = "Max Salinity",
                            "max_Temperature_C" = "Max Temperature",
                            "max_NN_umolL" = "Max Nitrite + Nitrate",
                            "max_Phosphate_umolL" = "Max Phosphate",
                            "min_pH" = "Min pH",
                            "min_Salinity" = "Min Salinity",
                            "min_Temperature_C" = "Min Temperature",
                            "min_NN_umolL" = "Min Nitrite + Nitrate",
                            "min_Phosphate_umolL" = "Min Phosphate",
                            "range_pH" = "pH Range",
                            "range_Salinity" = "Salinity Range",
                            "range_Temperature_C" = "Temperature Range",
                            "range_NN_umolL" = "Nitrate + Nitrate Range",
                            "range_Phosphate_umolL" = "Phosphate Range",
                            "mean_pH" = "Mean pH",
                            "mean_Salinity" = "Mean Salinity",
                            "mean_Temperature_C" = "Mean Temperature",
                            "mean_NN_umolL" = "Mean Nitrite + Nitrate",
                            "mean_Phosphate_umolL" = "Mean Phosphate")) %>%
  mutate(delAIC = AIC - min(AIC)) %>%
  arrange(delAIC)
```

## Try with my data 
## these are all linear models for both sites 
```{r}

## for Cabral 
recruits_SGDparams_MS_C <- CommComp_recruit_seasonavg_longpredictorvars_C %>%
  nest(data = -c(Predictor_Variable)) %>% # nest all the data measurement type
  mutate(fit = map(data, ~lm(Response_Variable~ Predictor_Values, data = .))) #create models
recruits_SGDparams_MS_C

recruits_SGDparams_MS_model_C <- recruits_SGDparams_MS_C$fit # shows you each of the models

recruits_SGDparams_results_C <- recruits_SGDparams_MS_C %>%
  mutate(coeff = map(fit, tidy),
         aic =  map(fit, glance)) %>% # R2 and others
  select(Predictor_Variable, coeff, aic) %>% # only keep the results
  unnest() %>%
  write_csv(here("Outputs", "RecruitmentTiles", "modelselectionC.csv"))

## for Varari 
recruits_SGDparams_MS_V <- CommComp_recruit_seasonavg_longpredictorvars_V %>%
  nest(data = -c(Predictor_Variable)) %>% # nest all the data measurement type
  mutate(fit = map(data, ~lm(Response_Variable~ Predictor_Values, data = .))) #create models
recruits_SGDparams_MS_V

recruits_SGDparams_MS_model_V <- recruits_SGDparams_MS_V$fit # shows you each of the models

recruits_SGDparams_results_V <- recruits_SGDparams_MS_V %>%
  mutate(coeff = map(fit, tidy),
         aic =  map(fit, glance)) %>% # R2 and others
  select(Predictor_Variable, coeff, aic) %>% # only keep the results
  unnest() %>%
  write_csv(here("Outputs", "RecruitmentTiles", "modelselectionV.csv"))

#############################################################
### for Cabral 
recruits_SGDparams_results_C_cleannames <- recruits_SGDparams_results_C %>%
  mutate(Predictor = dplyr::recode(Predictor_Variable,
                            "pH_max" = "Max pH",
                            "sal_max" = "Max Salinity",
                            "temp_max" = "Max Temperature",
                            "NN_max" = "Max Nitrite + Nitrate",
                            "phos_max" = "Max Phosphate",
                            "TA_max" = "Max TA", 
                            "silicate_max" = "Max Silicate", 
                            "ammonia_max" = "Max Ammonia", 
                            "pH_min" = "Min pH",
                            "sal_min" = "Min Salinity",
                            "temp_min" = "Min Temperature",
                            "NN_min" = "Min Nitrite + Nitrate",
                            "phos_min" = "Min Phosphate",
                            "TA_min" = "Min TA", 
                            "silicate_min" = "Min Silicate", 
                            "ammonia_min" = "Min Ammonia", 
                            "pH_CV" = "pH CV",
                            "sal_CV" = "Salinity CV",
                            "temp_CV" = "Temperature CV",
                            "NN_CV" = "Nitrate + Nitrate CV",
                            "phos_CV" = "Phosphate CV",
                            "TA_CV" = "TA CV", 
                            "silicate_CV" = "Silicate CV", 
                            "ammonia_CV" = "Ammonia CV", 
                            "ph_avg" = "Mean pH",
                            "sal_avg" = "Mean Salinity",
                            "temp_avg" = "Mean Temperature",
                            "NN_avg" = "Mean Nitrite + Nitrate",
                            "phos_avg" = "Mean Phosphate", 
                            "TA_avg" = "Mean TA", 
                            "silicate_avg" = "Mean Silicate", 
                            "ammonia_avg" = "Mean Ammonia")) %>%
  mutate(delAIC = AIC - min(AIC)) %>%
  arrange(delAIC)

## for Varari 
recruits_SGDparams_results_V_cleannames <- recruits_SGDparams_results_V %>%
  mutate(Predictor = dplyr::recode(Predictor_Variable,
                            "pH_max" = "Max pH",
                            "sal_max" = "Max Salinity",
                            "temp_max" = "Max Temperature",
                            "NN_max" = "Max Nitrite + Nitrate",
                            "phos_max" = "Max Phosphate",
                            "TA_max" = "Max TA", 
                            "silicate_max" = "Max Silicate", 
                            "ammonia_max" = "Max Ammonia", 
                            "pH_min" = "Min pH",
                            "sal_min" = "Min Salinity",
                            "temp_min" = "Min Temperature",
                            "NN_min" = "Min Nitrite + Nitrate",
                            "phos_min" = "Min Phosphate",
                            "TA_min" = "Min TA", 
                            "silicate_min" = "Min Silicate", 
                            "ammonia_min" = "Min Ammonia", 
                            "pH_CV" = "pH CV",
                            "sal_CV" = "Salinity CV",
                            "temp_CV" = "Temperature CV",
                            "NN_CV" = "Nitrate + Nitrate CV",
                            "phos_CV" = "Phosphate CV",
                            "TA_CV" = "TA CV", 
                            "silicate_CV" = "Silicate CV", 
                            "ammonia_CV" = "Ammonia CV", 
                            "ph_avg" = "Mean pH",
                            "sal_avg" = "Mean Salinity",
                            "temp_avg" = "Mean Temperature",
                            "NN_avg" = "Mean Nitrite + Nitrate",
                            "phos_avg" = "Mean Phosphate", 
                            "TA_avg" = "Mean TA", 
                            "silicate_avg" = "Mean Silicate", 
                            "ammonia_avg" = "Mean Ammonia")) %>%
  mutate(delAIC = AIC - min(AIC)) %>%
  arrange(delAIC)
recruits_SGDparams_results_V_cleannames

```

## try to get nonlinear models to work 
```{r}


## for Cabral 
recruits_SGDparams_MS_C_nonlinear <- CommComp_recruit_seasonavg_longpredictorvars_C %>%
  nest(data = -c(Predictor_Variable)) %>% # nest all the data measurement type
  mutate(fit = map(data, ~lm(Response_Variable~ poly(Predictor_Values, degree=2), data = .))) #create models
recruits_SGDparams_MS_C_nonlinear

recruits_SGDparams_MS_nonlinearmodel_C <- recruits_SGDparams_MS_C_nonlinear$fit # shows you each of the models

recruits_SGDparams_results_C_nonlinear <- recruits_SGDparams_MS_C_nonlinear %>%
  mutate(coeff = map(fit, tidy),
         aic =  map(fit, glance)) %>% # R2 and others
  select(Predictor_Variable, coeff, aic) %>% # only keep the results
  unnest() %>%
  write_csv(here("Outputs", "RecruitmentTiles", "nonlinear_modelselectionC.csv"))

## for Varari 
recruits_SGDparams_MS_V_nonlinear <- CommComp_recruit_seasonavg_longpredictorvars_V %>%
  nest(data = -c(Predictor_Variable)) %>% # nest all the data measurement type
  mutate(fit = map(data, ~lm(Response_Variable~ poly(Predictor_Values, degree=2), data = .))) #create models
recruits_SGDparams_MS_V_nonlinear

recruits_SGDparams_MS_model_V_nonlinear <- recruits_SGDparams_MS_V_nonlinear$fit # shows you each of the models

recruits_SGDparams_results_V_nonlinear <- recruits_SGDparams_MS_V_nonlinear %>%
  mutate(coeff = map(fit, tidy),
         aic =  map(fit, glance)) %>% # R2 and others
  select(Predictor_Variable, coeff, aic) %>% # only keep the results
  unnest() %>%
  write_csv(here("Outputs", "RecruitmentTiles", "nonlinear_modelselectionV.csv"))

#############################################################
### for Cabral 
recruits_SGDparams_results_C_nonlinear_cleannames <- recruits_SGDparams_results_C_nonlinear %>%
  mutate(Predictor = dplyr::recode(Predictor_Variable,
                            "pH_max" = "Max pH",
                            "sal_max" = "Max Salinity",
                            "temp_max" = "Max Temperature",
                            "NN_max" = "Max Nitrite + Nitrate",
                            "phos_max" = "Max Phosphate",
                            "TA_max" = "Max TA", 
                            "silicate_max" = "Max Silicate", 
                            "ammonia_max" = "Max Ammonia", 
                            "pH_min" = "Min pH",
                            "sal_min" = "Min Salinity",
                            "temp_min" = "Min Temperature",
                            "NN_min" = "Min Nitrite + Nitrate",
                            "phos_min" = "Min Phosphate",
                            "TA_min" = "Min TA", 
                            "silicate_min" = "Min Silicate", 
                            "ammonia_min" = "Min Ammonia", 
                            "pH_CV" = "pH CV",
                            "sal_CV" = "Salinity CV",
                            "temp_CV" = "Temperature CV",
                            "NN_CV" = "Nitrate + Nitrate CV",
                            "phos_CV" = "Phosphate CV",
                            "TA_CV" = "TA CV", 
                            "silicate_CV" = "Silicate CV", 
                            "ammonia_CV" = "Ammonia CV", 
                            "ph_avg" = "Mean pH",
                            "sal_avg" = "Mean Salinity",
                            "temp_avg" = "Mean Temperature",
                            "NN_avg" = "Mean Nitrite + Nitrate",
                            "phos_avg" = "Mean Phosphate", 
                            "TA_avg" = "Mean TA", 
                            "silicate_avg" = "Mean Silicate", 
                            "ammonia_avg" = "Mean Ammonia")) %>%
  mutate(delAIC = AIC - min(AIC)) %>%
  arrange(delAIC)

## for Varari 
recruits_SGDparams_results_V_nonlinear_cleannames <- recruits_SGDparams_results_V_nonlinear %>%
  mutate(Predictor = dplyr::recode(Predictor_Variable,
                            "pH_max" = "Max pH",
                            "sal_max" = "Max Salinity",
                            "temp_max" = "Max Temperature",
                            "NN_max" = "Max Nitrite + Nitrate",
                            "phos_max" = "Max Phosphate",
                            "TA_max" = "Max TA", 
                            "silicate_max" = "Max Silicate", 
                            "ammonia_max" = "Max Ammonia", 
                            "pH_min" = "Min pH",
                            "sal_min" = "Min Salinity",
                            "temp_min" = "Min Temperature",
                            "NN_min" = "Min Nitrite + Nitrate",
                            "phos_min" = "Min Phosphate",
                            "TA_min" = "Min TA", 
                            "silicate_min" = "Min Silicate", 
                            "ammonia_min" = "Min Ammonia", 
                            "pH_CV" = "pH CV",
                            "sal_CV" = "Salinity CV",
                            "temp_CV" = "Temperature CV",
                            "NN_CV" = "Nitrate + Nitrate CV",
                            "phos_CV" = "Phosphate CV",
                            "TA_CV" = "TA CV", 
                            "silicate_CV" = "Silicate CV", 
                            "ammonia_CV" = "Ammonia CV", 
                            "ph_avg" = "Mean pH",
                            "sal_avg" = "Mean Salinity",
                            "temp_avg" = "Mean Temperature",
                            "NN_avg" = "Mean Nitrite + Nitrate",
                            "phos_avg" = "Mean Phosphate", 
                            "TA_avg" = "Mean TA", 
                            "silicate_avg" = "Mean Silicate", 
                            "ammonia_avg" = "Mean Ammonia")) %>%
  mutate(delAIC = AIC - min(AIC)) %>%
  arrange(delAIC)
```


## now join datasets with nonlinear and linear models so just have one file per site 
```{r}
joinedresultsV <- recruits_SGDparams_results_V_nonlinear_cleannames %>% 
  full_join(recruits_SGDparams_results_V_cleannames) %>% 
  arrange(delAIC) %>% 
  write_csv(here("Outputs", "RecruitmentTiles", "joinedresultsV.csv"))


joinedresultsC <- recruits_SGDparams_results_C_nonlinear_cleannames %>% 
  full_join(recruits_SGDparams_results_C_cleannames) %>% 
  arrange(delAIC) %>% 
  write_csv(here("Outputs", "RecruitmentTiles", "joinedresultsC.csv"))


```

### now plot the top few models 
```{r}
longCV <- read_csv(here("Data", "longCV.csv"))
longMEAN <- read_csv(here("Data", "longMEAN.csv"))
longMAX <- read_csv(here("Data", "longMAX.csv"))
longMIN <- read_csv(here("Data", "longMIN.csv"))

### ONLY PLOT THE TOP MODELS 
## top models : 
## Cabral: Max Temp (nonlinear) , Min TA (linear), TA CV (linear), Temp CV (nonlinear) --> all less than 0.5 deltaAIC 
## Varari: pH CV (nonlinear), max pH (linear and nonlinear), Temp CV (nonlinear) --> all less than 1 deltaAIC 

#########################
###for Cabral 

#####################
##### max temp (nonlinear)
####################
maxtemp_C <- longMAX %>% 
  drop_na(sum_total) %>% ## gets rid of seep data 
  filter(MAXenv_params == "temp_max", 
         Location== "Cabral") %>% 
  ggplot(aes(x=max_value, 
             y=log(sum_total+1))) +
  geom_point(size=3, shape=21) + 
   geom_smooth(method=lm, formula="y~poly(x,2)") + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="Max Temp") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
maxtemp_C
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Cabral", "max_temp.jpg"))

#################
## min TA (linear)
#################
min_TA <- longMIN %>% 
  drop_na(sum_total) %>% ## gets rid of seep data 
  filter(MINenv_params == "TA_min", 
         Location== "Cabral") %>% 
  ggplot(aes(x=min_value, 
             y=log(sum_total+1))) +
  geom_point(size=3, shape=21) + 
  geom_smooth(method=lm) + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="Min TA") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
min_TA
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Cabral", "min_TA.jpg"))

#########################
## Temp CV (nonlinear)
#######################
temp_CV <- longCV %>% 
  drop_na(sum_total) %>% ## gets rid of seep data 
  filter(CVenv_params == "temp_CV", 
         Location== "Cabral") %>% 
  ggplot(aes(x=CV_value, 
             y=log(sum_total+1))) +
  geom_point(size=3, shape=21) + 
  geom_smooth(method=lm, formula="y~poly(x,2)") + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="CV Temp") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
temp_CV
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Cabral", "temp_CV.jpg"))

#######################
## TA CV (linear)
######################
TA_CV <- longCV %>% 
  drop_na(sum_total) %>% ## gets rid of seep data 
  filter(CVenv_params == "TA_CV", 
         Location== "Cabral") %>% 
  ggplot(aes(x=CV_value, 
             y=log(sum_total+1))) +
  geom_point(size=3, shape=21) + 
  geom_smooth(method=lm) + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="CV TA") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
TA_CV
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Cabral", "TA_CV.jpg"))

########################
## Varari 

#########################################
## pH CV (nonlinear)
#########################################
pH_CV <- longCV %>% 
  drop_na(sum_total) %>% ## gets rid of seep data 
  filter(CVenv_params == "pH_CV", 
         Location== "Varari") %>% 
  ggplot(aes(x=CV_value, 
             y=log(sum_total+1))) +
  geom_point(size=3, shape=21) + 
   geom_smooth(method=lm, formula="y~poly(x,2)") + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="CV pH") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
pH_CV
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Varari", "pH_CV.jpg"))

#########################################
## max pH (linear and nonlinear)
#########################################
max_pH_nonlinear <- longMAX %>% 
  drop_na(sum_total) %>% ## gets rid of seep data 
  filter(MAXenv_params == "pH_max", 
         Location== "Varari") %>% 
  ggplot(aes(x=max_value, 
             y=log(sum_total+1))) +
  geom_point(size=3, shape=21) + 
   geom_smooth(method=lm, formula="y~poly(x,2)") + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="max pH") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
max_pH_nonlinear
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Varari", "max_pH_nonlinear.jpg"))

#########################################
## max pH (linear and nonlinear)
#########################################
max_pH_linear <- longMAX %>% 
  drop_na(sum_total) %>% ## gets rid of seep data 
  filter(MAXenv_params == "pH_max", 
         Location== "Varari") %>% 
  ggplot(aes(x=max_value, 
             y=log(sum_total+1))) +
  geom_point(size=3, shape=21) + 
   geom_smooth(method=lm) + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="max pH") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
max_pH_linear
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Varari", "max_pH_linear.jpg"))

#########################################
## Temp CV (nonlinear)
#########################################
CV_temp_linear <- longCV %>% 
  drop_na(sum_total) %>% ## gets rid of seep data 
  filter(CVenv_params == "temp_CV", 
         Location== "Varari") %>% 
  ggplot(aes(x=CV_value, 
             y=log(sum_total+1))) +
  geom_point(size=3, shape=21) + 
  geom_smooth(method=lm, formula="y~poly(x,2)") + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="CV temp") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
CV_temp_linear
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Varari", "temp_CV.jpg"))


#########################################



#########################################
## plot sd(TA) instead of CV 
#########################################

CommComp_recruit_seasonavg_allenvdata_C2 <- CommComp_recruit_seasonavg_allenvdata_C %>% 
  select(!c(sal_avg:ammonia_CV)) 


sd_TA_plot <- CommComp_recruit_seasonavg_allenvdata_C2 %>% 
  ggplot(aes(x=sd_TA, 
             y=sum_total)) +
  geom_point(size=3, shape=21) + 
  geom_smooth(method=lm, formula="y~poly(x,2)") + 
 # facet_wrap(Location~MAXenv_params, scales = "free") + 
  theme_bw() + 
  labs(x="value", 
       y="recruit counts", 
       title="SD TA") + 
  theme(axis.text.x=element_text(size=18), 
        axis.text.y=element_text(size=18), 
        axis.title.x=element_text(size=18),
        axis.title.y=element_text(size=18), 
        plot.title=element_text(hjust=0.5, ## centers title 
                                size=20)) 
sd_TA_plot
ggsave(here("Outputs", "RecruitmentTiles", "ModelSelection", "Cabral", "sd_TA.jpg"))





### AIC plots 

AICplots_C <- joinedresultsC %>% 
  ggplot(aes(x=delAIC, 
             y=Predictor, 
             fill=term)) + 
  geom_col(position = "dodge") +
  labs(x = "DeltaAIC",
       y= "Parameters",
       fill = "Regression Type") + 
  theme_classic() + 
  scale_fill_manual(values=pnw_palette("Sailboat")) 
AICplots_C

```


### Findings: 
TA at Cabral might have an outlier point at 2100 - so need to redo that 

## make a PCA that has a TA/pH axis (biology) and a nutrient axis... You may need to play around with what goes in (like mean, min, max, sd... is a lot... try with that, but also mean just mean and standard deviation). Look at the PCA plot and the loading arrows before you use the scores in a regression and see if the loadings make sense


## And... the next step I want to try is a PCA approach. The max pH data is pretty clean here for Varari and the TA data for Cabral is interesting... Cabral SGD as really low TA and it generally shows that the lower TA has higher recruits (which is higher amounts of SGD)... or it could be just biologically driven and the places with the most corals, which would have the highest variance in TA because of calcification, have the most recruits.... this could kind be the same story for pH, which is really biologically driven at the site
